### 2017-10-11 analysis/fps_sanity_checks.Rmd
add sanity check rmd

---
title: "FPS Sanity Checks"
output: html_notebook
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
```

```{r, message=FALSE, warning=FALSE}
new <- read_csv("../data/new_movie_lengths.csv")
old <- read_csv("../data/old_movie_lengths.csv")
```

```{r}
new %>%
  rename(new_fps = fps, new_length = length) %>%
  inner_join(old %>%
              mutate(old_fps = last.frame / length) %>%
              rename(old_length = length), 
            by = "subid") %>%
  mutate(length_close_enough = abs(new_length - old_length) <= 1) %>%
  mutate(fps_close_enough =  abs(new_fps - old_fps) <= 1) %>%
  kable
```