### 2017-12-20 analysis/face_detection_analysis.Rmd

---
title: "XS-FACE preliminary analysis"
output: html_notebook
---

Imports
```{r, message=FALSE, warning=FALSE}
library(knitr)
library(tidyverse)
```

Load CSV
```{r}
data <- read_csv('../data/face_detections.csv')
```

Summarize
```{r}
data %>%
  distinct(group, video, frame, is_face, angle) %>% # ignore > 1 face detections
  group_by(group, angle) %>%
  summarise(face_frames=sum(is_face=="True"), total=n()) %>%
  mutate(percentage=face_frames/total * 100) %>%
  arrange(angle, group) %>%
  kable
```