### 2018-01-08 analysis/mtcnn_evaluation.Rmd
new eval script

---
title: "mtcnn evaluation"
output: html_notebook
---

Load data
```{r}
mtcnn_data <- read_csv("data/final_output/mtcnn2.csv") %>%
  select(-X1)
ground_truth_data <- read_csv("data/ground_truth/ground_truth.csv")
```

Join
```{r}
combined <- mtcnn_data %>%
  rename(mtcnn_is_face = is_face) %>%
  right_join(ground_truth_data, by = c("group", "video", "frame"))
```

Eval function
```{r}
evaluate <- function(a, b) {
  if (a == "True") {
    if (a == b) return ("TP")
    else return("FP")
  }
  else {
    if (a == b) return("TN")
    else return("FN")
  }
}
```

Evaluate
```{r}
results <- combined %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, mtcnn_is_face))
```

Precison / Recall / F-score
```{r}
results %>%
  group_by(result) %>%
  summarise(count = n())
```