### 2018-01-09 analysis-validation/openpose_analysis.Rmd
reorganized analyses & integrated new detector outputs

---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(knitr)
```

```{r}
data <- read_csv("../data/openpose.csv")
```

```{r}
check <- function(nose, l, r) {
  return(!is.na(nose) & !is.na(l) & !is.na(r))
}
```

```{r}
data2 <- data %>%
  #rowwise() %>%
  mutate(is_face = check(Nose, LEye, REye))

#!is.na(Nose) && !is.na(LEye) && !is.na(REye)
```

```{r}
data2 %>%
  #filter(w>100 | is.na(w)) %>%
  #distinct(group, video, frame, is_face) %>% # ignore > 1 face detections
  group_by(group) %>%
  summarise(face_frames=sum(is_face==TRUE), total=n()) %>%
  mutate(percentage=face_frames/total * 100) %>%
  arrange(group) %>%
  kable
```

```
```{r}

```