### 2018-01-10 analysis-validation/frame_validation/check_frame_counts_post_detection.Rmd

---
title: "Check frame counts after mtcnn detection"
output: html_notebook
---

```{r}
library(tidyverse)
library(knitr)
```

Load data
```{r}
ffprobe_frame_counts <- read_csv("data/video_stats/ffprobe_frame_count.csv")
extracted_frame_counts <- read_csv("data/video_stats/extracted_frame_count.csv")
mtcnn_data <- read_csv("data/final_output/mtcnn_face_detections.csv")
# openpose_data 
```

Check ffprobe results against actual extracted frames
```{r}
ffprobe_frame_counts %>%
  all.equal(extracted_frame_counts)
```

Check mtcnn results frame counts against extracted frames
```{r}
mtcnn_data %>%
  distinct(video, frame) %>%
  group_by(video) %>%
  summarise(num_frames = n()) %>%
  all.equal(extracted_frame_counts) 
```

Check openpose results frame counts against extracted frames
```{r}

```