### 2018-01-10 scripts/ground_truth_creation/get_gold_set_images.Rmd
gold set rmd

---
title: "get_gold_set_images"
output: html_document
---

```{r}

```