### 2018-01-10 scripts/ground_truth_creation/get_gold_set_images.Rmd
viola jones edits and gold set candidates

---
title: "get_gold_set_images"
output: html_document
---

```{r}
library(tidyverse)
library(magrittr)
```

```{r}
data <- read_csv("data/final_output/mtcnn3.csv") %>%
  distinct(video, frame, .keep_all = TRUE)
```

Sample high density faces, remove selected from original data
```{r}
high_density_faces_sample <- data %>%
  group_by(video) %>%
  filter(is_face == "True") %>%
  sample_n(7)

data %<>%
  anti_join(high_density_faces_sample)
```

Choose random frames from remaining
```{r}
random_sample <- data %>%
  group_by(video) %>%
  sample_n(7)
```

Combine and save result
```{r}
gold_set_candidates <- high_density_faces_sample %>%
  rbind(random_sample) %>%
  mutate(path = paste0(video, "/image-", frame, ".jpg")) %>%
  arrange(video, frame)

write_csv(gold_set_candidates, "data/ground_truth/gold_set_candidates.csv")
```