### 2018-01-10 analysis-main/main_analyses_bll.Rmd
integrated last detectors, reran sanity checks, added face size analyses

---
title: "XSFACE Analyses"
author: "Mike Frank; continued by Bria Long"
date: "July 14, 2016; January 8, 2018"
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
    highlight: tango
    code_folding: hide
---

```{r}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = TRUE)
options(error = browser)
```

```{r}
rm(list=ls())
library(readr)
library(plyr) # need to load this for scripts, always load before
library(dplyr)
library(tidyr)
library(langcog)
library(ggplot2)
library(lubridate)
```

Note that the dataset here is now the complete data (N=36); all pass length sanity checks.

```{r}
d <- read_csv("../data/consolidated_data.csv") %>%
  mutate(posture = factor(posture), 
         orientation = factor(orientation))
demo <- read_csv("../data/demographics/demographics.csv")
#cdis <- read_csv("../data/cdis/cdi_summary.csv") ## not in repo
```

# Postures and Orientations

## Postures

```{r}
ps <- d %>%
  filter(!is.na(posture)) %>%
  group_by(subid, age.grp, posture) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(age.grp, posture) %>%
  multi_boot_standard(col = "prop.time", na.rm=TRUE)
```

```{r}
ggplot(ps, aes(x = factor(age.grp), y = mean, 
               fill = posture, pch = posture)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), 
                 position = position_dodge(width = .9)) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  scale_fill_solarized() + 
  theme_bw()
```

## Orientations

```{r}
os <- d %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, subid, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(age.grp, orientation) %>%
  multi_boot_standard(col = "prop.time")
```

```{r}
ggplot(os, aes(x = factor(age.grp), y = mean, 
               fill = orientation, pch = orientation)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), 
                 position = position_dodge(width = .9)) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  scale_fill_solarized() + 
  theme_bw()
```

## Postures and orientations together

Data don't look interpretable typically when sliced this fine.

```{r}
pos <- d %>%
  filter(!is.na(orientation) & !is.na(posture)) %>%
  group_by(age.grp, subid, orientation, posture) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  group_by(age.grp, subid) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(age.grp, orientation, posture) %>%
  multi_boot_standard(col = "prop.time")
```

```{r}
ggplot(pos, aes(x = factor(age.grp), y = mean, 
               fill = posture)) + 
  geom_bar(stat = "identity", position = "dodge") + 
  geom_linerange(aes(ymin=ci_lower, ymax=ci_upper), 
                 position = position_dodge(width = .9)) +
  facet_grid(.~ orientation) + 
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  scale_fill_solarized() + 
  theme_bw()
```

# Face detections

Using automated face detection results.

## Basic detections descriptives

```{r}

mss <- d %>%
  group_by(age.grp, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

ms <- mss %>%
  multi_boot_standard(col = "face")

ggplot(mss, aes(x = age.grp, y = face)) + 
  geom_jitter(width = .25) + 
  geom_pointrange(data = ms, 
                 aes(x = age.grp, 
                     y = mean, 
                     ymin = ci_lower, 
                     ymax = ci_upper), col = "red") +
  geom_line(data = ms, 
             aes(x = age.grp, 
                     y = mean), col = "red") + 
  scale_x_continuous(breaks=c(8,12,16)) +
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

And with continuous age. This is probably the simplest option.

```{r}
mss <- d %>%
  filter(!is.na(age.at.test)) %>%
  group_by(subid, age.at.test) %>%
  summarise(face = mean(face)) %>%
  left_join(demo) %>%
  # mutate(len = as.numeric(lubridate::as.duration(lubridate::ms(len)))) ## error here, used simpler calculation.
  mutate(len = as.numeric(len_seconds/60))

ggplot(mss, aes(x = age.at.test, y = face)) + 
  geom_jitter(width = .25, aes(size = len)) + 
  geom_smooth() + 
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_size_continuous(name = "Length (s)") + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

Boxplot, though it's not as good.

```{r}
ggplot(mss, aes(x = factor(age.grp), y = face)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

Stats. Standard ANOVA (violates distributional assumptions).

```{r}
mod <- lm(face ~ factor(age.grp), data = mss)
summary(mod)
anova(mod)
```

Now a non-parametric test.

```{r}
kruskal.test(face ~ factor(age.grp), data = mss)
```

No real signal that the (predicted) U-shaped pattern is reliable, unfortunately.

## Detections timecourse

Do a 10s moving window average.

```{r}
d <- d %>%
  group_by(subid) %>%
  filter(!is.na(face)) %>%
  mutate(face_avg = zoo::rollapply(data = face, 
                                   width = 30*10, # 10s window
                                   FUN = mean, 
                                   align = "right", 
                                   fill = NA, 
                                   na.rm = TRUE))

ggplot(d, aes(x = frame, y = face_avg)) + 
  geom_line() + 
  facet_wrap(~subid) + 
  theme_bw()
```

# Detections by posture and orientation

## Posture

```{r}
mss <- d %>%
  filter(!is.na(posture)) %>%
  filter(!is.na(face)) %>%
  group_by(age.grp, posture, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

# ms <- mss %>%
#   multi_boot_standard(col = "face")

ggplot(mss, aes(x = factor(age.grp), y = face, col = posture)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

## Orientation

```{r}
mss <- d %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, orientation, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

# ms <- mss %>%
#   multi_boot_standard(col = "face")

ggplot(mss, aes(x = factor(age.grp), y = face, col = orientation)) + 
  geom_boxplot() + 
  ylab("Proportion Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

## Both

Not really enough data to say much when binned this finely.

```{r}
mss <- d %>%
  filter(!is.na(posture) & !is.na(orientation)) %>%
  group_by(age.grp, posture, orientation, subid) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp)

# ms <- mss %>%
#   multi_boot_standard(col = "face")

ggplot(mss, aes(x = factor(age.grp), y = face, col = posture)) +
  geom_boxplot() +
  ylab("Proportion Face Detections") +
  xlab("Age (months)") +
  facet_grid(~orientation) +
  ylim(c(0,.2)) +
  theme_bw()
```
# Face area: descriptives, posture and orientation
## When detected, how much of the field of view did faces occupy? (% of screen)

```{r}
screenPixels=720 * 480
d$faceArea = d$w * d$h /screenPixels

mss <- d %>%
  group_by(age.grp, subid) %>%
  summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
  group_by(age.grp)

ms <- mss %>%
  multi_boot_standard(col = "faceArea")

ggplot(mss, aes(x = age.grp, y = faceArea)) + 
  geom_jitter(width = .25) + 
  geom_pointrange(data = ms, 
                 aes(x = age.grp, 
                     y = mean, 
                     ymin = ci_lower, 
                     ymax = ci_upper), col = "red") +
  geom_line(data = ms, 
             aes(x = age.grp, 
                     y = mean), col = "red") + 
  scale_x_continuous(breaks=c(8,12,16)) +
  ylab("% ScreenArea of Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

## Face size as a function of continous age

```{r}
mssFaceArea <- d %>%
  filter(!is.na(age.at.test)) %>%
  group_by(subid, age.at.test) %>%
  summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
  left_join(demo) %>%
  mutate(len = as.numeric(len_seconds/60))

ggplot(mssFaceArea, aes(x = age.at.test, y = faceArea)) + 
  geom_jitter(width = .25, aes(size = len)) + 
  geom_smooth() + 
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_size_continuous(name = "Length (s)") + 
  ylab("% Area of Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

## Face size by infant's posture

```{r}
mss <- d %>%
  filter(!is.na(posture)) %>%
  filter(!is.na(face)) %>%
  group_by(age.grp, posture, subid) %>%
  summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
  group_by(age.grp)

# ms <- mss %>%
#   multi_boot_standard(col = "faceArea")

ggplot(mss, aes(x = factor(age.grp), y = faceArea, col = posture)) + 
  geom_boxplot() + 
  ylab("% Area of Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

## Face Area by Caregiver Orientation
```{r}
mss <- d %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, orientation, subid) %>%
  summarise(faceArea = mean(faceArea, na.rm="TRUE")) %>%
  group_by(age.grp)

# ms <- mss %>%
#   multi_boot_standard(col = "faceArea")

ggplot(mss, aes(x = factor(age.grp), y = faceArea, col = orientation)) + 
  geom_boxplot() + 
  ylab("% Area of Face Detections") + 
  xlab("Age (months)") + 
  theme_bw() 
```

# Naming data

Need some helper functions to merge in namings.

```{r}
source("helper.R")
```

```{r}
namings <- d %>%
  group_by(subid) %>%
  do(summarize.naming(.)) %>%
  left_join(demo %>% select(subid, gender, age.at.test, len_seconds))
```

First examine: how many namings do we have, generally?

```{r}
naming_rate <- namings %>%
  group_by(subid, age.grp, age.at.test) %>%
  summarise(num = n(),
            len = len_seconds[1]) %>%
  mutate(rate = num / as.numeric(len / 60)) # bll edits here to wrangle dates

ggplot(naming_rate,
       aes(x = age.at.test, y = rate)) +
  geom_point() +
  geom_smooth(span = 1) +
  ylab("Namings per minute") +
  xlab("Age (months)") +
  theme_bw()
```

Break down by novel and familiar.

```{r}
naming_rate_familiarity <- namings %>%
  group_by(subid, familiarity, age.at.test) %>%
  summarise(num = n(),
            len = len_seconds[1]) %>%
  mutate(rate = num / as.numeric(len / 60)) # bll edits here to wrangle dates

ggplot(naming_rate_familiarity,
       aes(x = age.at.test, y = rate, col = familiarity)) +
  geom_point() +
  geom_smooth(span = 1) +
  ylab("Namings per minute") +
  xlab("Age (months)") +
  theme_bw()
```

Item effects just for fun.

```{r}
namings %>%
  group_by(subid, name, familiarity, age.at.test) %>%
  summarise(num = n(),
            len = len_seconds[1]) %>%
  mutate(rate = num / as.numeric(len / 60)) %>%  # bll edits here to wrangle dates
  ggplot(aes(x = age.at.test, y = rate, col = name)) +
  geom_point() +
  facet_grid(.~familiarity) +
  geom_smooth(span = 1) +
  ylab("Namings per minute") +
  xlab("Age (months)") +
  theme_bw()
```

# Namings and face detections

## Are there more faces before or after naming instances?

```{r}
sb <- plyr::ddply(d,~subid, summarize.naming, window=c(-2,0)) ## used 2 second window as in ms draft
sa <- plyr::ddply(d,~subid, summarize.naming, window=c(0,2))

sb$window <- "before"
sa$window <- "after"
s <- rbind(sb,sa)
s$window <- factor(s$window)

# not used, commented out
# s$first.instance <- s$naming.instance == 1
# s$log.instance <- round(log10(s$naming.instance)*4)/4
# s$binned.instance <- 10^(round(log10(s$naming.instance)*4)/4)
# s$face.logical <- s$face != 0

## try plotting individuals
mss <- ddply(s,~subid+age.grp+window,summarise,
               face = mean(face, na.rm="TRUE"))
```

```{r}
  qplot(age.grp,face,colour=window,
        xlab="Age (Months)",ylab="Proportion faces detected in window",
        data=mss,geom=c("point")) +
    scale_x_continuous(breaks=c(8,12,16)) +
    geom_smooth()
```

## How do face detections change with sequence of naming instances?

```{r}
ms <- namings %>%
  group_by(age.grp, subid, naming.instance, familiarity, name) %>%
  summarise(face = mean(face)) %>%
  group_by(age.grp, naming.instance, familiarity) %>%
  multi_boot_standard(col = "face", na.rm = TRUE)
```

```{r}
ggplot(ms,
       aes(x = naming.instance, y = mean, col = familiarity)) +
  geom_point() +
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper)) +
  geom_smooth(se = FALSE) +
  facet_grid(~age.grp) +
  xlim(c(0,20)) +
  scale_color_solarized() +
  theme_bw()
```

## Now calculate face detection differences from base rate

```{r}

  n <- ddply(s,~subid+age.grp,summarise,
             face = mean(face, na.rm="TRUE"))

  n$br <- ddply(d,~subid+age.grp,summarise,
              face=mean(face, na.rm="TRUE"))$face

  n$br.diff <- n$face - n$br

#   pdf("~/Projects/xsface/writeup/figures/naming_faces_diff.pdf",width=4,height=3)
  qplot(age.grp,br.diff,position=position_jitter(.2),
        data=subset(n,age.grp>4),xlab="Age (Months)",ylab="Face detections relative to base rate") +
    scale_x_continuous(breaks=c(8,12,16)) +
    geom_abline(slope=0,intercept=0,lty=2) +
    stat_summary(fun.data="mean_cl_boot",color="red")
#   dev.off()
```

<!-- # CDIs, namings, and face detections -->

<!-- ```{r} -->
<!-- cdi_namings <- left_join(naming_rate, cdis) -->

<!-- qplot(comprehension, facets= .~ age.grp, data = cdi_namings) + -->
<!--   theme_bw() -->

<!-- qplot(production, facets= .~ age.grp, data = cdi_namings) + -->
<!--   theme_bw() -->

<!-- ``` -->

<!-- ## CDIs and namings -->

<!-- Just out of curiousity, are there any major differences in terms of CDIs by naming rate? (Would help to establish the validity of the naming rate analyses). -->

<!-- ```{r} -->
<!-- ggplot(cdi_namings, -->
<!--        aes(x = rate, y = comprehension)) + -->
<!--   geom_point() + -->
<!--   geom_smooth(method = "lm", se = FALSE) + -->
<!--   facet_grid(.~age.grp) + -->
<!--   xlab("Namings per minute") + -->
<!--   ylab("CDI Summed Comprehension Vocabulary") + -->
<!--   theme_bw() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- cdis <- read_csv("../data/cdis/cdi_summary.csv") -->
<!-- left_join(naming_rate, cdis) %>% -->
<!--   ggplot(aes(x = rate, y = production)) + -->
<!--   geom_point() + -->
<!--   geom_smooth(method = "lm", se = FALSE) + -->
<!--   facet_grid(.~age.grp) + -->
<!--   xlab("Namings per minute") + -->
<!--   ylab("CDI Summed Production Vocabulary") + -->
<!--   theme_bw() -->
<!-- ``` -->