### 2018-01-13 writeup-cogsci-copy/extra.Rmd

---
title: "Untitled"
author: "Bria Long"
date: "1/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}

Children (N = `r length(unique(joint$session_id))`, *M* = `r round(mean(d$age[d$trial_num==2]),2)` years, range 4-10 years) participated in an iPad-based drawing game where they were prompted with a verbal cue to draw one of sixteen familiar objects (e.g., “Can you draw a cup?”). Children drew each object category for 30 seconds, after which they were prompted to either make another drawing or to stop drawing altogether. Afterwards, a group of naive adults (N = `r length(unique(r$workerid))`) guessed the identity of each drawn object (`r length(joint$imNameShort)` drawings).

Number of drawings: `r length(joint$imNameShort)`

Number of drawers: `r length(unique(joint$session_id))`

Average age of drawers: `r round(mean(d$age[d$trial_num==2]),2)`

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.