### 2018-01-24 data/ground_truth/merge_ground_truth.Rmd
new detector stuff

---
title: "merge_ground_truth"
output: html_document
---

```{r}
ground_truth_mtcnn <- read_csv("intermediate/ground_truth_mtcnn.csv")
ground_truth_openpose <- read_csv("intermediate/ground_truth_openpose.csv")
candidates_mtcnn <- read_csv("intermediate/candidates_mtcnn.csv")
candidates_openpose <- read_csv("intermediate/candidates_openpose.csv")
```

```{r}
df1 <- ground_truth_mtcnn %>%
  inner_join(candidates_mtcnn %>%
               select(group, video, frame, sample_type), by = c("group", "video", "frame")) %>%
  mutate(detector_sample = "mtcnn")

df2 <- ground_truth_openpose %>%
  inner_join(candidates_openpose %>%
               select(group, video, frame, sample_type), by = c("group", "video", "frame")) %>%
  mutate(detector_sample = "openpose")

write_csv(rbind(df1, df2), "ground_truth_faces.csv")
```