### 2018-01-28 writeup-cogsci/cogsci2018_faceOnly.tex

% Template for Cogsci submission with R Markdown

% Stuff changed from original Markdown PLOS Template
\documentclass[10pt, letterpaper]{article}

\usepackage{cogsci}
\usepackage{pslatex}
\usepackage{float}
\usepackage{caption}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}

% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% hyperref package, useful for hyperlinks
\usepackage{hyperref}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% Sweave(-like)
\usepackage{fancyvrb}
\DefineVerbatimEnvironment{Sinput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{Soutput}{Verbatim}{}
\DefineVerbatimEnvironment{Scode}{Verbatim}{fontshape=sl}
\newenvironment{Schunk}{}{}
\DefineVerbatimEnvironment{Code}{Verbatim}{}
\DefineVerbatimEnvironment{CodeInput}{Verbatim}{fontshape=sl}
\DefineVerbatimEnvironment{CodeOutput}{Verbatim}{}
\newenvironment{CodeChunk}{}{}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace}
%\doublespacing

% % Text layout
% \topmargin 0.0cm
% \oddsidemargin 0.5cm
% \evensidemargin 0.5cm
% \textwidth 16cm
% \textheight 21cm

\title{Postural developments mediate children's visual access to social
information}

\author{{\large \bf Alessandro Sanchez} \\ \texttt{sanchez7@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Bria Long} \\ \texttt{bria@stanford.edu} \\ Department of Psychology \\ Stanford University \And {\large \bf Ally Kraus} \\ \texttt{allison.m.kraus@gmail.com} \\ Department of Psychology \\ Stanford University \And {\large \bf Michael C. Frank} \\ \texttt{mcfrank@stanford.edu} \\ Department of Psychology \\ Stanford University}

\begin{document}

\maketitle

\begin{abstract}
The ability to process social cues--including eye gaze--is a critical
component of children's early language and cognitive development.
However, as children reach their first birthday, they begin to locomote
themselves, walking and exploring their visual environment in an
entirely new way. How do these postural and locomotive changes affect
children's access to the social information relevant for word-learning?
Here, we explore this question by using head-mounted cameras to record
infants' (8-16 months of age) egocentric visual perspective and use
computer vision algorithms to detect the proportion of faces in infants'
environments. We find that infants' posture and orientation to their
caregiver largely mediate infants' access to faces, suggesting that
these postural and locomotive developments facilitate infants' emerging
linguistic and social capacities. Broadly, we suggest that the combined
use of head-mounted cameras and the application of new computer vision
techniques is a promising avenue for understanding the statistics of
infants' visual and linguistic experience.

\textbf{Keywords:}
social cognition; face-perception; infancy; locomotion; head-cameras;
deep learning
\end{abstract}

\section{Introduction}\label{introduction}

Children are deeply engaged in learning from others (Csibra \& Gergely,
2009; Meltzoff, 2007) and attend to the social information in their
environment from their earliest days. Even newborns tend to prefer to
look at faces with direct vs.~averted gaze (Farroni, Csibra, Simion, \&
Johnson, 2002) and follow overt gaze shifts by 6 months of age
(Gredebäck, Fikke, \& Melinder, 2010; Gredebäck, Theuring, Hauf, \&
Kenward, 2008). And when infants free-view videos, they tend to look
mostly at faces at the expense of other visual information--though older
infants start to look towards people's hands and the actions they are
performing (Frank, Amso, \& Johnson, 2014; Frank, Vul, \& Saxe, 2012).

However, as children are learning from others around them, their view of
the world is also radically changing (K. Adolph \& Berger, 2007).
Infants' motor abilities improve dramatically near the end of the first
year of life, allowing them to locomote independently. These motor
changes have drastic consequences for what children see; crawling and
walking infants have radically different views of the world. For
example, during spontaneous play in a laboratory playroom, toddlers are
more likely to look at the floor while crawling than while walking (J.
Franchak, Kretch, Soska, \& Adolph, 2011); in general, walking infants
tend to have full visual access to their environment and the people in
it, while crawling infants do not (K. S. Kretch, Franchak, \& Adolph,
2014).

One possibility is that these motor improvements have strong
developmental cascades, impacting children's emerging social, cognitive,
and linguistic abilities (Iverson, 2010). Indeed, these postural changes
also impact how children interact with their mothers; walking
(vs.~crawling) infants make different kinds of object-related bids for
attention from their mothers and tend to hear more action directed
statements (e.g., ``open it'') (Karasik, Tamis-LeMonda, \& Adolph,
2014). Further, in an observational study, Walle \& Campos (2014) found
that children who were able to walk had both higher receptive and
productive vocabularies. On their account, children's ability to stand
and independently locomote may fundamentally change their ability to
access social information (e.g., faces, gaze) relative to children who
are still crawling and sitting. In other words, the ability to walk and
move around independently and subsequently access more detailed social
information may accelerate infants' ability to learn the people in their
environments.

Recent technological developments allow for testing of this hypothesis
by documenting the experiences of infants and children from their own
perspective. By using head-mounted cameras, researchers have begun to
document visual experiences of infants and children --- which even for
walking children are radically different from the adult perspective (and
not easily predicted by our own adult intuitions) (Clerkin, Hart, Rehg,
Yu, \& Smith, 2017; J. Franchak et al., 2011; Yoshida \& Smith, 2008).
Children's views tend to be restricted and to be dominated by objects
and hands much more than that of adults (Yoshida \& Smith, 2008), and
both computational and empirical work suggest that this restricted
viewpoint may be more effective for learning objects and their labels
than the comparable adult perspective (Bambach, Crandall, Smith, \& Yu,
2017; D. Yurovsky, Smith, \& Yu, in press). Further, recent work also
suggests dramatic changes in the child's perspective over the first two
years of life, as views transition from primarily containing a close up
view of faces to capturing views of hands paired with the objects they
are acting on (Fausey, Jayaraman, \& Smith, 2016).

Here, we directly ask whether postural and locomotive developments
change the availability of social information--the presence of faces and
hands. To do so, we recorded the visual experience of a group of infants
in three age ranges (8,12, and 16 months) using head-mounted cameras
during a brief laboratory free-play session; children's posture and
orientation relative to their caregiver were also recorded from a
third-person perspective and hand-annotated. We then capitalize on
recent improvements in face detection algorithms (Cao, Simon, Wei, \&
Sheikh, 2017; K. Zhang, Zhang, Li, \& Qiao, 2016) to analyze the
frequencies of faces and wrists/hands in the child's visual environment,
both overall and relative to naming events by their caregivers. We
hypothesized that there would be differential access to social
information based on children's postural developments: that as infant's
learned how to crawl, they would see relatively fewer faces as they
would be looking primarily at the ground, and that as infant's learned
how to walk and locomote independently, that they will experience a much
greater portion of social information in their environment.

\section{Methods}\label{methods}

\subsection{Participants}\label{participants}

Our final sample consisted of 36 infants and children, with 12
participants in three age groups: 8 months (6 females), 12 months (7
females), and 16 months (6 females). Participants were recruited from
the surrounding community via state birth records, had no documented
disabilities, and were reported to hear at least 80\% English at home.
Demographics and exclusion rates are given in Table \ref{tab:pop}.

\begin{table}[H]
\centering
\begin{tabular}{rrrrrr}
  \hline
 Group & N & \% incl. & Mean age & Videos length (min) \\ 
  \hline
   8 mo. &   12 & 0.46 & 8.71 & 14.41 \\ 
   12 mo. &  12 & 0.40 & 12.62 & 13.48 \\ 
   16 mo. &  12 & 0.31 & 16.29 & 15.00\\ 
   \hline
\end{tabular}
\caption{\label{tab:pop} Demographics by age group.}
\end{table}

To obtain this final sample, we tested 95 children, excluding 59
children for the following reasons: 20 for technical issues related to
the headcam, 15 for failing to wear the headcam, 10 for fewer than 4
minutes of headcam footage, 5 for having multiple adults present, 5 for
missing Communicative Development Inventory (CDI) data, 2 for missing
scene camera footage, 1 for fussiness, and one excluded for sample
symmetry. All inclusion decisions were made independent of the results
of subsequent analyses. Some of these data were also used in Frank,
Simmons, Yurovsky, \& Pusiol (2013).

\subsection{Head-mounted camera}\label{head-mounted-camera}

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/headcam-1}

}

\caption[Field of view for three different headcam configurations, with the device we used in the middle]{Field of view for three different headcam configurations, with the device we used in the middle.}\label{fig:headcam}
\end{figure}
\end{CodeChunk}

We used a small, head-mounted camera (``headcam'') that was constructed
from a MD80 model camera attached to a soft elastic headband. Videos
captured by the headcam were 720x480 pixels with 25 frames per
second.Detailed instructions for creating this headcam can be found at
\url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.
A fisheye lens was attached to the camera to increase the view angle
from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
\ref{fig:headcam}, left).

Even with the fish-eye lens, the vertical field of view of the camera is
still considerably reduced compared to the child's approximate vertical
field of view, which spans around 100--120\(^{\circ}\) in the vertical
dimension by 6-7 months of age (Cummings, Van Hof-Van Duin, Mayer,
Hansen, \& Fulton, 1988; Mayer, Fulton, \& Cummings, 1988). As we were
primarily interested in the presence of faces in the child's field of
view, we chose to orient the camera upwards to capture the entirety of
the child's upper visual field where the child is likely to see the
faces of adults around them. This choice allowed us to maximize our
chances of capturing faces that the child would have seen during the
play session.

\subsection{Procedure}\label{procedure}

First, all parents signed consent documents in a waiting room where
children were fitted with the headcam. After the child was comfortable
in the waiting room and with the experimenter, the experimenter placed
the headcam on the child's head. If the child was uninterested in
wearing the headcam or tried to take it off, the experimenter presented
engaging toys to try to draw the child's focus away from the headcam
(Yoshida \& Smith, 2008).

After the child was comfortable wearing the headcam, the child and
caregiver were shown to a playroom for the free-play session--the focus
of the current study. Parents were shown a box containing three pairs of
novel and familiar objects (e.g., a ball and a microfiber duster, named
a ``zem''), and were instructed to play with the object pairs with their
child one at a time, ``as they typically would.'' All parents confirmed
that their child had not previously seen the novel toys and were
instructed to use the novel labels to refer to the novel toys.

The experimenter then left the playroom for approximately 15 minutes,
during which a tripod-mounted camera in the corner of the room recorded
the session and the headcam captured video from the child's perspective.

\subsection{Data Processing and
Annotation}\label{data-processing-and-annotation}

\begin{figure*}
\includegraphics[width=6in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

Headcam videos were trimmed such that they excluded the instruction
phase when the experimenter was in the room and were automatically
synchronized with the tripod-mounted videos using FinalCut Pro Software.
These sessions yielded videos of 516 minutes (almost a milion frames),
with an average video length of 8.6 minutes (min = 4.53, max = 19.35).

\subsubsection{Posture and Orientation
Annotation}\label{posture-and-orientation-annotation}

We created a set of custom annotations that described the child's
physical posture (e.g.~standing) and the orientation of the caregiver
relative to the child (e.g.~far away). The child's posture was
categorized as being held/carried, prone (crawling or lying), sitting,
or standing. The caregiver's orientation was characterized as being
close to the child, far from the child, and a global category of
caregiver behind the child. For the first two annotations (close/far
from the child), the caregiver could either be to the the front or to
the side of the child. All annotations were made by a trained coder
using the OpenSHAPA/Datavyu software (K. Adolph, Gilmore, Freeman,
Sanderson, \& Millman, 2012), and times when the child was out of view
of the tripod camera were marked as uncodable and were excluded from
these annotations.

\subsection{Face Detection}\label{face-detection}

We used three face detection systems to explore infants' changing access
to social information and to avoid the cost of hand-annotating every
frame. We first measured the performance of the most commonly-used and
widely available face detection algorithm, Viola-Jones. We used this as
a benchmark for performance, and while it can achieve impressive
accuracy in some situations, it is notoriously bad at dealing with
occluded faces (Scheirer, Anthony, Nakayama, \& Cox, 2014). We next
capitalized on recent improvements in computer vision, testing the
performance of two state-of-the-art face detectors that both made use of
Convolutional Neural Networks (CNNs) to extract face information. The
first algorithm was specifically optimized for face detection, and the
second algorithm was optimized to extract information about the position
of agent's bodyparts.

The second of these algorithms (OpenPose) also captures information
about the ``skeleton'' of people in a given frame, outputting
information about the entire set of body parts that were detected. We
analyze wrist detections as a good proxy for the presence of a hand. We
do so first because we did not capture the entire visual field
experienced by the infants/children. Thus, even though the video may
only show the presence of a wrist, the caregiver's hand may be within
the child's field of view. Second, we do so because hands are often
occluded by objects when caregivers are interacting with children, yet
still visually accessible by the child and part of their joint
interaction. For example, if a caregiver was holding a toy and
presenting it to the infant, their wrists may be visible but their hand
may not necessarily be (as it is occluded by the toy).

\subsubsection{Face Detection
Algorithms}\label{face-detection-algorithms}

The first face detection system made use of a series of Haar
feature-based cascade classifiers (Viola \& Jones, 2004) applied to each
individual frame. This detector provided information about the presence
of a face as well as its size and position.

The second algorithm was based on the work by K. Zhang et al. (2016)
using multi-task cascaded convolutional neural networks (MTCNNs). The
system was built using a novel cascaded CNN-based framework for joint
detection and alignment, built to perform well in real-world
environments where varying illuminations and occlusions are present. We
used a Tensorflow implementation of this algorithm provided by
(\url{https://github.com/davidsandberg/facenet}). Like Viola-Jones, this
detector provided information about the presence of a face as well as
its size and position.

The third algorithm was a CNN-based pose detector that provided the
locations of 18 body parts (ears, nose, wrists, etc.) called OpenPose
(Cao et al., 2017; Simon, Joo, Matthews, \& Sheikh, 2017; Wei,
Ramakrishna, Kanade, \& Sheikh, 2016) avaliable at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose. For the purposes of
our investigation we only made use of the body parts relevant to the
face (ears, eyes, nose). We operationalized face detections as any
frames containing a nose, as any half of a face necessarily contains a
nose. We operationalized wrist detections as any frames containing
either the left or the right wrist.

\subsubsection{Detector evaluation}\label{detector-evaluation}

To evaluate face detector performance, we needed to construct a ``gold
set'' of labeled frames that did not carry any biases that might be due
to the relatively rare appearance of faces in the dataset. We thus
hand-labeled two samples: a sample containing a high density of faces
(half reported by MTCNN, half by OpenPose) and a random sample from the
remaining frames. Each sample was comprised of an equal number of frames
taken from each participant (i.e., video). Faces were classifed as
present in a frame if at least half of the face was showing. Precision
(hits / hits + false alarms), recall (hits / hits + misses), and F-score
(harmonic mean of precision and recall) were calculated for all
detectors and are reported in Table 2. For wrist detections, the ``gold
set'' was constructed in the same way, except that frames with a high
density of wrists were taken by wrist detections made by OpenPose.

For face detection, MTCNN outperformed OpenPose when looking at the
composite F-score (0.89 vs.~0.83); thus the face presence analyses to
follow all result from MTCNN detections. We found that although MTCNN
and OpenPose perform comparably with the random sample, MTCNN had the
edge with the high density sample (specifically looking at precision),
suggesting that OpenPose generated more false positives than MTCNN.
ViolaJones performed quite poorly, missing many of the faces in the
randomly sampled frames.

For wrist detection, OpenPose performed moderately well (f = 0.72) with
relatively high precision, but low recall on the randomly sampled set of
frames (see Table 2). We thus analyze wrist detections, with the caveat
that we are likely underestimating the proportion of wrists overall in
the datatset.

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
  \hline
   Algorithm & Sample\ Type & P & R & F \\ 
  \hline
  MTCNN-Faces & High density & 0.89 & 0.92 & \textbf{0.90} \\ 
  MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
  OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
  OpenPose-Faces & Random & 0.72 & 0.80 & \textbf{0.76} \\ 
  ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
  ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
  OpenPose-Wrists & High density & 0.64 & 1.00 & 0.78 \\ 
  OpenPose-Wrists & Random & 0.76 & 0.26 & 0.39 \\ 
   \hline
\end{tabular}
\caption{Detector performance on both high density samples (where proportion of faces detected was high) and random samples (where frames were randomly selected). P, R, and F denote precision, recall, and F-score, respectively.} 
\end{table}

\section{Results}\label{results}

First, we report developmental shifts in infants' posture and their
orientation relative to their caregiver. Then, we explore how these
changes influence children's visual access to faces across this
developmental time range. Finally, we explore how these changes impact
the accessibility of faces during labeling events.

\subsection{Changes in Posture and
Orientation}\label{changes-in-posture-and-orientation}

\begin{CodeChunk}
\begin{figure}[h]

{\centering \includegraphics{figs/posture-1}

}

\caption[Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver]{Proportion time that infants in each age group spent in each posture/orientation relative to their caregiver.}\label{fig:posture}
\end{figure}
\end{CodeChunk}

We noted characteristic changes in infants' posture and orientation
across this developmental time range. The proportion of time infants
spent sitting decreased with age, and the proportion of time infants
spent standing increased with age. Both 8-month-olds and 12-month-olds
spent equivalent amounts of time either lying/crawling, which was
markedly decreased in the 16-month-olds, who spent most of their time
either sitting or standing (see Figure \ref{fig:posture}). We also
observed characteristic changes in children's orientation relative to
their caregivers: the 8-month-olds spent more time with their caregiver
behind them supporting their sitting positions (see Figure
\ref{fig:posture}).

\section{Changes in Access to Faces}\label{changes-in-access-to-faces}

\begin{CodeChunk}
\begin{figure}[h]

{\centering \includegraphics{figs/detByAge-1}

}

\caption[Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right panel) as a function of child's age]{Proportion of faces detected by the MTCNN model (left) and wrists detected by the OpenPose model (right panel) as a function of child's age. Larger dots indicate children who had longer playsessions and thus for whom there was more data}\label{fig:detByAge}
\end{figure}
\end{CodeChunk}

We first examined the proportion of face detections across age; a full
summary can be seen in Figure \ref{fig:detByAge}. We observed a slight
U-shaped function both when analyzing the output of the MTCNN and
OpenPose detectors, such that 12-month-olds appeared to experience
slighly fewer faces faces than 8 or 16-month-olds.

However, we found that any age related effects were much smaller
compared to the impact of postural and locomotive changes on children's
visual access to faces. Children's posture was a major factor both in
how many faces they saw during the play session. Infants who were
sitting saw more faces than infants who were lying down or being
carried, while infants who were standing saw the most faces. We also
examined how the child's orientation relative to their caregiver
impacted their visual access to faces: children who were far away from
their caregiver were more likely to see faces than children who were
close to their caregiver; this was true within all age groups and for
both face detectors.

\begin{CodeChunk}
\begin{figure*}[h]

{\centering \includegraphics{figs/detByPosOrient-1}

}

\caption[Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant]{Proportion face detections as a function of children's posture (top panel) and orientation (bottom panel), binned by the age of the participant.}\label{fig:detByPosOrient}
\end{figure*}
\end{CodeChunk}

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & -5.2469 & 0.0584 & -89.88 & 0.0000 \\ 
  Age & 0.0847 & 0.0041 & 20.89 & 0.0000 \\ 
  Prone & 0.2015 & 0.0564 & 3.57 & 0.0004 \\ 
  Sit & 1.4053 & 0.0541 & 25.98 & 0.0000 \\ 
  Stand & 1.4272 & 0.0542 & 26.33 & 0.0000 \\ 
  Close & 1.8239 & 0.0230 & 79.17 & 0.0000 \\ 
  Far & 2.5479 & 0.0239 & 106.42 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

\begin{table}[H]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & -5.0818 & 0.0776 & -65.46 & 0.0000 \\ 
  Age & 0.0564 & 0.0050 & 11.35 & 0.0000 \\ 
  Prone & 0.9499 & 0.0774 & 12.27 & 0.0000 \\ 
  Sit & 1.7282 & 0.0758 & 22.79 & 0.0000 \\ 
  Stand & 1.6618 & 0.0760 & 21.88 & 0.0000 \\ 
  Close & 0.7472 & 0.0188 & 39.81 & 0.0000 \\ 
  Far & 1.6357 & 0.0200 & 81.61 & 0.0000 \\ 
   \hline
\end{tabular}
\end{table}

To formalize these observations, we fit a generalized linear model to
the proportion faces infants saw in each posture and orientation (as
detected by MTCNN), with participant's age, orientation, and posture as
independent variables. A summary of the coefficients of a model with
only main effects (and no interactions) model can be found in Table 3.
When we did include interaction terms between age, posture, and
orientation, age no longer remained a significant predictor (b = -15.84,
SE = 113.98, z = -0.14 , p =0.89). We found the same pattern of results
for the OpenPose face detections, omitted here for brevity. We also
found the same pattern with respect to infant's visual access to hands
(see Table 4); including interaction terms also reduced any main effect
of age on the proportion of hands detected (b = -15.7, SE = 113.98, z =
-0.14 , p =0.89). Thus, these results suggest that infants access to
social information is heavily infleunced by their postural and
locomotive development.

\begin{CodeChunk}
\begin{figure}[H]

{\centering \includegraphics{figs/detByNaming-1}

}

\caption[Proportion face detections around a naming instance ('Look, a Zem']{Proportion face detections around a naming instance ('Look, a Zem'; +/- 2 seconds around each utterance) as a function of infants' posture.}\label{fig:detByNaming}
\end{figure}
\end{CodeChunk}

\subsection{Access to Faces and Wrists During Labeling
Events}\label{access-to-faces-and-wrists-during-labeling-events}

Finally, we analyzed how face and wrists detections changed during
object labeling events as a function of infant's posture and
orientation. Specfically, we analyzed a four-second window around each
labeling event (e.g., ``Look at the {[}zem{]}!''); these labeling events
were hand-annotated and automatically synchronized with the
frame-by-frame face detections. We again found that infant's posture and
orientations impacted the degree to which they saw their caregiver's
face and wrist during a labeling event; infants who were sitting or
standing were more likely to have access to this social information.

\section{General Discussion}\label{general-discussion}

We use a head-mounted camera to explore how children's postural and
locomotive development directly impacts their access to social
information, here operationalized as the presence of the faces and
wrists of their caregiver. We found that children's posture and
orientation towards their caregiver changed systematically across age,
and that both of these factors dramatically impacted the proportion of
faces and wrists that were avaliable in the child's visual field. Thus,
infants' postural and locomotive developments are mediating factors that
explain some of the age-related changes in the proportion of faces and
hands avaliable in infants' visual field. Broadly, this work suggests
that motoric developments mediate how infants experience their visual
world and the social information in it: infants that are sitting and
standing have a different view of their world, the people in it, and the
actions that are being performed.

This work also deploys novel advancements in computer vision to the
study of developmental psychology. The field of object detection and
recognition has advanced dramatically in the past five years since the
re-birth of deep learning algorithms (Krizhevsky, Sutskever, \& Hinton,
2012), creating a new generation of algorithmic tools. These tools are
substantially better equipped to deal with noisier, more complicated
datasets and can extract richer and more detailed information. Videos
from the infant perspective provided substantial challenges (e.g.,
partially occluded faces) for the classic models of face detection
(e.g., ViolaJones) (Viola \& Jones, 2004). Further, as the headcam
technologies employed here were inexpensive (\textasciitilde{}\$60 a
camera) and the computer vision algorithms freely avaliable, this method
is a promising avenue for quanitfying the visual and social information
avaliable to infant learners.

Thus, we suggest that the combined use of these new tools can be
leveraged to understand the changing infant perspecive on the visual
world and the implications of these changes for both linguistic,
cognitive, and social development.

\section{Acknowledgements}\label{acknowledgements}

Thanks to Kaia Simmons, Kathy Woo, Aditi Maliwal, and other members of
the Language and Cognition Lab for help in recruitment, data collection,
and annotation. This research was supported by a John Merck Scholars
grant to MCF. An earlier version of this work was presented to the
Cognitive Science Society in Frank et al. (2013). Please address
correspondence to Michael C. Frank, Department of Psychology, Stanford
University, 450 Serra Mall (Jordan Hall), Stanford, CA, 94305, tel:
(650) 724-4003, email: \texttt{mcfrank@stanford.edu}.

\section{References}\label{references}

\setlength{\parindent}{-0.1in} \setlength{\leftskip}{0.125in} \noindent

\hypertarget{refs}{}
\hypertarget{ref-adolph2007}{}
Adolph, K., \& Berger, S. (2007). Motor development. In \emph{Handbook
of child psychology}. Wiley Online Library.

\hypertarget{ref-adolph2012}{}
Adolph, K., Gilmore, R., Freeman, C., Sanderson, P., \& Millman, D.
(2012). Toward open behavioral science. \emph{Psychological Inquiry},
\emph{23}(3), 244--247.

\hypertarget{ref-bambach2017}{}
Bambach, S., Crandall, D. J., Smith, L. B., \& Yu, C. (2017). An
egocentric perspective on active vision and visual object learning in
toddlers. In \emph{Proceedings of the seventh joint ieee conference on
development and learning and on epigenetic robotics}.

\hypertarget{ref-cao2017realtime}{}
Cao, Z., Simon, T., Wei, S.-E., \& Sheikh, Y. (2017). Realtime
multi-person 2D pose estimation using part affinity fields. In
\emph{CVPR}.

\hypertarget{ref-clerkin2017}{}
Clerkin, E. M., Hart, E., Rehg, J. M., Yu, C., \& Smith, L. B. (2017).
Real-world visual statistics and infants' first-learned object names.
\emph{Phil. Trans. R. Soc. B}, \emph{372}(1711), 20160055.

\hypertarget{ref-csibra2009natural}{}
Csibra, G., \& Gergely, G. (2009). Natural pedagogy. \emph{Trends in
Cognitive Sciences}, \emph{13}(4), 148--153.

\hypertarget{ref-cummings1988}{}
Cummings, M., Van Hof-Van Duin, J., Mayer, D., Hansen, R., \& Fulton, A.
(1988). Visual fields of young children. \emph{Behavioural and Brain
Research}, \emph{29}(1), 7--16.

\hypertarget{ref-farroni2002eye}{}
Farroni, T., Csibra, G., Simion, F., \& Johnson, M. H. (2002). Eye
contact detection in humans from birth. \emph{Proceedings of the
National Academy of Sciences}, \emph{99}(14), 9602--9605.

\hypertarget{ref-fausey2016}{}
Fausey, C. M., Jayaraman, S., \& Smith, L. B. (2016). From faces to
hands: Changing visual input in the first two years. \emph{Cognition},
\emph{152}, 101--107.

\hypertarget{ref-franchak2011}{}
Franchak, J., Kretch, K., Soska, K., \& Adolph, K. (2011). Head-mounted
eye tracking: A new method to describe infant looking. \emph{Child
Development}.

\hypertarget{ref-frank2014visual}{}
Frank, M. C., Amso, D., \& Johnson, S. P. (2014). Visual search and
attention to faces during early infancy. \emph{Journal of Experimental
Child Psychology}, \emph{118}, 13--26.

\hypertarget{ref-frank2013}{}
Frank, M. C., Simmons, K., Yurovsky, D., \& Pusiol, G. (2013).
Developmental and postural changes in childrens visual access to faces.
In \emph{Proceedings of the 35th annual meeting of the cognitive science
society} (pp. 454--459).

\hypertarget{ref-frank2012measuring}{}
Frank, M. C., Vul, E., \& Saxe, R. (2012). Measuring the development of
social attention using free-viewing. \emph{Infancy}, \emph{17}(4),
355--375.

\hypertarget{ref-gredeback2010development}{}
Gredebäck, G., Fikke, L., \& Melinder, A. (2010). The development of
joint visual attention: A longitudinal study of gaze following during
interactions with mothers and strangers. \emph{Developmental Science},
\emph{13}(6), 839--848.

\hypertarget{ref-gredeback2008microstructure}{}
Gredebäck, G., Theuring, C., Hauf, P., \& Kenward, B. (2008). The
microstructure of infants' gaze as they view adult shifts in overt
attention. \emph{Infancy}, \emph{13}(5), 533--543.

\hypertarget{ref-iverson2010}{}
Iverson, J. M. (2010). Developing language in a developing body: The
relationship between motor development and language development.
\emph{Journal of Child Language}, \emph{37}(2), 229--261.

\hypertarget{ref-karasik2014}{}
Karasik, L. B., Tamis-LeMonda, C. S., \& Adolph, K. E. (2014). Crawling
and walking infants elicit different verbal responses from mothers.
\emph{Developmental Science}, \emph{17}(3), 388--395.

\hypertarget{ref-kretch2014}{}
Kretch, K. S., Franchak, J. M., \& Adolph, K. E. (2014). Crawling and
walking infants see the world differently. \emph{Child Development},
\emph{85}(4), 1503--1518.

\hypertarget{ref-krizhevsky2012imagenet}{}
Krizhevsky, A., Sutskever, I., \& Hinton, G. E. (2012). Imagenet
classification with deep convolutional neural networks. In
\emph{Advances in neural information processing systems} (pp.
1097--1105).

\hypertarget{ref-mayer1988}{}
Mayer, D., Fulton, A., \& Cummings, M. (1988). Visual fields of infants
assessed with a new perimetric technique. \emph{Investigative
Ophthalmology \& Visual Science}, \emph{29}(3), 452--459.

\hypertarget{ref-meltzoff2007like}{}
Meltzoff, A. N. (2007). ``Like me'': A foundation for social cognition.
\emph{Developmental Science}, \emph{10}(1), 126--134.

\hypertarget{ref-scheirer2014perceptual}{}
Scheirer, W. J., Anthony, S. E., Nakayama, K., \& Cox, D. D. (2014).
Perceptual annotation: Measuring human vision to improve computer
vision. \emph{IEEE Transactions on Pattern Analysis and Machine
Intelligence}, \emph{36}(8), 1679--1686.

\hypertarget{ref-simon2017hand}{}
Simon, T., Joo, H., Matthews, I., \& Sheikh, Y. (2017). Hand keypoint
detection in single images using multiview bootstrapping. In
\emph{CVPR}.

\hypertarget{ref-viola2004robust}{}
Viola, P., \& Jones, M. J. (2004). Robust real-time face detection.
\emph{International Journal of Computer Vision}, \emph{57}(2), 137--154.

\hypertarget{ref-walle2014}{}
Walle, E. A., \& Campos, J. J. (2014). Infant language development is
related to the acquisition of walking. \emph{Developmental Psychology},
\emph{50}(2), 336.

\hypertarget{ref-wei2016cpm}{}
Wei, S.-E., Ramakrishna, V., Kanade, T., \& Sheikh, Y. (2016).
Convolutional pose machines. In \emph{CVPR}.

\hypertarget{ref-yoshida2008}{}
Yoshida, H., \& Smith, L. (2008). What's in view for toddlers? Using a
head camera to study visual experience. \emph{Infancy}, \emph{13},
229--248.

\hypertarget{ref-yurovsky2012}{}
Yurovsky, D., Smith, L., \& Yu, C. (in press). Statistical word learning
at scale: The baby's view is better. \emph{Developmental Science}.

\hypertarget{ref-zhang2016}{}
Zhang, K., Zhang, Z., Li, Z., \& Qiao, Y. (2016). Joint face detection
and alignment using multitask cascaded convolutional networks.
\emph{IEEE Signal Processing Letters}, \emph{23}(10), 1499--1503.

\end{document}