### 2018-03-13 analysis-validation/fps_sanity_checks.Rmd
expand sanity check frame scripts

---
title: "FPS Sanity Checks"
output: html_notebook
---

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
```

```{r, message=FALSE, warning=FALSE}
new <- read_csv("../data/video_stats/new_movie_lengths.csv")
old <- read_csv("../data/video_stats/old_movie_lengths.csv")
```

```{r}
new %>%
  rename(new_fps = fps, new_length = length) %>%
  inner_join(old %>%
              mutate(old_fps = last.frame / length) %>%
              rename(old_length = length), 
            by = "subid") %>%
  mutate(length_close_enough = abs(new_length - old_length) <= 1) %>%
  mutate(fps_close_enough =  abs(new_fps - old_fps) <= 1) %>%
  kable
```
Notes:

XS_1229: Was trimmed from when headcam fell out and therefore is a different lengthj
XS_1628: Was exported with wrong frame count in Oct 2017; error was not caught until March 2018 (error was in CogSCi 2018
initial submission, corrected for revisions)