### 2018-05-11 analysis-validation/z-old/redo_mtcnn_csv.Rmd
reorg validation folder

---
title: "redo_mtcnn_csv"
output: html_document
---

```{r}
data <- read_csv("data/final_output/mtcnn2.csv") %>%
  select(-X1)

new_frames <- read_csv("data/final_output/new_frames.csv")
```

Filter out messed up videos (contain less frames)
```{r}
data %<>% 
  filter(!(video %in% c("XS_1612", "XS_1614", "XS_1631", "XS_1635")))
```

```{r}
result <- data %>%
  rbind(new_frames) %>%
  arrange(video, frame)
```

Sanity check # of frames (should be 970695)
```{r}
result %>% 
  distinct(video, frame, .keep_all = TRUE) %>%
  nrow
```

```{r}
write_csv(result, "data/final_output/mtcnn3.csv")
```