### 2018-11-06 analysis-validation/z-old/check_frame_counts_post_detection.Rmd
Get rid of duplicate sanity check files / old unused scripts

---
title: "Check frame counts after MTCNN detection"
output: html_notebook
---

```{r}
library(tidyverse)
library(knitr)
```

Load data
```{r}
showframes_frame_counts <- read_csv("../data/video_stats/frametimechecks_showframes.csv")

mtcnn_data <- read_csv("../data/final_detector_output/mtcnn3.csv")
```

Check mtcnn results frame counts against extracted frames
```{r}

showframes_frame_counts <- showframes_frame_counts %>%
  group_by(subid) %>%
  mutate(video = strsplit(subid,'.csv',2)) %>%
  mutate(video = str_replace(video,'-','_'))

check_frame_times <- mtcnn_data %>%
  distinct(video, frame) %>%
  group_by(video) %>%
  summarise(num_frames = n()) %>%
  left_join(showframes_frame_counts) %>%
  mutate(checkOK = num_frames == last.frame)

kable(check_frame_times)
```