### 2020-04-29 writeup-paper/draft_apa_march2020.tex

\documentclass[english,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Postural developments modulate children's visual access to social information},
            pdfkeywords={Postural developments modulate children's visual access to social
information},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Posture modulates social access}
\author{Bria L. Long\textsuperscript{1}, Alessandro Sanchez\textsuperscript{1}, Allison M. Kraus\textsuperscript{1}, Ketan Agrawal\textsuperscript{1}, \& Michael C. Frank\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\authornote{

Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
\keywords{Postural developments modulate children's visual access to social information\newline\indent Word count: X}
\usepackage{csquotes}

\title{Postural developments modulate children's visual access to social
information}

\date{}

\abstract{
The ability to process social information is a critical component of
children's early linguistic and cognitive development. However, as
children reach their first birthday, they begin to locomote themselves,
dramatically changing the way they see the world around them. How do
these postural and locomotor developments affect children's access to
the social information relevant for word-learning? We explored this
question by using head-mounted cameras to record the egocentric visual
perspective of 36 infants' (at 8, 12, and 16 months of age) during a
naturalistic play session with their caregiver. To estimate the
proportion of faces and hands in the infant view, we used a computer
vision algorithm to detect the presence of faces and hands in the entire
dataset. We found that infants' posture and orientation to their
caregiver changed dramatically across this age range and modulated their
access to this social information; infants who were sitting or standing
with their caregiver at a close distance tended to have the most
faces/hands in their visual field. We also applied our automated
analysis to the egocentric video data from a recent paper documenting
how posture changes one-year-olds visual access to faces (Franchak et
al., 2017), finding convergence across both methodologies and dataset
and confirming previous work that motoric developments play a
significant role in the emergence of children's linguistic and social
capacities. We suggest that the combined use of head-mounted cameras and
the application of new computer vision techniques is a promising avenue
for understanding the statistics of infants' visual and linguistic
experience as they change over development.
}

\begin{document}
\maketitle

From their earliest months, infants are deeply engaged in learning from
others. Even newborns tend to prefer to look at faces with direct
vs.~averted gaze (Farroni, Csibra, Simion, \& Johnson, 2002) and young
infants follow overt gaze shifts. And as infants learn to parse and
segment their native language(s), the social cues provided by speakers
(e.g., eye-gaze) may provide strong scaffolding for early word learning.
Indeed, empirical work suggests that children's ability to process
social cues is a key factor in early language development: in a
longitudinal study, children's level of joint engagement with their
mother at 9-12 months was found to predict both their receptive and
productive vocabularies (Carpenter, Nagell, \& Tomasello, 1998). More
specifically, 10 month-olds who follow an adult's gaze in an
experimental context have larger vocabularies at 18 months and
throughout the second year of life (Brooks \& Meltzoff, 2005, 2008).

Around their first birthday, however, children's ability to interact
with their world radically changes (Adolph \& Berger, 2006) as they are
no longer constrained to the same spot that their caregivers last placed
them in. Before the first birthday, children often begin crawling; soon
after, they begin to walk independently. While not all children crawl
(Adolph, Vereijken, \& Denny, 1998)---or crawl in the same
way---children tend to find creative ways of moving (e.g., scooting) or
getting others to help them move (e.g cruising) to begin to explore
their world on their own (Patrick, Noah, \& Yang, 2012). As independent
agents in the world, children are more active in the construction of
their own learning environments--yet spend much of their time in a world
primarily populated by knees

One possibility is that these motor improvements have strong
developmental cascades, impacting children's emerging social, cognitive,
and linguistic abilities (Iverson, 2010). Indeed, these postural changes
change how children interact with their mothers; walking infants make
different kinds of object-related bids for attention from their mothers
than crawling infants, and tend to hear more action directed statements
(e.g., \enquote{open it}) (Karasik, Tamis-LeMonda, \& Adolph, 2014). In
an observational study, Walle and Campos (2014) also found that children
who were able to walk had both higher receptive and productive
vocabularies. On their account, children's ability to stand and
independently locomote may fundamentally change their ability to access
the social information (e.g., facial expressions, gaze cues, pointing)
relative to children who are still crawling and sitting. In other words,
the ability of walking infants to access more detailed social
information may allow infants to learn words quicker and more
efficiently, facilitating language growth.

Over the past decade, researchers have started to use egocentric,
head-mounted cameras to document the social information that infants and
children have access to across early development, and to understand the
degree to which there are substantial shifts in their viewpoints that
may have downstream developmental consequences (Yoshida \& Smith, 2008).
Indeed, the infant visual perspective has proved to be both remarkably
different than the adult perspective and not easily predicted by our own
intuitions (Franchak, Kretch, Soska, \& Adolph, 2011, Yoshida and Smith
(2008), Clerkin, Hart, Rehg, Yu, and Smith (2017)). Over the first two
years of life, the infant viewpoint seems to transition from primarily
containing close up view of faces to capturing relatively restricted
views of hands paired with the objects they are acting on (Fausey,
Jayaraman, \& Smith, 2016); both computational and empirical work
suggest that this restricted viewpoint may be more effective for
learning about objects and their labels than the comparable adult
perspective (D. Yurovsky, Smith, \& Yu, 2012, Bambach, Crandall, Smith,
and Yu (2017)).

We hypothesized that children's postural developments may be partially
responsible for some of these changes in the infant perspective. During
spontaneous play, toddlers are more likely to look at the floor while
crawling than while walking (Franchak et al., 2011), when they have full
visual access to their environment and the people in it (Kretch,
Franchak, and Adolph, 2014). More recently, when 12-month-olds
(Franchak, Kretch, \& Adolph, 2017) participated in a free-play session
with their caregivers, their own posture and as well as their
caregiver's posture influenced the proportion of time they spent looking
at the faces and bodies of their caregivers (Franchak et al., 2017).

Here, we directly examine the role that postural developments that
infants' experience as they reach their first birthday change the social
information in the infant view. We expand on previous findings in three
key ways. First, using head-mounted cameras during a brief laboratory
free-play session, we record the visual experience of three groups of
children at 8, 12, and 16 months-of-age, covering a broad range of ages
and locomotor abilities around the first birthday; Children's posture
and orientation relative to their caregiver were also recorded from a
third-person perspective and hand-annotated. This cross-sectional design
thus allows us to directly examine the relative contributions of age
vs.~postural developments on children's visual access to social
information. Second, we use a modern computer vision algorithm for the
automated detection of faces and hands in the egocentric viewpoints of
infants. In particular, we capitalize on recent improvements in face and
pose detection algorithms (Cao, Simon, Wei, \& Sheikh, 2017; K. Zhang,
Zhang, Li, \& Qiao, 2016) to analyze the frequencies of faces and hands
in the child's visual environment, both overall and relative to naming
events by their caregivers. Use of this emerging technology allows us to
annotate the entirety of the dataset, allowing a more complete picture
of the changing infant perspective. We apply this same automated method
to the dataset used in Franchak et al. (2017), validating the
generalizability of this technique. Third, as parents were presented
with pairs of novel and familiar objects with their children during the
play sessions, we were able to explore the availability of social
signals during naming events.

Thus, the current study allows us to analyze changes in the visual
access to faces and hands according to children's age, posture, and
linguistic input. Broadly, we predicted that there would be differential
access to social information based on children's postural developments:
crawling infants would see fewer faces/hands because they would
primarily be looking at the ground, while walking toddlers would have
access to a richer visual landscape with greater access to the social
information in their environment.

\section{Methods}\label{methods}

\subsubsection{Participants}\label{participants}

Our final sample consisted of 36 infants and children, with 12
participants in three age groups: 8 months (6 F), 12 months (7 F), and
16 months (6 F). Participants were recruited from the surrounding
community via state birth records, had no documented disabilities, and
were reported to hear at least 80 percent English at home. Demographics
and exclusion rates are given in the table below.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
\hline
Group & N & \% incl. & Avg age & Avg video length (min) \\
\hline
8 mo. &   12 & 0.46 & 8.71 & 14.41 \\
12 mo. &  12 & 0.40 & 12.62 & 12.71 \\
16 mo. &  12 & 0.31 & 16.29 & 15.10\\
\hline
\end{tabular}
\end{table}

To obtain this final sample, we tested 95 children, excluding 59
children for the following reasons: 20 for technical issues related to
the headcam, 15 for failing to wear the headcam, 10 for fewer than 4
minutes of headcam footage, 5 for having multiple adults present, 5 for
missing Communicative Development Inventory (CDI) data, 2 for missing
scene camera footage, 1 for fussiness, and one for sample symmetry. All
inclusion decisions were made independent of the results of subsequent
analyses. These data were also analyzed in Frank, Simmons, Yurovsky, and
Pusiol (2013) and Sanchez, Long, Kraus, and Frank (2018).

\subsubsection{Procedure}\label{procedure}

All parents signed consent documents while children were fitted with the
headcam. If the child was uninterested in wearing the headcam or tried
to take it off, the experimenter presented engaging toys to try to draw
the child's focus away from the headcam. When the child was comfortable
wearing the headcam, the child and caregiver were shown to a playroom
for the free-play session. Parents were shown a box containing three
pairs of novel and familiar objects (e.g., a ball and a microfiber
duster, named a \enquote{zem}), and were instructed to play with the
object pairs with their child one at a time, \enquote{as they typically
would.} All parents confirmed that their child had not previously seen
the novel toys and were instructed to use the novel labels to refer to
the toys. The experimenter then left the playroom for approximately 15
minutes, during which a tripod-mounted camera in the corner of the room
recorded the session and the headcam captured video from the child's
perspective.

\subsubsection{Data Processing and
Annotation}\label{data-processing-and-annotation}

Headcam videos were trimmed such that they excluded the instruction
phase when the experimenter was in the room and were automatically
synchronized with the tripod-mounted videos using FinalCut Pro Software.
These sessions yielded 507 minutes (almost a million frames) of video,
with an average video length of 14.07 minutes (min = 4.53, max = 19.35).

\subsubsection{Posture and Orientation
Annotation}\label{posture-and-orientation-annotation}

We created custom annotations to describe the child's physical posture
(e.g.~standing) and the orientation of the child relative to the
caregiver (e.g.~far away). The child's posture was categorized as being
held/carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations.

\subsection{Face and Hand Detection}\label{face-and-hand-detection}

We used three face detection systems to measure infants' access to
faces. The first of these is the most commonly-used and widely available
face detection algorithm: Viola-Jones. We used this algorithm as a
benchmark for performance, as while it can achieve impressive accuracy
in some situations, it is notoriously bad at dealing with occluded faces
(Scheirer, Anthony, Nakayama, \& Cox, 2014). We next tested the
performance of two face detectors that both made use of recently
developed Convolutional Neural Networks (CNNs) to extract face
information. The first algorithm was specifically optimized for face
detection, and the second algorithm was optimized to extract information
about the position of 18 different body parts. For the second algorithm
(OpenPose; Cao et al., 2017), we used the agent's nose (one of the body
parts detected) to operationalize the presence of faces, as any half of
a face necessarily contains a nose.

The OpenPose detector also provided us with the location of an agent's
wrists, which we used as a proxy for hands for two reasons. First, as we
did not capture children's entire visual field, the presence of a wrist
is likely often indicative of the presence of a hand within the field of
view. Second, hands are often occluded by objects when caregivers are
interacting with children, yet still visually accessible by the child
and part of their joint interaction.

\begin{figure*}
\centering
\includegraphics[width=5.5in]{images/detector_samples_banner.pdf}
\caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.}
\end{figure*}

\subsubsection{Algorithms}\label{algorithms}

The first face detection system made use of a series of Haar
feature-based cascade classifiers (Viola \& Jones, 2004) applied to each
individual frame. The second algorithm (based on work by K. Zhang et al.
(2016)) uses multi-task cascaded convolutional neural networks (MTCNNs)
for joint face detection and alignment, built to perform well in
real-world environments where varying illuminations and occlusions are
present. We used a Tensorflow implementation of this algorithm available
at \url{https://github.com/davidsandberg/facenet}.

The CNN-based pose detector (OpenPose; Cao et al., 2017; Simon, Joo,
Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016)
provided the locations of 18 body parts (ears, nose, wrists, etc.) and
is available at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a CNN for initial anatomical detection and subsequently
applies part affinity fields (PAFs) for part association, producing a
series of body part candidates. The candidates are then matched to a
single individual and finally assembled into a pose; here, we only made
use of the body parts relevant to the face and hands (nose and wrists).

\subsubsection{Detector evaluation}\label{detector-evaluation}

To evaluate face detector performance, we hand-labeled a \enquote{gold
set} of labeled frames. To account for the relatively rare appearance of
faces in the dataset, we hand-labeled two types of samples: a sample
containing a high density of faces (half reported by MTCNN, half by
OpenPose) and a random sample from the remaining frames. Each sample was
comprised of an equal number of frames taken from each child's video.
For wrist detections, the \enquote{gold set} was constructed in the same
manner, except frames with a high density of wrists came only from
detections made by OpenPose. Faces were classified as present if at
least half of the face was showing; wrists were classified as present if
any part of the wrist was showing. Two authors labelled the frames
independently and resolved disagreements on a case-by-case basis.
Precision (hits / hits + false alarms), recall (hits / hits + misses),
and F-score (harmonic mean of precision and recall) were calculated for
all detectors and are reported in Table 1.

For face detection, MTCNN outperformed OpenPose when taking into account
only the composite F-score (0.89 MTCNN vs.~0.83 OpenPose). MTCNN and
OpenPose performed comparably with the random sample, with MTCNN having
higher precision on the high density sample, suggesting that OpenPose
generated slightly more false positives than MTCNN. ViolaJones performed
quite poorly relative to the other detectors, especially with respect to
the random sample. For wrist detection, OpenPose performed moderately
well (F = 0.74) with relatively high precision but low recall on the
randomly sampled frames (see Table 1). We thus analyze wrist detections,
with the caveat that we are likely underestimating the proportion of
hands in the dataset.

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
\hline
Algorithm & Sample\ Type & P & R & F \\ 
\hline
MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
OpenPose-Wrists & High density & 0.66 & 1.00 & 0.79 \\ 
OpenPose-Wrists & Random & 0.88 & 0.29 & 0.44 \\ 
\hline
\end{tabular}
\caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected). P, R, and F denote precision, recall, and F-score, respectively.} 
\vspace{-1em}
\end{table}

\section{Results}\label{results}

First, we report developmental shifts in infants' posture and their
orientation relative to their caregiver, consistent with previous
literature (Adolph \& Berger, 2006; Franchak et al., 2017). Then, we
examine how these changes influence children's visual access to faces
and wrists/hands across this developmental time range, examining the
relative contributions of age vs.~postural developments. We also explore
how these changes impact the accessibility of faces and wrists/hands
during labeling events, both when parents were labeling familiar objects
(e.g., \enquote{cat}) as well as novel objects (e.g., \enquote{zem}).
Finally, we apply the same automated detection method to a egocentric
video dataset collected by a different lab (Franchak et al., 2017)
during which children's in-the-moment posture was annotated.

\begin{figure*}[!h]

{\centering \includegraphics[width=\textwidth]{figs/posture-1}

}

\caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:posture}
\end{figure*}

\begin{figure*}[!h]

{\centering \includegraphics[width=\textwidth]{figs/indivPosOrient-1}

}

\caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:indivPosOrient}
\end{figure*}

\subsection{Changes in Posture and
Orientation}\label{changes-in-posture-and-orientation}

How do childen's in-the-moment posture and orientation change across
age? In our dataset, the proportion of time infants spent sitting
decreased with age, and the proportion of time infants spent standing
increased with infants' age. As children got older, their locomotive
abilities allowed them to become more independent. Both 8-month-olds and
12-month-olds spent relatively equivalent amounts of time lying/crawling
(i.e., \enquote{prone}) which was markedly decreased in the
16-month-olds, who spent most of their time sitting or standing (see
Figure \ref{fig:posture}). We also observed changes in children's
orientation relative to their caregivers: the 8-month-olds spent more
time with their caregiver behind them supporting their sitting positions
than did children at other ages (see Figure \ref{fig:posture}). However,
we also saw considerable variability across children: some infants spent
almost their entire time sitting at a close distance from their
caregiver, whereas others showed more considerable variability (see
Figure \ref{fig:indivPosOrient}).

\subsection{Changes in Access to Faces and
Hands}\label{changes-in-access-to-faces-and-hands}

We first examined the proportion of face and hand detections as a
function of children's age without considering their posture (see Figure
\ref{fig:detByAge}); here, we report face and wrist detections using
OpenPose, although the same pattern of results for face detections was
found using the outputs from MTCNN. While faces tended to be in the
field-of-view overall more often than hands, children's head-mounted
cameras were angled slightly upward to capture the presence of faces,
and hand detections suffered from somewhat lower recall than face
detections. Going forward, we thus analyze differences in the relative
proportion of faces or hands in view as a function of age, posture, and
orientation, rather than comparing them directly. Overall, we observed
that 12-month-olds appeared to have visual access to slightly fewer
faces than 8 or 16-month-olds, creating a slight U-shaped function in
face detections; conversely, hand detections were showed a slight
increase across this age range, as reported in prior literature (Fausey
et al., 2016).

However, these age-related trends were much smaller than the effect of
infant's postural developments on children's visual access to faces and
hands. Children's in-the-moment posture was a major factor both in how
many faces and hands were in view during the play session, as was their
orientation relative to their caregiver. Infants who were sitting or
standing had more faces in view than infants who were lying
down/crawling (i.e.~prone), which was most frequent among 12-month-olds
relative to the other age groups (Figure \ref{fig:detByPosOrient}). When
caregivers were behind their children, supporting their children's
sitting or standing positions, children saw fewer faces and wrists. In
particular, children who were sitting or in front of their caregiver had
a high proportion of faces and hands in their field of view (Figure
\ref{fig:detByPosOrient}).

These trends were quantified using two generalized linear mixed-effect
models estimating the proportion of faces and hands that were in view,
with orientation, posture, their interaction, and scaled participant's
age as fixed effects, and with random slopes for infants' orientation
and posture. A summary of the coefficients of the models can be found in
Table 2, confirming that infants who were sitting/standing and in front
of their caregivers saw the most faces and hands. When modeling the
proportion of faces seen, age was not a significant predictor; however,
age remained a significant predictor when modeling the proportion of
hands seen by infants. Overall, these results suggest that infants'
visual access to social information is largely modulated by their
posture and orientation to their caregiver, which is in turn a function
of their general locomotor development. Nonetheless, these results also
suggest that infants may still increase their attention to hands
throughout this age range as they continue to learn about objects, their
functions, and their names.

\subsection{Access to Faces and Hands During Labeling
Events}\label{access-to-faces-and-hands-during-labeling-events}

Our play session was designed to provide parents with opporunities to
label objects--both familiar and novel--such that we could examine
whether children sought out different kinds of social information around
naming events. We thus explored how face and hand detections changed
during object labeling events, analyzing a four-second window around
each labeling event (e.g., \enquote{Look at the {[}zem{]}!}). Every
utterance of one of the labelled objects (e.g., \enquote{ball}) was
counted as a \enquote{labeling event}; timestamps of the beginning of
each word were hand-annotated from transcripts ad synchronized with the
frame-by-frame detections. Overall, we did not find major differences in
face/hand detections during naming events relative to baseline, either
for novel objects or for familiar objects (see Figure
\ref{fig:detByNaming}). This was true both when we analyzed naming rates
with or without taking into account children's posture and orientation
relative to their caregiver. These results suggest that either children
did not actively seek out social information during this particular play
session around naming events, perhaps because were a limited number of
possible referents at any given time.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\
  \hline
Intercept & -3.35 & 0.17 & -19.49 & 0.000 \\
  Sit & 0.31 & 0.20 & 1.58 & 0.114 \\
  Stand & -0.02 & 0.20 & -0.12 & 0.905 \\
  Close & 0.03 & 0.19 & 0.15 & 0.882 \\
  Far & 0.46 & 0.24 & 1.93 & 0.053 \\
  Age (Scaled) & 0.14 & 0.11 & 1.22 & 0.223 \\
  Sit*Close & 0.92 & 0.07 & 13.31 & 0.000 \\
  Stand*Close & 1.30 & 0.08 & 16.02 & 0.000 \\
  Sit*Far & 0.54 & 0.07 & 7.37 & 0.000 \\
  Stand*Far & 1.24 & 0.09 & 14.18 & 0.000 \\
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants (as detected by OpenPose).}
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\
  \hline
Intercept & -4.28 & 0.23 & -18.67 & 0.000 \\
  Sit & 0.54 & 0.19 & 2.90 & 0.004 \\
  Stand & 0.61 & 0.23 & 2.63 & 0.009 \\
  Close & 0.67 & 0.22 & 3.05 & 0.002 \\
  Far & 0.33 & 0.24 & 1.35 & 0.176 \\
  Age (Scaled) & 0.45 & 0.13 & 3.35 & 0.001 \\
  Sit*Close & 0.34 & 0.09 & 3.93 & 0.000 \\
  Stand*Close & 0.73 & 0.09 & 7.92 & 0.000 \\
  Sit*Far & -0.15 & 0.11 & -1.39 & 0.165 \\
  Stand*Far & 1.32 & 0.11 & 11.70 & 0.000 \\
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants (as detected by OpenPose).}
\end{table}

\begin{figure*}[!h]

{\centering \includegraphics[width=\textwidth]{figs/detByAge-1}

}

\caption{Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data.}\label{fig:detByAge}
\end{figure*}

\begin{figure*}[!h]

{\centering \includegraphics[width=\textwidth]{figs/detByPosOrient-1}

}

\caption{Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\% bootstrapped confidence intervals}\label{fig:detByPosOrient}
\end{figure*}

\begin{figure*}[!h]

{\centering \includegraphics[width=\textwidth]{figs/detByNaming-1}

}

\caption{Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95\% bootstrapped confidence intervals.}\label{fig:detByNaming}
\end{figure*}

\subsection{Extension to Franchak et al.,
2017}\label{extension-to-franchak-et-al.-2017}

In the present work, we found that infant's in-the-moment posture
changed with their age, as did infants' orientation relative their
caregiver. In a related study with 12-month-olds, Franchak et al. (2017)
also that children's in-the-moment posture changed the amount of time
infants spent looking at faces. Here, we sought to replicate the
findings from Franchak et al. (2017) using our automated methodology
(OpenPose detections), using the footage from their head-mounted cameras
(D. A. Simon, Gordon, Steiger, \& Gilmore, 2015).

This dataset differed from the present in two key ways that could
present a challenge for our automated methodology. First, the
environment that infants were immersed in with their caregivers (and
experimenters) was much larger and more varied than the play room used
in the present dataset, containing multiple structures and toys in
different parts of the room for infants to explore, unlike the room used
for the present dataset which was relatively small (approximately 10' x
10'). Thus, our automated methodology could fail to generalize to scenes
from these more complex environments, where detecting faces and hands
could arguably be a much harder task. Second, Franchak et al. (2017)
used a head-mounted eye-tracker, finding that children's posture
affected where children were looking within their visual field.
Nonetheless, if children are often orienting their heads towards where
they are allocating their attention (Yoshida \& Smith, 2008) then we
should expect to find the same pattern of results only when analyzing
the information in view.

\begin{figure*}[!h]

{\centering \includegraphics[width=\textwidth]{figs/franchak-1}

}

\caption{Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals.}\label{fig:franchak}
\end{figure*}

Overall, we found convergence between our two methodologies: we
replicated the main results from Franchak et al. (2017), finding that
the proportion of faces in view was greater when infants were sitting or
standing vs.~prone. We also found the same patterns of results with
hands (i.e., wrist detections), though these were not originally
annotated in Franchak et al. (2017) (see Figure \ref{fig:franchak});
results were validated with generalized mixed effect models as in our
previous analysis. Interestingly, there was a relatively higher
proportion of hands in view in Franchak et al. (2017) relative to the
present dataset; this is likely mostly due to the fact that there were
often multiple people in view: experimenters stayed in the room to film
children and assist the caregiver, while in our present dataset the
experimenters left the room for the majority of the play session.
Nonetheless, we still found that posture modulated the proportion of
hands in view, suggesting that this is a major factor that structures
infants' access to visual information.

\section{General Discussion}\label{general-discussion}

How do postural and locomotive developments influence the social
information that infants see? We used a head-mounted camera to explore
how these emerging motoric abilities affect children's visual access to
social information, here operationalized as the presence of the faces
and hands of their caregiver. First, we found systematic changes around
the first birthday in children's in-the-moment posture and their
orientation relative to the caregivers; older children spent more time
standing and less time sitting; older children's caregivers spent less
time supporting their standing or sitting postures. Children's changing
posture and orientation to their caregiver jointly shaped the amount of
social information that was in their view during one-on-one play
sessions with their caregivers: Children saw the most faces/hands when
they were sitting or standing and close to their caregiver. Motor
development appears to modulate how infants experience their visual
world and the social information in it.

However, we did not see differences in visual access to social
information during naming events, suggesting that children did not
change where they were looking when hearing a label for either a novel
or familiar object. We see several potential explanations for this:
first, other work (Yoshida \& Smith, 2008; Yu \& Smith, 2013, 2017),
including Franchak et al. (2017), has found that infants spend much more
time looking at the toys vs.~their caregiver's faces during these play
sessions; infants may have thus been primarily interested in exploring
these new toys rather than learning their names. Furthermore, moving
their own neck upwards towards their caregiver still requires quite a
physical effort at this age (especially when prone or sitting) and
simply may not have been a priority for children in this context. A
second factor is that this particular play session did not present many
opportunities where children would need to use social cues to
disambiguate referents. Indeed, there were only two possible referents
in the room at a time----and one of them was always a familiar category
(i.e., car, kitty).

Broadly,these results show promise for the use of automated methods to
detect the social information in the egocentric infant viewpoint during
more naturalistic parent-child interactons. The use of this automated
methodology allowed us to easily annotate the entirety of the dataset
and analyze all of the frames without the need for hundreds of hours of
human coding. Importantly, we also applied these same methods to an
addiontal dataset that contained more variation in the type and
structure of the play session, replicating and extending the effects of
children's in-the-moment posture on visual access to social information
(Franchak et al., 2017).

Yet more work is needed to understand how these results relate to
children's home experiences (Fausey et al., 2016). Both this play
session and that of Franchak et al. (2017) were relatively controlled
interactions where caregivers are highly aware that they are being
montiored by experimenters. Furthermore, as children grow and change,
the activities in which they engage with their caregivers are likely to
also vary, leading to differences in the distribution of contexts they
experience that may not be captured in these one-on-one play sessions.
Finally, the ability to walk is of course only of a cascade of changes
in children's abilities and experiences, and this study captures only a
cross-sectional slice of this broader, multifaceted trajectory.

Understanding these changes and how they relate to one another has been
a persistent challenge for developmental psychology, but the field of
computer vision has advanced dramatically in recent years, creating a
new generation of algorithmic tools. These tools deal better with
noisier, more complicated datasets and extract richer information than
previous system. We hope that these new tools can be leveraged to
understand the changing infant perspective on the visual world, both in
controlled experimental contexts and in children's home environments. By
doing so, we believe that we can understand the implications of the
changing infant perspective consequences for linguistic, cognitive, and
social development.

\section{Acknowledgements}\label{acknowledgements}

Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in
recruitment, data collection, and annotation. This work was funded by a
Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF,
and NSF \#1714726 to BLL. An earlier version of this work was presented
to the Cognitive Science Society in Frank et al. (2013) and Sanchez et
al. (2018).

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-adolph2006motor}{}
Adolph, K. E., \& Berger, S. E. (2006). Motor development.
\emph{Handbook of Child Psychology}.

\hypertarget{ref-adolph2012toward}{}
Adolph, K. E., Gilmore, R. O., Freeman, C., Sanderson, P., \& Millman,
D. (2012). Toward open behavioral science. \emph{Psychological Inquiry},
\emph{23}(3), 244--247.

\hypertarget{ref-adolph1998roles}{}
Adolph, K. E., Vereijken, B., \& Denny, M. (1998). Roles of variability
and experience in development of crawling. \emph{Child Development},
\emph{69}(1299), 312.

\hypertarget{ref-bambach2017}{}
Bambach, S., Crandall, D. J., Smith, L. B., \& Yu, C. (2017). An
egocentric perspective on active vision and visual object learning in
toddlers. In \emph{Proceedings of the seventh joint ieee conference on
development and learning and on epigenetic robotics}.

\hypertarget{ref-brooks2005}{}
Brooks, R., \& Meltzoff, A. (2005). The development of gaze following
and its relation to language. \emph{Developmental Science}, \emph{8}(6),
535--543.

\hypertarget{ref-brooks2008}{}
Brooks, R., \& Meltzoff, A. N. (2008). Infant gaze following and
pointing predict accelerated vocabulary growth through two years of age:
A longitudinal, growth curve modeling study. \emph{Journal of Child
Language}, \emph{35}(1), 207--220.

\hypertarget{ref-cao2017realtime}{}
Cao, Z., Simon, T., Wei, S.-E., \& Sheikh, Y. (2017). Realtime
multi-person 2D pose estimation using part affinity fields. In
\emph{CVPR}.

\hypertarget{ref-carpenter1998}{}
Carpenter, M., Nagell, K., \& Tomasello, M. (1998). Social cognition,
joint attention, and communicative competence from 9 to 15 months of
age. \emph{Monographs of the Society for Research in Child Development},
\emph{63}(4).

\hypertarget{ref-clerkin2017}{}
Clerkin, E. M., Hart, E., Rehg, J. M., Yu, C., \& Smith, L. B. (2017).
Real-world visual statistics and infants' first-learned object names.
\emph{Phil. Trans. R. Soc. B}, \emph{372}(1711), 20160055.

\hypertarget{ref-farroni2002eye}{}
Farroni, T., Csibra, G., Simion, F., \& Johnson, M. H. (2002). Eye
contact detection in humans from birth. \emph{Proceedings of the
National Academy of Sciences}, \emph{99}(14), 9602--9605.

\hypertarget{ref-fausey2016}{}
Fausey, C. M., Jayaraman, S., \& Smith, L. B. (2016). From faces to
hands: Changing visual input in the first two years. \emph{Cognition},
\emph{152}, 101--107.

\hypertarget{ref-franchak2017see}{}
Franchak, J. M., Kretch, K. S., \& Adolph, K. E. (2017). See and be
seen: Infant--caregiver social looking during locomotor free play.
\emph{Developmental Science}.

\hypertarget{ref-franchak2011}{}
Franchak, J. M., Kretch, K. S., Soska, K. C., \& Adolph, K. E. (2011).
Head-mounted eye tracking: A new method to describe infant looking.
\emph{Child Development}, \emph{82}(6), 1738--1750.

\hypertarget{ref-frank2013}{}
Frank, M. C., Simmons, K., Yurovsky, D., \& Pusiol, G. (2013).
Developmental and postural changes in children's visual access to faces.
In \emph{Proceedings of the 35th annual meeting of the cognitive science
society} (pp. 454--459).

\hypertarget{ref-iverson2010}{}
Iverson, J. M. (2010). Developing language in a developing body: The
relationship between motor development and language development.
\emph{Journal of Child Language}, \emph{37}(2), 229--261.

\hypertarget{ref-karasik2014}{}
Karasik, L. B., Tamis-LeMonda, C. S., \& Adolph, K. E. (2014). Crawling
and walking infants elicit different verbal responses from mothers.
\emph{Developmental Science}, \emph{17}(3), 388--395.

\hypertarget{ref-patrick2012developmental}{}
Patrick, S. K., Noah, J. A., \& Yang, J. F. (2012). Developmental
constraints of quadrupedal coordination across crawling styles in human
infants. \emph{Journal of Neurophysiology}, \emph{107}(11), 3050--3061.

\hypertarget{ref-sanchez2018postural}{}
Sanchez, A., Long, B., Kraus, A. M., \& Frank, M. C. (2018). Postural
developments modulate children's visual access to social information.

\hypertarget{ref-scheirer2014perceptual}{}
Scheirer, W. J., Anthony, S. E., Nakayama, K., \& Cox, D. D. (2014).
Perceptual annotation: Measuring human vision to improve computer
vision. \emph{IEEE Transactions on Pattern Analysis and Machine
Intelligence}, \emph{36}(8), 1679--1686.

\hypertarget{ref-simon2015databrary}{}
Simon, D. A., Gordon, A. S., Steiger, L., \& Gilmore, R. O. (2015).
Databrary: Enabling sharing and reuse of research video. In
\emph{Proceedings of the 15th acm/ieee-cs joint conference on digital
libraries} (pp. 279--280).

\hypertarget{ref-simon2017hand}{}
Simon, T., Joo, H., Matthews, I., \& Sheikh, Y. (2017). Hand keypoint
detection in single images using multiview bootstrapping. In
\emph{CVPR}.

\hypertarget{ref-viola2004robust}{}
Viola, P., \& Jones, M. J. (2004). Robust real-time face detection.
\emph{International Journal of Computer Vision}, \emph{57}(2), 137--154.

\hypertarget{ref-walle2014}{}
Walle, E. A., \& Campos, J. J. (2014). Infant language development is
related to the acquisition of walking. \emph{Developmental Psychology},
\emph{50}(2), 336.

\hypertarget{ref-wei2016cpm}{}
Wei, S.-E., Ramakrishna, V., Kanade, T., \& Sheikh, Y. (2016).
Convolutional pose machines. In \emph{CVPR}.

\hypertarget{ref-yoshida2008}{}
Yoshida, H., \& Smith, L. (2008). What's in view for toddlers? Using a
head camera to study visual experience. \emph{Infancy}, \emph{13},
229--248.

\hypertarget{ref-yu2013joint}{}
Yu, C., \& Smith, L. B. (2013). Joint attention without gaze following:
Human infants and their parents coordinate visual attention to objects
through eye-hand coordination. \emph{PloS One}, \emph{8}(11).

\hypertarget{ref-yu2017hand}{}
Yu, C., \& Smith, L. B. (2017). Hand--eye coordination predicts joint
attention. \emph{Child Development}, \emph{88}(6), 2060--2078.

\hypertarget{ref-yurovsky2012}{}
Yurovsky, D., Smith, L., \& Yu, C. (2012). Statistical word learning at
scale: The baby's view is better. \emph{Developmental Science}.

\hypertarget{ref-zhang2016}{}
Zhang, K., Zhang, Z., Li, Z., \& Qiao, Y. (2016). Joint face detection
and alignment using multitask cascaded convolutional networks.
\emph{IEEE Signal Processing Letters}, \emph{23}(10), 1499--1503.

\endgroup

\end{document}