### 2020-08-11 writeup-paper/render_examples.Rmd
scripts for naming/posture reliability and rendering examples

---
title: "render_examples"
author: "Bria Long"
date: "7/31/2020"
output: html_document
---

# Load libraries

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(lme4)
library(gridExtra)
library(assertthat)
library(tidyverse)
library(viridis)
library(papaja)
library(png)
library(magick)

# base_size in theme_few (changes font size in plots)
plot_size=16
```

####  Load detections
```{r loadDetections}

d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_july2020.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

## make long form version for plotting both face detections on same graph
dlong <- d %>%
  mutate(id = paste(subid,frame)) %>%
  gather(key = detectorType, value = detection, faceMT, faceOP, faceOPStrict, faceVJ, wristOP, wristOPStrict) %>%
  mutate(detectorType=as.factor(detectorType))  %>%
  mutate(detection = replace_na(detection,FALSE)) ## few frames where OP didn't have detections

# for better plotting names
levels(dlong$detectorType) 
#[1] "faceMT"        "faceOP"        "faceOPStrict"  "faceVJ"       
#[5] "wristOP"       "wristOPStrict"
levels(dlong$detectorType) <- c("MTCNN-Faces","OpenPose-Faces", "OpenPose-Faces-Strict","ViolaJones-Faces","OpenPose-Wrists","OpenPose-Wrists-Strict")

```

## Load ground truth
```{r groundTruthSetup, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN GROUND TRUTH and load function
ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv") %>%
  mutate(sample_type=as.factor(sample_type))

ground_truth_wrists <- read_csv("../data/ground_truth/ground_truth_wrists.csv") %>%
  mutate(sample_type=as.factor(sample_type))

# Renaming for table output
levels(ground_truth_faces$sample_type)=c("High density", "Random")
levels(ground_truth_wrists$sample_type)=c("High density", "Random")

# Function to evaluate detectors
evaluate <- function(a, b) {
  if (a == TRUE) {
    if (a == b) return ("TP") # was face/wrist, detected face/wrist
    else return("FN") # was face/wrist, missed face/wrist
  }
  else {
    if (a == b) return("TN") # was not face/wrist, did not detect face/wrist
    else return("FP") # was not face/wrist, detected face/wrist
  }
}
```

## Evaluate face detectors
```{r}
performanceFaces <- ground_truth_faces  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong)   %>%
  filter(!detectorType %in% c('OpenPose-Wrists-Strict','OpenPose-Wrists')) %>%
  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face)) # always have the gronud truth first in evaluate argument
```

## Set up paths for getting detection images
```{r}
example_op_images = paste0(here::here(),'/data/openpose_examples/3sample/')
goldset_image_dir = paste0(here::here(),'/data/ground_truth/goldset_frames/face_images/')
dir.create(paste0(here::here(),'/data','/example_montages/'))
save_dir = paste0(here::here(),'/data','/example_montages/')
no_consent_show_images = c('XS_1224','XS_1622','XS_1628','XS_1631','XS_1634')

performanceFacesToRender <- performanceFaces %>%
  filter(!subid %in% no_consent_show_images) %>% #don't render frames from non-consenting fams
  mutate(frame_string = sprintf("image-%05d", frame)) %>%
  # mutate(image_path = paste0(goldset_image_dir, video, '/',frame_string,'.jpg')) %>%
  mutate(image_path = paste0(example_op_images, video, '_',frame_string,'_rendered.png')) %>% 
  #NOTE: only renders high confidence OP detections
  mutate(file_exists = file.exists(image_path)) %>%
  filter(file_exists==TRUE)
```

## Render out example errors for Viola Jones
```{r}
false_negs_VJ <- performanceFacesToRender %>%
  filter(detectorType=='ViolaJones-Faces') %>%
  filter(result=="FN") %>%
  filter(sample_type=='Random')

false_pos_VJ <- performanceFacesToRender %>%
  filter(detectorType=='ViolaJones-Faces') %>%
  filter(result=="FP") %>%
  filter(sample_type=='Random')

## note these will take a random sample, so set seed if trying to reproduce everything
# set.seed(30)
OpenPose_TP_ViolaJones_FN <- performanceFacesToRender %>%
  filter(image_path %in% false_negs_VJ$image_path) %>%
  filter(detectorType=='OpenPose-Faces') %>%
  filter(result=='TP')  %>%
  sample_n(8)

OpenPose_TN_ViolaJones_FP <- performanceFacesToRender %>%
  filter(image_path %in% false_pos_VJ$image_path) %>%
  filter(detectorType=='OpenPose-Faces') %>%
  filter(result=='TN') %>%
  sample_n(8)
```

### Make montages 
```{r}
image_read(false_negs_VJ$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "false_negs_VJ.png")))

image_read(false_pos_VJ$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "false_pos_VJ.png")))

image_read(OpenPose_TP_ViolaJones_FN$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "OpenPose_TP_ViolaJones_FN.png")))

image_read(OpenPose_TN_ViolaJones_FP$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "OpenPose_TN_ViolaJones_FP.png")))

false_negs_MTCNN <- performanceFacesToRender %>%
  filter(detectorType=='MTCNN-Faces') %>%
  filter(result=="FN")

OpenPose_TP_MTCNN_FN <- performanceFacesToRender %>%
  filter(image_path %in% false_negs_MTCNN$image_path) %>%
  filter(detectorType=='OpenPose-Faces') %>%
  filter(result=='TP') %>%
  sample_n(16)

image_read(OpenPose_TP_MTCNN_FN$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "OpenPose_TP_MTCNN_FN.png")))
```

## Where OpenPose had false alarms that MTCNN correctly rejected 
```{r}
false_pos_OP <- performanceFacesToRender %>%
  filter(detectorType=='OpenPose-Faces') %>%
  filter(result=="FP")

OpenPose_FP_MTCNN_TN <- performanceFacesToRender %>%
  filter(image_path %in% false_pos_OP$image_path) %>%
  filter(detectorType=='MTCNN-Faces') %>%
  filter(result=='TN') %>%
  sample_n(16)

image_read(OpenPose_FP_MTCNN_TN$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(goldset_image_dir,"example_montages/", "OpenPose_FP_MTCNN_TN.png")))
```

## by posture
```{r}
sit <- performanceFacesToRender %>%
   filter(posture=='sit') %>%
   # filter(is_face==TRUE) %>%
   # filter(detectorType=='OpenPose-Faces-Strict') %>%
   # filter(detection==TRUE) %>%
   sample_n(8)

stand <- performanceFacesToRender %>%
   filter(posture=='stand') %>%
   # filter(is_face==TRUE) %>%
   # filter(detectorType=='OpenPose-Faces-Strict') %>%
   # filter(detection==TRUE) %>%
   sample_n(8)

prone <- performanceFacesToRender %>%
   filter(posture=='prone') %>%
   # filter(is_face==TRUE) %>%
   # filter(detectorType=='OpenPose-Faces-Strict') %>%
   # filter(detection==TRUE) %>%
   sample_n(8)

# carry <- performanceFacesToRender %>%
#    filter(posture=='carry') %>%
#    filter(is_face==TRUE) %>%
#    filter(detectorType=='OpenPose-Faces-Strict') %>%
#    filter(detection==TRUE) %>%
#    sample_n(8)

image_read(prone$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(save_dir, "prone.png")))

image_read(sit$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(save_dir, "sit.png")))

image_read(stand$image_path) %>%
    image_append(stack = FALSE) %>%
    image_write(file.path(paste0(save_dir, "stand.png")))

```