### 2020-08-11 writeup-paper/old-drafts/backup.Rmd
code for possible extra plots

### Procedure
All parents signed consent documents while children were fitted with the headcam. If the child was uninterested in wearing the headcam or tried to take it off, the experimenter presented engaging toys to try to draw the child's focus away from the headcam. When the child was comfortable wearing the headcam, the child and caregiver were shown to a playroom for the free-play session. Parents were shown a box containing three pairs of novel and familiar objects (e.g., a ball and a microfiber duster, named a "zem"), and were instructed to play with the object pairs with their child one at a time, "as they typically would." All parents confirmed that their child had not previously seen the novel toys and were instructed to use the novel labels to refer to the toys. The experimenter then left the playroom for approximately 15 minutes, during which a tripod-mounted camera in the corner of the room recorded the session and the headcam captured video from the child's perspective.

### Data Processing and Annotation
Headcam videos were trimmed such that they excluded the instruction phase when the experimenter was in the room and were automatically synchronized with the tripod-mounted videos using FinalCut Pro Software. These sessions yielded 
`r round(sum(inc_subs$len.min), digits = 0)` minutes (almost a million frames) of video, with an average video length of
`r round(mean(inc_subs$len.min), digits = 2)` minutes
(min = `r round(min(inc_subs$len.min), digits = 2)`, max = `r round(max(inc_subs$len.min), digits = 2)`).

### Posture and Orientation Annotation
We created custom annotations to describe the child's physical posture (e.g. standing) and the orientation of the child relative to the caregiver (e.g. far away). The child's posture was categorized as being held/carried, prone (crawling or lying), sitting, or standing. The caregiver's orientation was characterized as being close, far, or behind the child (independent of distance). For the first two annotations (close/far from the child), the caregiver could either be to the front or side of the child. All annotations were made by a trained coder using the OpenSHAPA/Datavyu software [@adolph2012toward]. Times when the child was out of view of the tripod camera were marked as uncodable and were excluded from these annotations.

## Face and Hand Detection
We used three face detection systems to measure infants' access to faces. The first of these is the most commonly-used and widely available face detection algorithm: Viola-Jones. We used this algorithm as a benchmark for performance, as while it can achieve impressive accuracy in some situations, it is notoriously bad at dealing with occluded faces [@scheirer2014perceptual]. We next tested the performance of two face detectors that both made use of recently developed Convolutional Neural Networks (CNNs) to extract face information. The first algorithm was specifically optimized for face detection, and the second algorithm was optimized to extract information about the position of 18 different body parts. For the second algorithm [OpenPose; @cao2017realtime], we used the agent's nose (one of the body parts detected) to operationalize the presence of faces, as any half of a face necessarily contains a nose.

The OpenPose detector also provided us with the location of an agent's wrists, which we used as a proxy for hands for two reasons. First, as we did not capture children's entire visual field, the presence of a wrist is likely often indicative of the presence of a hand within the field of view. Second, hands are often occluded by objects when caregivers are interacting with children, yet still visually accessible by the child and part of their joint interaction.

<!-- \begin{figure*} -->
<!-- \centering -->
<!-- \includegraphics[width=5.5in]{images/detector_samples_banner.pdf} -->
<!-- \caption{\label{fig:frames} Example face and pose detections made by OpenPose (top row) and MTCNN (bottom row) from a child in each age group. The last column features a false positive from OpenPose and a false negative from MTCNN.} -->
<!-- \end{figure*} -->

### Algorithms
The first face detection system made use of a series of Haar feature-based cascade classifiers [@viola2004robust] applied to each individual frame. The second algorithm (based on work by @zhang2016) uses multi-task cascaded convolutional neural networks (MTCNNs) for joint face detection and alignment, built to perform well in real-world environments where varying illuminations and occlusions are present. We used a Tensorflow implementation of this algorithm avaliable at https://github.com/davidsandberg/facenet.

The CNN-based pose detector  [OpenPose; @cao2017realtime; @simon2017hand; @wei2016cpm] provided the locations of 18 body parts (ears, nose, wrists, etc.) and is available at https://github.com/CMU-Perceptual-Computing-Lab/openpose. The system uses a CNN for initial anatomical detection and subsequently applies part affinity fields (PAFs) for part association, producing a series of body part candidates. The candidates are then matched to a single individual and finally assembled into a pose; here, we only made use of the body parts relevant to the face and hands (nose and wrists).

### Detector evaluation
To evaluate face detector performance, we hand-labeled a "gold set" of labeled frames. To account for the relatively rare appearance of faces in the dataset, we hand-labeled two types of samples: a sample containing a high density of faces (half reported by MTCNN, half by OpenPose) and a random sample from the remaining frames. Each sample was comprised of an equal number of frames taken from each child's video. For wrist detections, the "gold set" was constructed in the same manner, except frames with a high density of wrists came only from detections made by OpenPose.  Faces were classified as present if at least half of the face was showing; wrists were classified as present if any part of the wrist was showing. Two authors labelled the frames independently and resolved disagreements on a case-by-case basis. Precision (hits / hits + false alarms), recall (hits / hits + misses), and F-score (harmonic mean of precision and recall) were calculated for all detectors and are reported in Table 1.

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN DETECTIONS
d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_march2020.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

## make long form version for plotting both face detections on same graph
dlong <- d %>%
  mutate(id = paste(subid,frame)) %>%
  gather(key = detectorType, value = detection, faceMT, faceOP, faceVJ, wristOP) %>%
  mutate(detectorType=as.factor(detectorType))

# for better plotting names
levels(dlong$detectorType) 
# [1] "faceMT"  "faceOP"  "faceVJ"  "wristOP"
levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces","OpenPose-Wrists")

```

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN GROUND TRUTH and load function
ground_truth_faces <- read_csv("../data/ground_truth/ground_truth_faces.csv") %>%
  mutate(sample_type=as.factor(sample_type))

ground_truth_wrists <- read_csv("../data/ground_truth/ground_truth_wrists.csv") %>%
  mutate(sample_type=as.factor(sample_type))

# Renaming for table output
levels(ground_truth_faces$sample_type)=c("High density", "Random")
levels(ground_truth_wrists$sample_type)=c("High density", "Random")

# Function to evaluate detectors
evaluate <- function(a, b) {
  if (a == TRUE) {
    if (a == b) return ("TP") # was face/wrist, detected face/wrist
    else return("FN") # was face/wrist, missed face/wrist
  }
  else {
    if (a == b) return("TN") # was not face/wrist, did not detect face/wrist
    else return("FP") # was not face/wrist, detected face/wrist
  }
}
```

```{r  include=FALSE, echo=FALSE}
## Evalute face and wrist detectors
performanceFaces <- ground_truth_faces  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong)   %>%
  filter(detectorType!="OpenPose-Wrists") %>%
  mutate(detector_is_face = detection, is_face = as.logical(is_face)) %>%
  rowwise() %>%
  mutate(result = evaluate(is_face, detector_is_face)) # always have the gronud truth first in evaluate argument

perfFacesbySampleType <- performanceFaces %>%
  group_by(detectorType, sample_type) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN"))  %>%
  group_by(detectorType, sample_type) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
  # xtable %>%
  # print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)

perfFacesOverall <- performanceFaces %>%
  group_by(detectorType) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  group_by(detectorType) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))

## EVALUATE WRIST DETECTOR
performanceWrists <- ground_truth_wrists  %>%
  mutate(frame=as.numeric(frame), subid=video) %>%
  left_join(dlong) %>%
  filter(detectorType=="OpenPose-Wrists") %>%
  mutate(detector_is_wrist = detection, is_wrist = as.logical(is_wrist)) %>% 
  rowwise() %>%
  mutate(result = evaluate(is_wrist, detector_is_wrist))

perfWristsbySampleType<- performanceWrists %>%
  group_by(detectorType, sample_type) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  group_by(detectorType, sample_type) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r)) 
  xtable %>%
  print(digits = 2, include.rownames=FALSE,table.placement = "H", type="latex", comment = F)

perfWristsOverall<- performanceWrists %>%
  group_by(detectorType) %>%
  summarise(tp=sum(result == "TP"), fp=sum(result == "FP"), fn=sum(result=="FN")) %>%
  summarise(p = tp / (tp + fp), r = tp / (tp + fn), f=( 2 * p * r )/ (p + r))

## get rid of ViolaJones detections from data structure for future analyses
dlong <- dlong %>%
  filter(detectorType!="faceVJ")

```

For face detection, MTCNN outperformed OpenPose when taking into account only the composite F-score (`r round(perfFacesOverall$f[1],2)` MTCNN vs. `r round(perfFacesOverall$f[2],2)` OpenPose). Although MTCNN and OpenPose performed comparably with the random sample, MTCNN performed better on the high density sample (specifically looking at precision), suggesting that OpenPose generated more false positives than MTCNN. ViolaJones performed quite poorly relative to the other detectors, especially with respect to the random sample. We thus use MTCNN detections in the following analyses. For wrist detection, OpenPose performed moderately well (F = `r round(perfWristsOverall$f,2)`) with relatively high precision but low recall on the randomly sampled frames (see Table 1). We thus analyze wrist detections, with the caveat that we are likely underestimating the proportion of hands in the dataset.

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
\hline
Algorithm & Sample\ Type & P & R & F \\ 
\hline
MTCNN-Faces & High density & 0.89 & 0.92 & \textbf{0.90} \\ 
MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
OpenPose-Faces & Random & 0.72 & 0.80 & \textbf{0.76} \\ 
ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
OpenPose-Wrists & High density & 0.66 & 1.00 & 0.79 \\ 
OpenPose-Wrists & Random & 0.88 & 0.29 & 0.44 \\ 
\hline
\end{tabular}
\caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected). P, R, and F denote precision, recall, and F-score, respectively. Scores in bold are the highest F-scores for each sample type.} 
\vspace{-1em}
\end{table}

# Results
First, we report developmental shifts in infants' posture and their orientation relative to their caregiver, consistent with previous literature [@adolph2006; @franchak2017]. Then, we examine how these changes influence children's visual access to faces and wrists/hands across this developmental time range, examining the relative contributions of age vs. postural developments. We also explore how these changes impact the accessibility of faces and wrists/hands during labeling events, both when parents were labeling familiar objects (e.g., 'cat') as well as novel objects (e.g., 'zem'). Finally, we apply the same automated detection method to a egocentric video dataset collected by a different lab [@franch2017] during which children's in-the-moment posture was annotated, replicating their existing results and finding converging support across methods, labs, and paradigms for the impact of postural developments on children's visual access to social information.

```{r}
## Calculate distributions of infants spent in each category separately for posture / orientation
ages_inc_subs <- inc_subs %>%
  select(age.at.test, age.grp, subid)

orientation <- d %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, subid, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(subid, age.grp, orientation) %>%
  left_join(ages_inc_subs)

posture <- d %>%
  filter(!is.na(posture)) %>%
  group_by(age.grp, subid, posture) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  mutate(prop.time = time/sum(time)) %>%
  group_by(subid, age.grp, posture) %>%
  left_join(ages_inc_subs)
```

```{r}
## Make plots of time spent in each orientation / posture
orientation_labels <- c(
  behind ='Caregiver behind',
  close = 'Caregiver close',
  far = 'Caregiver far')

posture_labels = c(
  carry = 'Infant carried',
  prone = 'Infant prone',
  sit = 'Infant sitting',
  stand ='Infant standing'
)

pos_plot <- ggplot(posture, aes(x = age.at.test, y = prop.time, col = posture, size=time)) + 
  geom_point(alpha=.6) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  theme_few() +
  ylim(0,1) +
  facet_grid(~posture, labeller = labeller(posture = posture_labels)) +
  geom_smooth(span=5, alpha=.2) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_color_manual(name = "Posture", values=c("#8878cc","#0099cc","#339966","#eb7928")) +
  theme(legend.position = 'none')

orient_plot <- ggplot(orientation, aes(x = age.at.test, y = prop.time, col = orientation, size=time)) + 
  geom_point(alpha=.8) +
  ylab("Proportion Time") + 
  xlab("Age (months)") + 
  theme_few() +
  ylim(0,1) +
  facet_grid(~orientation,  labeller = labeller(orientation= orientation_labels)) +
  geom_smooth(span=10, alpha=.2) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_color_manual(name = "CG", values=c("#171615","#5c5753","#99928d")) +
  theme(legend.position = 'none')
```

```{r posture, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}
cowplot::plot_grid(pos_plot, orient_plot, nrow=2)
```

```{r}
### Calculate individual amounts of time spent in each orientation by subject
posture_by_orientation <- d %>%
  filter(!is.na(posture)) %>%
  # filter(!is.na(orientation)) %>%
  group_by(subid, posture, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  group_by(subid) %>%
  mutate(prop.time = time/sum(time)) %>%
  mutate(pos_orient = paste0('Infant ',posture, ', CG ', orientation))

posture_by_orientation <- posture_by_orientation %>%
  left_join(ages_inc_subs) %>%
  ungroup() %>%
  mutate(subid = fct_reorder(subid, age.at.test))
```

```{r individ_pos_orient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=3, fig.cap = "Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes"}

## should be a two column figure
ggplot(posture_by_orientation, aes(x = subid, y = prop.time)) + 
  geom_col(aes(fill = pos_orient), position=position_stack()) +
  ylab("Proportion Time") + 
  xlab("Individual subjects (ordered by age)") +
  ylim(0,1) +
  theme_few(base_size=18) +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  scale_fill_manual(name = "Posture & CG Orientation", values=c("#8878cc","#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc", "#98a69f", "#e67300","#ffa64d", "#ffe6cc","#d1b673")) 
```

## Changes in Posture and Orientation
How do childen's in-the-moment posture and orientation change across age? In our dataset, the proportion of time infants spent sitting decreased with age, and the proportion of time infants spent standing increased with infants' age. As children got older, their locomotive abilities allowed them to become more independent. Both 8-month-olds and 12-month-olds spent relatively equivalent amounts of time lying/crawling (i.e., 'prone') which was markedly decreased in the 16-month-olds, who spent most of their time sitting or standing (see Figure \ref{fig:posture}). We also observed changes in children's orientation relative to their caregivers: the 8-month-olds spent more time with their caregiver behind them supporting their sitting positions than did children at other ages (see Figure \ref{fig:posture}). However, we also saw considerable variability across children: some infants spent almost their entire time sitting at a close distance from their caregiver, whereas others showed more considerable variability (see Figure \ref{fig:individ_pos_orient}).

## Changes in Access to Faces and Hands
We first examined the proportion of face and hand detections as a function of children's age without considering their posture (see Figure \ref{fig:detByAge}). While faces tended to be in the field-of-view overall more often than hands, children's head-mounted cameras were angled slightly upward to capture the presence of faces, and hand detections suffered from somewhat lower recall than face detections. Going forward, we thus analyze differences in the relative proportion of faces or hands in view as a function of age, posture, and orientation, rather than comparing them directly. Overall, we observed that 12-month-olds appeared to have visual access to slightly fewer faces than 8 or 16-month-olds, creating a slight U-shaped function in face detections; conversely, hand detections were showed a slight increase across this age range, as reported in prior literature [@Fausey2016].

However, these age-related trends were much smaller than the effect of infant's postural developments on children's visual access to faces and hands. Children's in-the-moment posture was a major factor both in how many faces and hands were in view during the play session, as was their orientation relative to their caregiver. Infants who were sitting or standing had more faces in view than infants who were lying down/crawling (i.e. prone), which was most frequent among 12-month-olds relative to the other age groups  (Figure \ref{fig:detByPosOrient}). When caregivers were behind their children, supporting their children's sitting or standing positions, children saw fewer faces and wrists. In particular, children who were sitting or in front of their caregiver had a high proportion of faces and hands in their field of view. (Figure \ref{fig:detByPosOrient}, lower panel).

These trends were quantified using two generalized linear mixed-effect models estimating the proportion of faces and hands that were in view, with orientation, posture, their interaction, and scaled participant's age as fixed effects, and with random slopes for infants' orientation and posture. A summary of the coefficients of the models can be found in Table 2, confirming that infants who were sitting/standing and in front of their caregivers saw the most faces and hands. When modeling the proportion of faces seen, age was not a significant predictor; however, age remained a significant predictor when modeling the proportion of hands seen by infants. Overall, these results suggest that infants' visual access to social information is largely modulated by their posture and orientation to their caregiver, which is in turn a function of their general locomotor development. Nonetheless, these results also suggest that infants may still increase their attention to hands throughout this age range as they continue to learn about objects, their functions, and their names.

## Access to Faces and Hands During Labeling Events
Our play session was designed to provide parents with opporunities to label objects--both familiar and novel--such that we could examine whether children seek out different kinds of social information around  naming events. We thus explored how face and hand detections changed during object labeling events, analyzing a four-second window around each labeling event (e.g., "Look at the [zem]!"). Every utterance of one of the labelled objects (e.g., 'ball') was counted as a 'labeling event'; timestamps of the beginning of each word were hand-annotated from transcripts ad synchronized with the frame-by-frame detections. Overall, we did not find major differences in face/hand detections during naming events relative to baseline, either for novel objects or for familiar objects (see Figure \ref{fig:detByNaming}). This was true both when we analyzed naming rates with or without taking into account children's posture and orientation relative to their caregiver. These results suggest that either children did not actively seek out social information during this particular play session around naming events, perhaps because were a limited number of possible referents at any given time.

```{r eval=FALSE, include=FALSE}
#wrangle data
## Using MTCNN detections for GLM since it was the best
dGLM_faces <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countFaces=sum(faceMT==TRUE), countNotFaces = sum(faceMT==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

dGLM_wrists <- d %>%
  group_by(subid, posture, orientation) %>%
  summarize(countWrists=sum(wristOP==TRUE), countNotWrists = sum(wristOP==FALSE), age.at.test=age.at.test[1]) %>%
  ungroup()

## run models 
faceGLM <- glmer(cbind(countFaces,countNotFaces) ~ posture * orientation + scale(age.at.test) + 
               (posture + orientation | subid), 
             dGLM_faces, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

wristGLM <- glmer(cbind(countWrists,countNotWrists) ~ orientation * posture + scale(age.at.test) + 
               (posture + orientation | subid), 
             dGLM_wrists, family = "binomial", control=glmerControl(optCtrl=list(maxfun=20000),optimizer=c("bobyqa")))

## output ---- no interaction
# cleanedNames_NoInt = c("Intercept","Close","Far","Sit","Stand","Age") 
# 
# faceGLMOut=(summary(faceGLM))
# rownames(faceGLMOut$coefficients)<-cleanedNames_NoInt
# 
# wristGLMOut=(summary(wristGLM))
# rownames(wristGLMOut$coefficients)<-cleanedNames_NoInt
# 
# xtable(wristGLMOut$coefficients, digits=c(2,2,2,2,3),"Model coefficients from a generalized linear mixed model predicting the proportion of faces vs. wrists seen by infants.") 
# xtable(faceGLMOut$coefficients, digits=c(2,2,2,2,3))

```

```{r detByAge, fig.env="figure", fig.pos = "h", fig.align = "center", fig.width=3, fig.height=2, fig.cap = "Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data."}
detectionsByAge <- dlong %>%
  filter(detectorType!="MTCNN-Faces") %>% # for brevity
  filter(detectorType!="ViolaJones-Faces") %>% # low performance
  filter(!is.na(age.at.test)) %>% # filters NAs for combos of orientation/posture that were empty (see merge_data.Rmd)
  group_by(subid, age.at.test, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  left_join(inc_subs) %>%
  mutate(len = as.numeric(len.min)) ## allows us to scale dots by length of video

ggplot(detectionsByAge, aes(x = age.at.test, y = prop_detected, color = detectorType, size=num_detections)) +
  theme_few(base_size=16) + 
  geom_point(alpha=.5) +
  geom_smooth(size=1, alpha=.3, span=10) +
  scale_x_continuous(breaks=c(8,12,16)) +
  scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
  scale_size_continuous(name = "Number of frames") +
  labs(y = "Proportion detected", x = element_blank()) +
  scale_colour_discrete(name = element_blank()) +
  theme(legend.position="none") +
  facet_grid(~detectorType) 
```

```{r}
### Make dataframes for main plot (detections by orientation & posture)
# filter out detectors we aren't plotting
# filter out detections w/o posture/orientation codings (wasn't always in view) -- this elimiantes "carry"
dlongPlot <-dlong %>%
  filter(detectorType!="MTCNN-Faces") %>%
  filter(detectorType!="ViolaJones-Faces") %>%
  filter(!is.na(posture)) %>%
  filter(!is.na(orientation))

# pop out of actual ages of included subs
ages_inc_subs <- inc_subs %>%
  select(subid, age.at.test, age.grp)

# raw data
detectionsByPostureAndOrient <- dlongPlot %>%
  group_by(age.grp, posture, orientation, subid, detectorType) %>%
  summarise(prop_detected = mean(detection, na.rm=TRUE), num_detections = length(detection)) %>%
  left_join(ages_inc_subs) %>%
  mutate(age.at.test = as.numeric(age.at.test))

# with CIs
detectionsByPostureAndOrientByGroup <- dlongPlot %>%
  group_by(age.grp, orientation, posture, subid, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  group_by(age.grp, orientation, posture, detectorType) %>%
  multi_boot_standard(col = 'prop_detected')

```

```{r detByPosOrient, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95/% bootstrapped confidence intervals"}

# better labels 
orientation_labels <- c(
  behind ='Caregiver behind',
  close = 'Caregiver close',
  far = 'Caregiver far')

posture_labels = c(
  prone = 'Infant prone',
  sit = 'Infant sitting',
  stand ='Infant standing'
)

 ## plot data all together now, with CIs and individual points from each infant
ggplot(detectionsByPostureAndOrientByGroup, aes(x = age.grp, y = mean, col = detectorType)) + 
  # plot raw data
  geom_point(data=detectionsByPostureAndOrient, alpha=.2, aes(x = age.at.test, y=prop_detected, size=num_detections)) +
  # plot trend lines & SEs of trends based on raw data
  geom_smooth(data=detectionsByPostureAndOrient, aes(x = age.at.test, y=prop_detected, weight = num_detections), alpha=0.1, size=0, span=5) +
  stat_smooth(data=detectionsByPostureAndOrient, aes(x = age.at.test, y=prop_detected, weight = num_detections), geom="line", alpha=0.5, size=1, span=5) +
  # plot means and CIs based on age groups
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=1), alpha=.9) +
  # fix up the plot to look nice
  labs(y = "Proportion detections", x = "Age (months)") +
  theme_few(base_size=16) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_discrete(name = 'Detection type') +
  scale_size_continuous(name = 'Number of frames') +
  scale_y_continuous(breaks=seq(0,.60,.30), limits=c(0, .6)) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  facet_grid(orientation ~ posture, labeller = labeller(posture = posture_labels, orientation= orientation_labels))

```

```{r}
source("helper.R")
## summarize detections around naming events
aroundNamings_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.))

aroundNamings_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceOP) %>%
  do(summarize.naming(.)) 
```

```{r}
### Analyze detections around naming events (without posture/orientation)

### first by faces
aroundNamings_Detections_BySub_Faces <- aroundNamings_Faces %>%
  group_by(subid, familiarity, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE))

aroundNamings_Detections_Grouped_Faces <- aroundNamings_Detections_BySub_Faces %>%
  group_by(familiarity, age.grp) %>%
  multi_boot_standard(col = 'prop_detected')

### now by wrists 
aroundNamings_Detections_BySub_Wrists <- aroundNamings_Wrists %>%
  group_by(subid, familiarity, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE))

aroundNamings_Detections_Grouped_Wrists <- aroundNamings_Detections_BySub_Wrists %>%
  group_by(familiarity, age.grp) %>%
  multi_boot_standard(col = 'prop_detected')

### get individual points
detectionsByGroup <- dlongPlot %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  group_by(age.grp, detectorType) %>%
  multi_boot_standard(col = 'prop_detected')
```

```{r}
## Plot faces/wrists around namings by age

faces_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
  # plot baseline
  geom_pointrange(data = (detectionsByGroup %>% filter(detectorType == 'OpenPose-Faces')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=3), alpha=.9) +
  # plot averages by familiarity 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
  # plot familairity detections for each sub
  geom_point(data=aroundNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
  theme_few() +
  labs(y = "Prop. face detections", x = "Age (months)") +
  theme_few(base_size=16) +
  ylim(0,.6) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")

## Faces
wrists_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Wrists, aes(x=age.grp, y=mean, color=familiarity)) +
  # plot baseline
  geom_pointrange(data = (detectionsByGroup %>% filter(detectorType == 'OpenPose-Wrists')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=3), alpha=.9) +
  # plot averages by familiarity 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
  # plot familairity detections for each sub
  geom_point(data=aroundNamings_Detections_BySub_Wrists, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
  theme_few() +
  labs(y = "Prop. wrist detections", x = "Age (months)") +
  theme_few(base_size=16) +
  ylim(0,.6) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")

```

```{r detByNaming, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections during naming events (+/- 2 seconds around label)for familiar and novel objects; these rates are put into context relative to baseline Error bars represent 95 /% bootstrapped confidence intervals."}
cowplot::plot_grid(faces_aroundnamings_plot, wrists_aroundnamings_plot)
```

# Extension to Franchak et al., 2017
In the present work, we found that infant's in-the-moment posture changed with their age, as did infants' orientation relative their caregiver. In a related study with 12-month-olds, @franchak2017 also that children's in-the-moment posture changed the amount of time infants spent looking at faces. Here, we sought to replicate the findings from @franchak2017 using our automated methodology (OpenPose detections), using the footage from their head-mounted cameras (shared on Databrary, @Databrary).

This dataset differed from the present in two key ways that could present a challenge for our automated methodology. First, the environment that infants were immersed in with their caregivers (and experimenters)  was much larger and more varied than the play room used in the present dataset, containing multiple structures and toys in different parts of the room for infants to explore, unlike the present room  dataset which was relatively small (approximately 10' x 10'). Thus, our automated methodology could fail to generalize to scenes from these more complex environments, where detecting faces and hands could arguably be a much harder task. Second, @franchak2017 used a head-mounted eye-tracker, finding that children's posture affected where children were looking within their visual field. Nonetheless, if children are often orienting their heads towards where they are allocating their attention  [@Yoshida20008] then we should expect to find the same pattern of results only when analyzing the information in view.

```{r}
# loads pos_detections used below for franchak dataset
load(file = '../data/franchak_dataset/preprocessed/franchak_dataset_detections.RData')
```

```{r}
faces_by_pos  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_faces = mean(face_openpose)) %>%
  group_by(infant_posture_label) %>%
  multi_boot_standard(col = 'prop_faces')

faces_by_pos_by_sub  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_faces = mean(face_openpose), faces = sum(face_openpose), no_faces = sum(face_openpose==0))

hands_by_pos  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_hands = mean(hand_openpose)) %>%
  group_by(infant_posture_label) %>%
  multi_boot_standard(col = 'prop_hands')

hands_by_pos_by_sub  <- pos_detections %>%
  group_by(sub_id, infant_posture_label) %>%
  summarize(prop_hands = mean(hand_openpose), hands = sum(hand_openpose), no_hands = sum(hand_openpose==0))
```

```{r}
# glmers for franchak dataset
posture_model_faces_franchak <- glmer(cbind(faces,no_faces) ~ infant_posture_label + (1|sub_id), data=faces_by_pos_by_sub, family="binomial")

posture_model_hands_franchak <- glmer(cbind(hands,no_hands) ~ infant_posture_label + (1|sub_id), data=hands_by_pos_by_sub, family="binomial")
```

```{r}
(faces_by_pos_plot <- ggplot(faces_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
   geom_point(data=faces_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label), alpha=.1) +
  geom_line(data=faces_by_pos_by_sub, aes(x=infant_posture_label, y=prop_faces, col=infant_posture_label, group=sub_id), alpha=.1) +
  theme_few(base_size=18) +
  ylab('Prop faces detected') +
  xlab('Infant posture') + 
  ylim(c(0,.3)) +
  theme(legend.position = 'none') +
  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928")) )

(hands_by_pos_plot <- ggplot(hands_by_pos, aes(x=infant_posture_label, y=mean, col=infant_posture_label)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) +
  geom_point(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_hands, col=infant_posture_label), alpha=.1) +
  geom_line(data=hands_by_pos_by_sub, aes(x=infant_posture_label, y=prop_hands, col=infant_posture_label, group=sub_id), alpha=.1) +
  theme_few(base_size=18) +
  ylab('Prop hands detected') +
   xlab('Infant posture') + 
  ylim(c(0,.4)) +
  theme(legend.position = 'none') +
  scale_color_manual(name = "Posture", values=c("#0099cc","#339966","#eb7928")) )
```

```{r franchak, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=8, fig.height=6, fig.cap = "Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals."}
cowplot::plot_grid(faces_by_pos_plot,hands_by_pos_plot)
```

Overall, we found convergence between our two methodologies: we replicated @Franchak2017's main results, finding that the proportion of faces in view was greater when infants were sitting or standing vs. prone. We also found the same patterns of results with hands (i.e., wrist detectinos), which were not originally annotated in @Franch2017 (see Figure \ref{fig:franchak}); results were validated with generalized mixed effect models as in our previous analysis. Interestingly, there was a relatively higher proportion of hands in view in @Franchak2017 relative to the present dataset; this is likely mostly due to the fact that there were often multiple people in view: experimenters stayed in the room to film children and assist with technical difficulties, while in our present dataset the experiments left the room for the majority of the play session. Nonetheless, we still found that posture modulated the hands in view as well, suggesting that this is a major factor that structures infants' access to visual information.

# Discussion