### 2020-08-11 writeup-paper/draft_apa_july2020.tex

\documentclass[english,man]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[unicode=true]{hyperref}
\hypersetup{
            pdftitle={Using automated detections to quantify how postural developments shape the social information in view for young infants},
            pdfkeywords={Postural developments, headcam, social information, face detection, pose
detection},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\makeatother
\shorttitle{Automated detection of the social information}
\author{Bria L. Long\textsuperscript{1}, Alessandro Sanchez\textsuperscript{1}, Allison M. Kraus\textsuperscript{1}, Ketan Agrawal\textsuperscript{1}, \& Michael C. Frank\textsuperscript{1}}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Psychology, Stanford University}
\authornote{

Correspondence concerning this article should be addressed to Bria L. Long, 450 Serra Mall, Stanford CA 94305. E-mail: bria@stanford.edu}
\keywords{Postural developments, headcam, social information, face detection, pose detection\newline\indent Word count: XXX}
\DeclareDelayedFloatFlavor{ThreePartTable}{table}
\DeclareDelayedFloatFlavor{lltable}{table}
\DeclareDelayedFloatFlavor*{longtable}{table}
\makeatletter
\renewcommand{\efloat@iwrite}[1]{\immediate\expandafter\protected@write\csname efloat@post#1\endcsname{}}
\makeatother
\usepackage{csquotes}

\title{Using automated detections to quantify how postural developments shape
the social information in view for young infants}

\date{}

\abstract{
Social information plays a key role in infants' early linguistic and
cognitive development. Yet infants locomotor abilities also develop
continuously as they learn to sit, crawl, stand, and walk -- influencing
the way that infants see the world around them. How do these postural
developments affect infants access to the social information relevant
for early learning? To address this question, we first created a rich
annotated dataset of naturalistic play sessions between infants and
their caregivers with sets of novel and familiar toys (N=36, 8-16 months
of age), recording both egocentric and third-person video. We then
developed an automated analysis method using a pose detection model
(Zhang et al., 2017) to detect the faces and hands in the infant view,
allowing us to analyze the entirety of this dataset. We also applied
this automated method to second egocentric video dataset where
one-year-olds explored a large play area with their caregivers (Franchak
et al., 2017). We found that infants' posture and orientation to their
caregiver changed dramatically across this age range and modulated
infants access to social information; infants who were sitting or
standing with their caregiver at a close distance tended to have the
most faces and hands in their visual field. We also found convergence
across both datasets, confirming the finding that motoric developments
play a significant role in shaping the social information that infants
have access to over development, and suggesting that they may play a
role in emergence of infants' cognitive and linguistic abiltiies. We
suggest that the combined use of head-mounted cameras and the
application of new computer vision techniques is a promising avenue for
understanding the statistics of infants' visual and linguistic
experience as they change over development and for re-utilizing rich
video datasets collected across different labs and diverse contexts.
}

\begin{document}
\maketitle

From their earliest months, infants are deeply engaged in learning from
others. Even newborns tend to prefer to look at faces with direct
vs.~averted gaze (Farroni, Csibra, Simion, \& Johnson, 2002) and young
infants follow overt gaze shifts (Bruner, 1975; Gredeback, Fikke, \&
Melinder, 2010). And as infants reach their first birthday, they also
tend to follow (Yu \& Smith, 2013, 2017) and imitate the gestures of
their caregivers (e.g., pointing).

Infants ability to process these social cues may provide strong
scaffolding for early word learning. Longitudinal studies provide some
evidence for this link: children's level of joint engagement with their
mother at 9-12 months predicts both their receptive and productive
vocabularies (Carpenter, Nagell, \& Tomasello, 1998) and 10 month-olds
who follow an adult's gaze (in an experimental context) have larger
vocabularies at 18 months and throughout the second year of life (Brooks
\& Meltzoff, 2005, 2008). While the relationship between hand-following
and language development has been less well characterized, infants who
follow their caregivers' hands tend to be those who spend more time
jointly attending to events with their caregivers (Yu \& Smith, 2017).

Relatively little work, however, has quantified how often infants see
and use these kinds of social cues in naturalistic learing environments.
By using head-mounted cameras to record what infants see, researchers
have begun to document the infant egocentric perspective (Yoshida \&
Smith, 2008) and to quantify the information -- social and otherwise --
available to infants as they learn. While head-mounted camera data do
not provide explicit information on where infants are attending, some
work suggests that infants orient their head towards what they are
focusing on -- putting those people or objects in view (Yoshida \&
Smith, 2008). Initial recordings from in-lab play sessions revealed a
different view than many imagined: instead of being dominated by their
faces, the infant perspective contained close up views of primarily toys
and hands (Franchak, Kretch, Soska, \& Adolph, 2011; Yoshida \& Smith,
2008; Yu \& Smith, 2017). Subsequent research has revealed that the
infant view is far from a unitary construct, undergoing dramatic changes
as infants grow. Recordings from home environments suggest that the
viewpoints of very young infants -- less than 4 months of age -- do
indeed contain persistent and frequent views of faces (Fausey,
Jayaraman, \& Smith, 2016; Jayaraman, Fausey, \& Smith, 2017) but that
the infant view tends to contain more and more hands as infants grow
older.

These changes in perspective are likely the downstream consequence of a
myriad factors, chief of which may be infants' evolving locomotive
abilities: an infants ability to sit, crawl, stand, or walk structures
the way they interact with the things and people in their world. These
changes can be rapid: the physical body of an infant can lengthen up to
2cm in 24 hours (Lampl, 1993), leading to changes in their motoric
abilities from one day to the next. These motoric developments can be
thought of as gateways that open up entirely new phases of development
(Iverson, 2010), and cause a cascade of changes in infants ability to
interact with their world and the people in it (Karasik, Tamis-LeMonda,
\& Adolph, 2014).

Thus, infants' changing locomotor abilities may shape the social cues
that infants' see and seek out, in turn impacting their cognitive and
linguistic abilities. Supporting this idea, infants experience with
sitting predicts their success at 3D object completion tasks (Soska,
Adolph, \& Johnson, 2010) as well as their receptive vocabulary
(Libertus \& Violi, 2016). And as children begin crawling (Adolph,
Vereijken, \& Denny, 1998) -- or scooting or cruising (Patrick, Noah, \&
Yang, 2012) -- they are no longer constrained to the same spot that
their caregivers last placed them in. Yet while crawlers can choose
where to go and what they see to a much greater degree, they also appear
to spend much of their time in a world populated by floors and knees;
during spontaneous play, toddlers are more likely to look at the floor
while crawling than while walking (Franchak et al., 2011), when they
have full visual access to their environment and the people in it
(Kretch, Franchak, \& Adolph, 2014).

On one view, it is children's ability to stand and walk that
fundamentally changes their ability to access social information (e.g.,
facial expressions, gaze cues, pointing) relative to children who are
still crawling and sitting (Walle, 2016). In turn, this ability to
access more detailed social information could allow infants to learn
words quicker and more efficiently, facilitating language growth (Walle,
2016). Supporting this idea, walking infants make different kinds of
object-related bids for attention from their caregivers than crawling
infants, and tend to hear more action directed statements (e.g.,
\enquote{open it}) (Karasik et al., 2014); further, in an observational
study, Walle and Campos (2014) also found that children who were able to
walk had both higher receptive and productive vocabularies. However, not
all evidence supports this view: some parental report data suggest a
weak relationship between walking and the onset of language (Moore,
Dailey, Garrison, Amatuni, \& Bergelson, 2019), and head-mounted
eye-tracking data have highlighted the role of infants' in-the-moment
posture -- and that of their caregivers' -- in what infants' see,
finding that these factors are better predictors of the social
information in view than their locomotor status per se (i.e.,
\enquote{walker} vs \enquote{crawler}) (Franchak, Kretch, \& Adolph,
2017).

Understanding this nuanced relationship between the social information
in the infant view and their motoric and linguistic development has
remained challenging. Indeed, different kinds of play sessions (e.g.,
exploring novel environments vs.~playing with novel objects) might
modulate the degree to which infants' posture changes the social
information in view. Furthermore, no work to date has directly examined
how the availability of social information changes in naturalistic
language learning contexts. For example, it is unclear if and how the
social information in view changes during the naming events for
different objects (e.g, \enquote{Look, a {[}ball{]}}).

More broadly, the field is in need of computational tools to reutilize
these rich head-mounted camera datasets and to understand the
generalizability of findings across different datasets, populations,
tasks, and age-ranges. While the field has assembled many head-mounted
camera datasets, conducting new analyses on these videos has remained
prohibitively time-consuming due to a lack of computational tools for
annotations. Instead, hundreds of hours of manual annotations have been
required to analyze a fraction of the available frames for a given
analysis. Thus, despite containing a wealth of information about the
structure of parent-child interactions, these datasets have thus gone
dramatically underused.

Recent innovations in computer vision hold promise for automated
annotations of the infant view. Over the past decade, deep neural
networks have become dramatically better at a wide range of visual
tasks, including object classification, Simonyan and Zisserman
(2014){]}, scene categorization (Zhou, Lapedriza, Khosla, Oliva, \&
Torralba, 2017), and pose detection (K. Zhang, Zhang, Li, \& Qiao,
2016), arguably facilitating our understanding of visual perception
(Peterson, Abbott, \& Griffiths, 2018, VanRullen (2017)) and improving
computational neuroscience (Kietzmann, McClure, \& Kriegeskorte, 2018).
However, as most models have been trained on photographs or videos taken
from the adult perspective, it is still unclear how easily these models
can be applied to videos taken from the infant perspective.

In this paper, we thus first validate a model for the automated
detection of faces and hands in the infant view, using a open-source
model of pose detection -- OpenPose(Cao, Simon, Wei, \& Sheikh, 2017; K.
Zhang et al., 2016) -- that provides pose, face, and hand keypoints for
every person in an image. We compare the detection accuracy of OpenPose
with that of both older and more specialized models of face detection
and demonstrate the usability of this off-the-shelf model for
quantifying the faces and hands in the infant view.

We then apply this automated method to a head-mounted camera dataset
collected in our lab to directly examine the role of postural
developments on the social information in the infant view during
language learning. Specifically, we constructed a dataset where infants
and caregivers played freely using a set of novel and familiar objects;
8, 12, and 16 months-olds (N=26; N=12 in each age group) participated to
ensure a wide range of locomotor abilities. Infants' in-the-moment
posture and orientation relative to their caregiver were hand annotated
and all play sessions were transcribed. This cross-sectional design thus
allowed us to directly examine the relative contributions of age
vs.~postural developments on children's visual access to social
information, and how the avaliability of social cues changes relative to
naming events (e.g., \enquote{Yes, you like the {[}ball!{]}}). Broadly,
we predicted that there would be differential access to social
information based on children's postural developments: crawling infants
would see fewer faces/hands because they would primarily be looking at
the ground, while walking toddlers would have access to a richer visual
landscape with greater access to the social information in their
environment.

To examine the robustness of our automated method and the
generalizability of our findings, we applied OpenPose (Cao et al., 2017)
to the dataset collected in Franchak et al. (2017), where one-year-olds
wore head-mounted eye-tracking cameras during a play session and their
in-the-moment posture was hand-annotated. Unlike the present dataset,
infants and caregivers roamed a large, open playroom and explored
different toys placed throughout. We thus analyzed this dataset with the
goal of both (1) validating our automated method on a very different
kind of dataset, attempting to replicate their primary findings
originally obtained with head-mounted eye-tracking (2) extending their
findings to ask whether whether the proportion of hands in view also
varies with infants' posture in an exploratory play session.

\section{Methods}\label{methods}

\subsection{Free-play video dataset}\label{free-play-video-dataset}

A primary goal of this project was to provide a rich, open dataset for
other researchers to examine the effects of postural developments and
labeling behavior during naturalistic parent-child interactions. We thus
invited caregivers of 8, 12, and 16-month-olds to participate in play
sessions where they were provided with pairs of novel and familiar
objects (e.g., a ball and a microfiber duster, called a \enquote{zem})
in a small playroom in a lab (approximately 10' x 10'). Infants wore
head-mounted cameras equipped with a fish-eye lens (see Head-mounted
camera), a third-person camera captured a birds-eye view of the play
session. Using these video data, infants' posture and orientation to
their caregiver were hand-coded and annotated for the entirety of the
play session. All videos were transcribed, and we collected collected
MacArthur CDIs for all children who particpated. All materials have been
made publicly avaliable on Databrary for whom the parents provided
sharing consent (29/36 dyads) via
\url{https://nyu.databrary.org/volume/101}

\subsubsection{Participants}\label{participants}

Our final sample consisted of 36 infants and children, with 12
participants in three age groups: 8 months (6 F), 12 months (7 F), and
16 months (6 F). Participants were recruited from the surrounding
community via state birth records, had no documented disabilities, and
were reported to hear at least 80 percent English at home. Demographics
and exclusion rates are given in the table below (see Table 1)/

\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
\hline
Group & N & \% incl. & Avg age & Avg video length (min) \\
\hline
8 mo. &   12 & 0.46 & 8.71 & 14.41 \\
12 mo. &  12 & 0.40 & 12.62 & 12.71 \\
16 mo. &  12 & 0.31 & 16.29 & 15.10\\
\hline
\end{tabular}
\caption{Exclusion rates and summary demographics for the infants included in the study.} 
\end{table}

To obtain this final sample, we tested 95 children, excluding 59
children for the following reasons: 20 for technical issues related to
the headcam (e.g., failure to record, ran out of battery), 15 for
failing to wear the headcam, 10 for fewer than 4 minutes of headcam
footage, 5 for having multiple adults present, 5 for missing
Communicative Development Inventory (CDI) data, 2 for missing scene
camera footage, 1 for fussiness, and one for sample symmetry.
16-month-olds tolerated the head-mounted camera less well than younger
infants, leading to a higher exclusion rate. All inclusion decisions
were made independent of the results of subsequent analyses.

\subsubsection{Head-mounted camera}\label{head-mounted-camera}

\begin{figure}[H]
\includegraphics[width=0.4\linewidth]{images/view_angle_redone} \caption{Vertical field of view for two different headcam configurations (we used the lower in our current study).}\label{fig:headcam}
\end{figure}

We used a small, head-mounted camera (\enquote{headcam}) that was
constructed from a MD80 model camera attached to a soft elastic
headband. Videos captured by the headcam were 720x480 pixels with 25
frames per second.\footnote{Detailed instructions for creating this
  headcam can be found at
  \url{http://babieslearninglanguage.blogspot.com/2013/10/how-to-make-babycam.html}.}
A fisheye lens was attached to the camera to increase the view angle
from \(32^{\circ}\) horizontal by \(24^{\circ}\) vertical to
\(64^{\circ}\) horizontal by \(46^{\circ}\) vertical (see Figure
\ref{fig:headcam}, above).

Even with the fish-eye lens, the vertical field of view of the camera is
still considerably reduced compared to the child's field of view, which
spans around 100--120\(^{\circ}\) in the vertical dimension by 6-7
months of age (Cummings, Van Hof-Van Duin, Mayer, Hansen, \& Fulton,
1988; Mayer, Fulton, \& Cummings, 1988). As we were originally primarily
interested in the presence of faces in the child's field of view, we
chose to orient the camera upwards to capture the entirety of the
child's upper visual field where the child is likely to see adult faces,
understanding that this decision limited our ability to detect hands
(especially those of the child, which are typically found at the bottom
of the visual field).

\subsubsection{Procedure}\label{procedure}

All parents signed consent documents while children were fitted with the
headcam. If the child was uninterested in wearing the headcam or tried
to take it off, the experimenter presented engaging toys to try to draw
the child's focus away from the headcam. When the child was comfortable
wearing the headcam, the child and caregiver were shown to a playroom
for the free-play session. Parents were shown a box containing three
pairs of familiar and novel objects. These pairs consisted of a ball
paired with a microfiber duster (a \enquote{zem}), a toy car paired with
a cheese grater (a \enquote{manu}), and a brush paired with a back
massager (a \enquote{tima}). Parents were instructed to play with the
object pairs with their child one at a time, \enquote{as they typically
would.} {[}\^{}3{]}:The first few participants played with a different
pair of objects (a toy cat and pedicure foam piece) that was replaced as
some infants chewed persistently on the foam.

All parents confirmed that their child had not previously seen the novel
toys and were instructed to use the novel labels to refer to the toys.
The experimenter then left the playroom for approximately 15-20 minutes,
during which a tripod-mounted camera in the corner of the room recorded
the session and the headcam captured video from the child's perspective.

\subsubsection{Data processing and
annotations}\label{data-processing-and-annotations}

Headcam videos were trimmed such that they excluded the instruction
phase when the experimenter was in the room and were automatically
synchronized with the tripod-mounted videos using FinalCut Pro Software.
These sessions yielded 507 minutes (almost a million frames) of video,
with an average video length of 14.07 minutes (min = 4.53, max = 19.35).

\paragraph{Posture and caregiver orientation
annotations}\label{posture-and-caregiver-orientation-annotations}

We created custom annotations to describe the child's physical posture
(i.e.~standing) and the orientation of the caregiver relative to the
child (e.g.~far away). The child's posture was categorized as being
carried, prone (crawling or lying), sitting, or standing. The
caregiver's orientation was characterized as being close, far, or behind
the child (independent of distance). For the first two annotations
(close/far from the child), the caregiver could either be to the front
or side of the child. All annotations were made by a trained coder using
the OpenSHAPA/Datavyu software (Adolph, Gilmore, Freeman, Sanderson, \&
Millman, 2012). Times when the child was out of view of the tripod
camera were marked as uncodable and were excluded from these
annotations; similarily, times when the child was being carried or the
caregivers were out of the frame were marked as uncodable for caregiver
orientation. On average, posture or orientation was uncodable from 1-2
minutes of data in each child (seconds excluded from analysis for
posture;. M =94.01, SD =225.18; orientation; M =94.01, SD =225.18), and
these rates did not vary substantially with the age of the child. To
assess the reliability of these annotations, a second coder coded videos
from three different children to calculate Cohen's kappa (posture, irr =
XX, orientation, irr = XX).

\paragraph{Naming event annotations}\label{naming-event-annotations}

One coder listened to all of the audio from the play sessions and marked
the exact timestamps whenever one of the novel or familiar objects was
named in any instance (e.g., \enquote{Look at the {[}ball{]}},
\enquote{Can you say {[}zem{]}?}); a second coder listened to the
majority of the play sessions (N=23 sessions) and also annotated all
naming events. To asesess reliability, we calculated the proportion of
naming events detected by the first coder that were also annotated by a
second coder within a sliding window. We found that 82.1\% of naming
events were detected within a 4 second window (+/- 2s), and 70.9\% of
namings events were detected within a 2 second window. We also obtained
full text transcriptions of the entire play sessions (with time stamps
marking 10s intervals). While these full transcriptions are not used in
the present analyses, they have been made publically avaliable for
future research.

\subsection{Face and hand detection}\label{face-and-hand-detection}

We evaluated three detection systems for the ability to measure infants'
access to faces. The first of these is the most commonly-used and widely
available face detection algorithm: Viola-Jones (Viola \& Jones, 2004).
We used this algorithm as a benchmark for performance, as while it can
achieve impressive accuracy in some situations, it is notoriously bad at
dealing with occluded faces (Scheirer, Anthony, Nakayama, \& Cox, 2014).
We next tested the performance of two face detectors that both made use
of relatively recently developed Convolutional Neural Networks (CNNs) to
extract face information. The first algorithm was specifically optimized
for face detection, and the second algorithm was optimized to extract
pose information of all the individuals in an image, operationalized as
information about the position of 18 different body parts. For this
second algorithm (OpenPose; Cao et al., 2017), we used the agent's nose
(one of the body keypoints detected) to operationalize the presence of
faces, as any half of a face necessarily contains a nose.

The OpenPose detector also provided us with the location of an agent's
wrists, which we used as a proxy for hands for two reasons. First, as we
did not capture children's entire visual field, the presence of a wrist
is likely often indicative of the presence of a hand within the field of
view. Second, hands are often occluded by objects when caregivers are
interacting with children, yet still visually accessible by the child
and part of their joint interaction.

\subsubsection{Algorithms}\label{algorithms}

Viola Jones, the first face detection system, made use of a series of
Haar feature-based cascade classifiers (Viola \& Jones, 2004) applied to
each individual frame. The second algorithm (based on work by K. Zhang
et al. (2016)) uses multi-task cascaded convolutional neural networks
(MTCNNs) for joint face detection and alignment, built to perform well
in real-world environments where varying illuminations and occlusions
are present. We used a Tensorflow implementation of this algorithm
available at \url{https://github.com/davidsandberg/facenet}.

The CNN-based pose detector (OpenPose; Cao et al., 2017; Simon, Joo,
Matthews, \& Sheikh, 2017; Wei, Ramakrishna, Kanade, \& Sheikh, 2016)
provided the locations of 18 body parts (ears, nose, wrists, etc.) and
is available at
\url{https://github.com/CMU-Perceptual-Computing-Lab/openpose}. The
system uses a convolutional neural network for initial anatomical
detection and subsequently applies part affinity fields for part
association, producing a series of body part candidates. The candidates
are then matched to a single individual and finally assembled into a
pose; here, we only made use of the body parts relevant to the face and
hands (nose and wrists), though the entire set of keypoints is publicly
avaliable. Each keypoint was accompanied by a confidence score made by
the detector.

\subsubsection{Detector evaluation}\label{detector-evaluation}

To evaluate face detector performance, we hand-labeled a \enquote{gold
set} of frames extracted from the video dataset. To account for the
relatively rare appearance of faces in the dataset, we hand-labeled two
types of samples: a sample containing a high density of faces (half
reported by MTCNN, half by OpenPose) and a random sample from the
remaining frames. Each sample was comprised of an equal number of frames
taken from each child's video. For wrist detections, the \enquote{gold
set} was constructed in the same manner, except frames with a high
density of wrists came only from detections made by OpenPose. Faces were
classified as present if at least half of the face was showing; wrists
were classified as present if any part of the wrist was showing. Two
authors labelled the frames independently and resolved disagreements on
a case-by-case basis. Precision (hits / hits + false alarms), recall
(hits / hits + misses), and F-score (harmonic mean of precision and
recall) were calculated for all detectors.

\section{Results}\label{results}

First, we report the accuracy of the automated detectors, as assessed by
comparison to hand-labelled frames from the free-play video dataset
described above. We then apply one of these automated detectors
(OpenPose) to the entirety of this video dataset, and use these outputs
to examine how postural developments influence children's visual access
to faces and hands from 8-16 months of age, as well as how access to
these social cues changes during labeling events (e.g., do you see the
{[}zem{]}?). Finally, we apply this same automated detector to another
video dataset (Franchak et al., 2017), replicating and extending their
findings on the effects of posture on visual access to social
information.

\begin{figure}[H]
\includegraphics[width=1\linewidth]{images/exampe_detections} \caption{Example detections made by OpenPose from children in each age group.}\label{fig:exampledetections}
\end{figure}

\subsection{Accuracy of automated
detections}\label{accuracy-of-automated-detections}

For face detection, we found that both OpenPose and MTCNN dramatically
outperformed ViolaJones, our baseline model, especially with respect to
the random sample, where ViolaJones generated a relatively high
proportion of both false negatives and false positives. When considering
only the composite F-score across all frames, MTCNN slightly
outperformed OpenPose (0.89 MTCNN vs.~0.83 OpenPose), and MTCNN and
OpenPose performed comparably with the random sample. Generally, MTCNN
exhibited higher precision, whereas OpenPose exhibited higher recall,
and these differences were most pronounced on the randomly sample
frames. In other words, while OpenPose generated slightly more false
positives than MTCNN, MTCNN missed several faces that were accurately
detected by OpenPose. When we restricted our analysis to high-confidence
detections from OpenPose (\textgreater{}.5 confidence; default threshold
for visualization), we found very high precision (P = 0.97), but much
lower recall (R = 0.64) and thus overall performance (F = 0.77),
indicating that these low-confidence detections often indexed actual
faces that were in the infant view. Figure \ref{fig:exampledetections}
shows example successful detections from OpenPose in each age group, and
Figure \ref{fig:failures} shows examples of missed faces as well as
false positives for context.

We next analyzed the viability of OpenPose as a hand detector. Despite
the fact that hand detection is a more computationally challenging
problem (Bambach, Crandall, Smith, \& Yu, 2017), and the fact that we
used wrist keypoints as a proxy for hands, OpenPose performed moderately
well as a hand detector (F = 0.74). OpenPose achieved relatively high
precision -- generating relatively few false positives -- but showedlow
recall on the randomly sampled frames (see Table 1). As with face
detections, when we restricted our analysis to high-confidence
detections, we found much higher precision (P = 0.95), but much lower
recall (R = 0.37) and thus overall performance (F = 0.53).

Thus, one major advantage of OpenPose relative to specialized face
detectors, such as MTCNN, is that it allows the analysis of both the
faces and hands in the infant view with the outputs of only one
algorithm, and analyzing the results of all detections (regardless of
confidence) yielded reasonably accurate results. Going forward, we
analyze face and wrist detections using all detections from OpenPose,
with the caveat that we are likely underestimating the proportion of
hands in the dataset given the lower recall for hand detections.

\begin{figure}[H]
\includegraphics[width=0.5\linewidth]{images/example_failures} \caption{Example failed detections from OpenPose, showing both false positives and false negatives (i.e. missed detections).}\label{fig:failures}
\end{figure}

\begin{table}[ht]
\centering
\begin{tabular}{rllrrr}
\hline
Algorithm & Sample\ Type & P & R & F \\ 
\hline
MTCNN-Faces & High density & 0.89 & 0.92 & 0.90 \\ 
MTCNN-Faces & Random & 0.94 & 0.62 & 0.75 \\ 
OpenPose-Faces & High density & 0.78 & 0.93 & 0.84 \\ 
OpenPose-Faces & Random & 0.72 & 0.80 & 0.76 \\ 
ViolaJones-Faces & High density & 0.96 & 0.44 & 0.60 \\ 
ViolaJones-Faces & Random & 0.44 & 0.38 & 0.41 \\ 
OpenPose-Wrists & High density & 0.66 & 1.00 & 0.79 \\ 
OpenPose-Wrists & Random & 0.88 & 0.29 & 0.44 \\ 
\hline
\end{tabular}
\caption{Detector performance on both high density samples (where proportion of targets detected was high) and random samples (where frames were randomly selected). P, R, and F denote precision, recall, and F-score, respectively.} 
\vspace{-1em}
\end{table}

\subsection{Access to social information during postural
developments}\label{access-to-social-information-during-postural-developments}

\begin{figure}[H]
\includegraphics[width=\textwidth]{figs/posture-1} \caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times where posture was not codable are ommitted for visualization purposes}\label{fig:posture}
\end{figure}

\begin{figure}[H]
\includegraphics[width=\textwidth]{figs/indivPosOrient-1} \caption{Proportion of time spent by each infant in different postures and orientations relative to their caregivers (CG); times when infant was carried or when posture/orientation were not codable are ommitted for visualization purposes}\label{fig:indivPosOrient}
\end{figure}

\subsubsection{Developmental changes in infant posture and caregiver
orientation}\label{developmental-changes-in-infant-posture-and-caregiver-orientation}

We report developmental shifts in infants' posture and their orientation
relative to their caregiver, consistent with previous literature (Adolph
\& Berger, 2006; Franchak et al., 2017). The proportion of time infants
spent sitting decreased with age, and the proportion of time infants
spent standing increased with infants' age. As children got older, their
locomotor abilities allowed them to become more independent. Both
8-month-olds and 12-month-olds spent relatively equivalent amounts of
time lying/crawling (i.e., \enquote{prone}) which was markedly decreased
in the 16-month-olds, who spent most of their time sitting or standing
(see Figure \ref{fig:posture}). We also observed changes in infants'
orientation relative to their caregivers: the 8-month-olds spent more
time with their caregiver behind them supporting their sitting positions
than did children at other ages (see Figure \ref{fig:posture}). However,
we also saw considerable variability across children: some infants spent
almost their entire time sitting at a close distance from their
caregiver, whereas others showed more considerable variability (see
Figure \ref{fig:indivPosOrient}).

\subsubsection{Access to faces and
hands}\label{access-to-faces-and-hands}

First, we examined the proportion of face and hand detections as a
function of infants' age without considering their posture (see Figure
\ref{fig:detByAge}). While faces tended to be in the field-of-view
overall more often than hands, infants' head-mounted cameras were angled
slightly upward to capture the presence of faces, and hand detections
suffered from somewhat lower recall than face detections. We thus only
considered differences in the relative proportion of faces or hands in
view as a function of age, posture, and orientation, rather than
comparing them directly. Overall, we observed that 12-month-olds
appeared to have visual access to slightly fewer faces than 8 or
16-month-olds, creating a slight U-shaped function in face detections;
conversely, hand detections were showed a slight increase across this
age range, as reported in prior work (Fausey et al., 2016).

\begin{figure}[H]
\includegraphics[width=\textwidth]{figs/detByAge-1} \caption{Proportion of faces (left) and wrists (right) detected by the OpenPose model as a function of child's age. Larger dots indicate children who had longer play sessions and thus for whom there was more data.}\label{fig:detByAge}
\end{figure}

\begin{figure}[H]

{\centering \includegraphics[width=\textwidth]{figs/detByPosOrient-1}

}

\caption{Proportion of face / wrist detections by children's age, their posture, and their caregivers orientation. Data points are scaled by the amount of time spent in each orientation/posture combination; times when posture/orientation annotatinos were unavaliable or the infant was carried are not plotted. Error bars represent 95\% bootstrapped confidence intervals}\label{fig:detByPosOrient}
\end{figure}

However, these age-related trends were much smaller than the effect of
infant's postural developments on infants' visual access to faces and
hands. Infants' in-the-moment posture was a major factor both in how
many faces and hands were in view during the play session, as was their
orientation relative to their caregiver. This was quantified using two
generalized linear mixed-effect models estimating the proportion of
faces and hands that were in view, with orientation, posture, their
interaction, and scaled participant's age as fixed effects, and with
random slopes for infants' orientation and posture. A summary of the
coefficients of the models can be found in Tables 2 and 3.

The interaction between infants' posture and their caregiver's
orientation had the most dramatic effect on the social information in
view. When caregivers were behind their infants, supporting their
infants' sitting or standing positions, infants saw fewer faces. When
caregivers were relatively close to their infants, infants who were
sitting or standing had more faces in view (Face detections; infant
sitting and CG close, b=0.99, SE = 0.07, Z = 14.54, P = 0; infant
standing and CG close, b=1.36, SE = 0.08, Z = 16.95, P = 0 than infants
who were lying down/crawling (i.e.~prone). When caregivers were far away
from their infants, face detections were similarity higher (Face
detections; infant sitting and CG far, b = 0.75, SE = 0.07, Z = 10.48, P
= 0; infant standing and CG far, b = 1.40, SE = 0.09, Z = 16.16, P = 0.
Overall, we found that age did not remain a significant predictor in
accounting for the faces in view when modeling infants posture, their
orientation to their caregiver, and the interaction between these two
factors (Face detections; Age (scaled), b = 0.05, SE = 0.12, Z = 0.39, P
= 0.70).

We found a similar pattern of results for wrist detections, even though
there were fewer wrist detections overall in the dataset. Infants saw
fewer wrists when caregivers were behind their infants, supporting their
infants' sitting or standing positions vs when caregivers were
relatively closer to their infants (Hand detections; infant sitting and
CG close, b=0.31, SE = 0.08, Z = 3.72, P = 0.00; infant standing and CG
close, b=0.76, SE = 0.09, Z = 8.46, P = 0 than infants who were lying
down/crawling (i.e.~prone). Wrist detections were clearly highest when
caregivers were far away from their infants (Wrist detections; infant
sitting and CG far, b=0.13, SE = 0.10, Z = 1.27, P = 0.21; infant
standing and CG far, b=1.76, SE = 0.11, Z = 15.87, P = 0. As with faces,
age did not remain a significant predictor in these models (Wrist
detections; Age (scaled), b = 0.16, SE = 0.11, Z = 1.48, P = 0.14).

We directly examined the contributions of posture and orientation by
fitting a reduced version of the full model (Nakagawa \& Schielzeth,
2013) without their fixed effects (both models were run with the maximal
random effects structure). The fixed effects in a model with only the
age of the participants accounted for relatively little variance in the
proportion of faces (marginal \(R^2\) \textless{} 0.01) or hands in view
(marginal \(R^2\) = 0.02). However, when adding infants' posture and
orientation to their caregiver to the model (and their interaction) the
marginal \(R^2\) were higher for both faces (marginal \(R^2\) = 0.28)
and wrists (marginal \(R^2\) = 0.31).\\
Overall, these results suggest that infants' visual access to social
information is largely modulated by their posture and orientation to
their caregiver, which is in turn a function of their general locomotor
development.

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & -4.29 & 0.18 & -23.87 & 0.000 \\ 
  Sit & 0.59 & 0.18 & 3.20 & 0.001 \\ 
  Stand & 0.53 & 0.30 & 1.76 & 0.079 \\ 
  Close & 0.53 & 0.19 & 2.75 & 0.006 \\ 
  Far & 0.15 & 0.26 & 0.56 & 0.574 \\ 
  Age (Scaled) & 0.16 & 0.11 & 1.48 & 0.140 \\ 
  Sit*Close & 0.31 & 0.08 & 3.72 & 0.000 \\ 
  Stand*Close & 0.76 & 0.09 & 8.46 & 0.000 \\ 
  Sit*Far & 0.13 & 0.10 & 1.27 & 0.205 \\ 
  Stand*Far & 1.76 & 0.11 & 15.87 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed models predicting the proportion of wrists seen by infants.} 
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & z value & Pr($>$$|$z$|$) \\ 
  \hline
Intercept & -3.25 & 0.18 & -18.04 & 0.000 \\ 
  Sit & 0.05 & 0.17 & 0.33 & 0.744 \\ 
  Stand & -0.24 & 0.18 & -1.34 & 0.180 \\ 
  Close & 0.01 & 0.19 & 0.07 & 0.942 \\ 
  Far & 0.45 & 0.27 & 1.63 & 0.102 \\ 
  Age (Scaled) & 0.05 & 0.12 & 0.39 & 0.697 \\ 
  Sit*Close & 0.99 & 0.07 & 14.54 & 0.000 \\ 
  Stand*Close & 1.36 & 0.08 & 16.95 & 0.000 \\ 
  Sit*Far & 0.75 & 0.07 & 10.48 & 0.000 \\ 
  Stand*Far & 1.40 & 0.09 & 16.16 & 0.000 \\ 
   \hline
\end{tabular}
\caption{Model coefficients from a generalized linear mixed models predicting the proportion of faces seen by infants.} 
\end{table}

\subsubsection{Social information during naming
events}\label{social-information-during-naming-events}

Our play session was designed to provide parents with opportunities to
label objects -- both familiar and novel -- such that we could examine
whether children saw different kinds of social information around naming
events. In a set of exploratory analyses, we thus analyzed how face and
hand detections changed during object labeling events relative to
baseline, analyzing a four-second window around each labeling event
(e.g., \enquote{Look at the {[}zem{]}!}). Every utterance of one of the
labelled objects (e.g., \enquote{ball}) was counted as a
\enquote{labeling event}; timestamps of the beginning of each word were
hand-annotated and synchronized with the frame-by-frame detections.

To assess whether there were differences in the social information in
view during naming events, we first calculated the proportion of
detections that were in view during this four second window, and
averaged across naming events for each subject as function of whether
the named object as a novel or a familiar objet; this was then then
compared to the baseline proportion of faces in view for each subject in
linear mixed-effect models, with random effects of subjects and fixed
effects of (scaled) age.

Face detections were numerically slightly higher around these novel
naming events relative to baseline, with similar effects across age
groups (8 m.o., \(M_{fam - baseline}\) = =0.02, 12 m.o.,
\(M_{nov - baseline}\) =0.05, 16 m.o. \(M_{nov - baseline}\) =0.02)
however, they were not higher during familiar naming events vs.~baseline
(8 m.o., \(M_{fam - baseline}\) =0.00, 12 m.o., \(M_{fam - baseline}\)
=0.00, 16 m.o. \(M_{fam - baseline}\) =0.02, Conversely, wrist
detections were higher during both familiar (8 m.o.,
\(M_{fam - baseline}\) =0.07, 12 m.o., \(M_{fam - baseline}\) =0.04, 16
m.o. \(M_{fam - baseline}\) =0.08) and novel naming events relative to
baseline across all age groups (Novel; 8 m.o., (8 m.o.,
\(M_{fam - baseline}\) =0.09, 12 m.o., \(M_{fam - baseline}\) =0.08, 16
m.o. \(M_{fam - baseline}\) =0.08). These results were confirmed by a
linear mixed-effect model with scaled aged as a fixed effect and random
intercepts for each subject (Wrist detections; familiar objects
vs.~baseline, b=0.06, SE = 0.01, t = 4.33, P = 0.00'; Novel objects
vs.~baseline, b=0.08, SE = 0.02, t = 5.21, P = 0.

Overall, these exploratory results suggest that children may tend to see
more hands around naming events. This finding is consistent with the
possibility that caregivers may change how they interact with their
infant when presenting them with objects (Gogate, Bahrick, \& Watson,
2000, Gogate, Bolzani, and Betancourt (2006)). For example, caregivers
may tend to simultaneously name objects when demonstrating their
affordances or simply when pointing to them. In turn, infants may be
sensitive to these naming events and orient their attention towards
their caregiver, consistent with other accounts positing infants'
sensitivity to social cues in early word learning environments
(Yurovsky, 2018, Yurovsky and Frank (2017), Frank, Simmons, Yurovsky,
and Pusiol (2013)).

\begin{figure}[H]

{\centering \includegraphics[width=\textwidth]{figs/detByNaming-1}

}

\caption{Proportion of face / wrist detections during naming events (+/- 2 seconds around label) for familiar and novel objects; these rates are put into context relative to baseline. Error bars represent 95\% bootstrapped confidence intervals. Grey lines connect points from individual subjects.}\label{fig:detByNaming}
\end{figure}

\subsection{Extension to Franchak et a.,
2017}\label{extension-to-franchak-et-a.-2017}

In the present work, we found that infant's in-the-moment posture
changed with their age, as did infants' orientation relative their
caregiver. In a related study with 12-month-olds, Franchak et al. (2017)
also found that infants' in-the-moment posture changed the proportion of
time infants spent looking at faces. Here, we sought to replicate these
findings using our automated methodology (OpenPose detections), using
the footage from their head-mounted cameras (D. A. Simon, Gordon,
Steiger, \& Gilmore, 2015) for two reasons. First, we sought to validate
our novel method, which could fail to generalize to scenes from these
more complex environments, where detecting faces and hands could
arguably be a much harder task. Second, we sought to replicate the
effects of infants' in-the-moment posture on differences in visual
access to hands in an independent dataset.

\subsubsection{Comparison between video
datasets}\label{comparison-between-video-datasets}

First, despite having different head-mounted camera setups, the view
angle of the two head-mounted cameras were only slightly different
(\(52.2^{\circ}\) horizontal by \(42.2^{\circ}\) in Franchak et al.
(2017), here \(64^{\circ}\) horizontal by \(46^{\circ}\) vertical).
However, the environment that infants were immersed in with their
caregivers (and experimenters) was much larger and more varied than the
play room used in the present dataset, containing multiple structures
and toys in different parts of the room for infants to climb, explore,
and interact with, and infants were unconstrained and allowed to freely
wander the room. In contrast, the play room used in Study 1 was
relatively small (approximately 10' x 10') and was set-up with focused
play with the pairs of novel and familiar objects. Further, multiple
people were present during the play session in Franchak et al. (2017)--
including their caregiver and two experimenters -- whereas in Study 1
the experimenters left the room during the play session.

\begin{figure}[H]

{\centering \includegraphics[width=\textwidth]{figs/franchak-1}

}

\caption{Proportion of face / wrist detections for 12-month-olds in Franchak et al., 2017 as a function of children's in-the-moment posture. Error bars represent 95 percent bootstrapped confidence intervals.}\label{fig:franchak}
\end{figure}

\subsubsection{Replication using automated
detections}\label{replication-using-automated-detections}

Overall, we found convergence between our two methodologies, replicating
the main results from Franchak et al. (2017), and finding that the
proportion of faces detected was greater when infants were sitting or
standing vs.~prone. We found this result regardless of whether we used
all detections (see Figure \ref{fig:franchak}, panel A; (proportion of
frames with face detections; Prone: M =0.13\%, Sitting: M =0.20\%,
Upright: M =0.22\%) or restricted our analyses to high-confidence
detections (percentage of frames with face detections; Prone: M =0.03,
Sitting: M =0.06, Upright: M =0.05). These results were confirmed in
generalized linear-mixed models, as in the first study (Sitting
vs.~Prone, b=0.34, SE = 0.02, Z = 23.37, P = 0. Upright vs.~Prone,
b=0.38, SE = 0.02, Z = 24.01, P = 24.01).

The most noticeable difference between our analysis methods is the
overall proportion of frames with faces detected by OpenPose vs.~the
proportion of frames in which infants were judged to be looking at faces
in Franchak et al. (2017). Across the entire session in Franchak et al.
(2017), infants looked at faces on 4.7\% of frames (Franchak et al.,
2017). When we used all detections from OpenPose, we found a higher
proportion of faces: M = 0.22, or 21.80\%. When we restricted our
results to only high-confidence detections, we found M = 0.06, closer to
the original values reported by Franchak et al. (2017). However, the
above analyses on the accuracy of these two methods suggest that
high-confidence detections dramatically underestimate the number of
faces in view. Thus, we suspect that these differences are due to both
(1) the fact head-mounted eye-trackers may sometimes underestimate the
proportion of faces attended due to calibration issues, and that (2)
OpenPose may sometimes detect faces that are in view for infants but
that infants may not be foveating.

We also found that infants' in-the-moment posture also modulated the
proportion of hands that were in view (i.e., wrist detections), though
these were not originally annotated in Franchak et al. (2017) (see
Figure \ref{fig:franchak}, (proportion of frames with wrist detections;
Prone: M =0.15; Sitting: M =0.21; Upright: M =0.22). These results were
confirmed in generalized linear-mixed models, as in the first study
(Sitting vs.~Prone, b=0.27, SE = 0.01, Z = 19.09, P = 0. Upright
vs.~prone, b=0.30, SE = 0.02, Z = 19.73, P = 19.73).

Overall, these analyzes extend and validate previous work, replicating
the results in Study 1 that infants' in-the-moment posture modulated the
proportion of hands in view, and suggesting that posture is a major
factor that structures infants' access to visual information broadly
construed.

\section{General Discussion}\label{general-discussion}

What social cues do infants see as they learn language, and how does
this change as they grow and are start to locomote themselves in their
environments? We explored this question using video data from
head-mounted cameras from two naturalistic datasets of parent-child
interactions: a cross-sectional database of naturalistic play sessions
from 8-16 month-olds with sets of novel and familiar toys, and database
of head-mounted camera videos from one-year-olds who explored a large
play area from Franchak et al. (2017). To analyze these datasets, we
developed a novel method using a pose detection model to automate the
annotation of the social information in the infant view, here
operationalized as the presence of the faces and hands of their
caregiver; these annotations were then synced with manual annotations of
infants in-the-moment posture from third-person videos.

Despite being trained on the adult perspective, the pose detector we
used (OpenPose, Cao et al. (2017)) was able to generalize relatively
well to the infant viewpoint, achieving comparable precision and
accuracy as a face detector relative to a state-of-the-art model
optimized specifically for detecting faces in natural scenes (MTCNN, K.
Zhang et al. (2016){]}. While OpenPose had relatively low recall as a
hand detector -- missing some hands that were in the infant view -- it
made comparable rates of false alarms. In both cases, we found that
overall performance was maximized when all detections were included,
regardless of their confidence, suggesting that some low-confidence face
and hand detection still index actual faces and hands that were seen by
infants. Thus, while imperfect, we suggest that OpenPose can be applied
to infant egocentric videos for the extraction of the social information
in the infant viewpoint, reducing the burden of manual annotations and
promoting the reusability of rich video datasets for further analyses.
The use of this automated methodology allowed us to easily annotate the
entirety of our dataset -- additionally analyzing the social information
around naming events -- and to re-analyze the data from Franchak et al.
(2017) to replicate our findings in a very different kind of play
session.

Broadly, our results replicate and extend previous work, first by
showing systematic changes in infants' in-the-moment posture and their
orientation relative to their caregivers (Adolph \& Franchak, 2017);
older children spent more time standing and less time sitting, and older
infants' caregivers spent less time supporting their standing or sitting
postures. Motor development changes dramatically at the same time that
children are breaking into language learning. Using these automated
detections, we found that infants' changing posture and orientation to
their caregiver jointly shaped the amount of social information that was
in their view during one-on-one play sessions with their caregivers:
Children saw the most faces/hands when they were sitting or standing and
close to their caregiver vs.~crawling or prone. These same findings were
recapitulated in a second dataset collected by Franchak et al. (2017)
with one-year-olds: sitting and upright infants saw more faces--and
hands--than infants who were prone. Motor development appears to
modulate how infants experience their visual world and the social
information in it.

While exploratory, our results also suggest that infants saw a greater
proportion of hands during naming events, hinting that children may have
been orienting towards their caregiver when they heard labels for
objects. While this effect was not present for faces, other work
(Yoshida \& Smith, 2008; Yu \& Smith, 2013, 2017), including Franchak et
al. (2017), has found that infants spend much more time looking at the
toys vs.~their caregiver's faces during these play sessions, and
highlighted the importance of hand-following as a component of joint
attention (Yu \& Smith, 2017). Further, given that there were only two
possible referents in the room at a time----and one of them was always a
familiar category (i.e., car, kitty) -- this particular play session did
not present many opportunities where children would need to use gaze
cues to disambiguate referents.

Importantly, however, all of these findings come from observational,
in-lab datasets, posing limits on their generalizability. Future work is
needed to relate these the slices of experience captured within in-lab
play sessions and infants' everyday experiences (Clerkin, Hart, Rehg,
Yu, \& Smith, 2017; Fausey et al., 2016). More broadly, though
observational findings allow us to document developmental change and
identify potential causal pathways, they cannot confirm them. Locomotive
abilities are of course only part of a cascade of changes in infants'
abilities and experiences, and these analyzes document only a fraction
of this broader, multifaceted trajectory. As children grow and change,
the activities in which they engage with their caregivers are likely to
also vary, leading to differences in the distribution of social cues
that they experience that may not be captured in these in-lab play
sessions.

Nonetheless, we think that this approach hold promise for documenting
these developmental trajectories and for generating hypotheses about
potential causal pathways. Indeed, understanding the relationship
between different domains of developmental changes in naturalistic
contexts has been a persistent challenge for developmental psychology.
However, the field of computer vision has advanced dramatically in
recent years, creating a new generation of algorithmic tools that deal
better with noisier, more complicated datasets and extract richer
information. We hope that these new tools can now be leveraged to
examine the consequences of the changing infant perspective for
linguistic, cognitive, and social development.

\section{Acknowledgements}\label{acknowledgements}

Thanks to Kaia Simmons, Kathy Woo, and Aditi Maliwal for help in
recruitment, data collection, and annotation. This work was funded by a
Jacobs Foundation Fellowship to MCF, a John Merck Scholars award to MCF,
and NSF \#1714726 to BLL. An earlier version of this work was presented
to the Cognitive Science Society in Frank et al. (2013) and Sanchez,
Long, Kraus, and Frank (2018).

\newpage

\section{References}\label{references}

\begingroup
\setlength{\parindent}{-0.5in} \setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\hypertarget{ref-adolph2006motor}{}
Adolph, K. E., \& Berger, S. E. (2006). Motor development.
\emph{Handbook of Child Psychology}.

\hypertarget{ref-adolph2017development}{}
Adolph, K. E., \& Franchak, J. M. (2017). The development of motor
behavior. \emph{Wiley Interdisciplinary Reviews: Cognitive Science},
\emph{8}(1-2), e1430.

\hypertarget{ref-adolph2012toward}{}
Adolph, K. E., Gilmore, R. O., Freeman, C., Sanderson, P., \& Millman,
D. (2012). Toward open behavioral science. \emph{Psychological Inquiry},
\emph{23}(3), 244--247.

\hypertarget{ref-adolph1998roles}{}
Adolph, K. E., Vereijken, B., \& Denny, M. (1998). Roles of variability
and experience in development of crawling. \emph{Child Development},
\emph{69}(1299), 312.

\hypertarget{ref-bambach2017}{}
Bambach, S., Crandall, D. J., Smith, L. B., \& Yu, C. (2017). An
egocentric perspective on active vision and visual object learning in
toddlers. In \emph{Proceedings of the seventh joint ieee conference on
development and learning and on epigenetic robotics}.

\hypertarget{ref-brooks2005}{}
Brooks, R., \& Meltzoff, A. (2005). The development of gaze following
and its relation to language. \emph{Developmental Science}, \emph{8}(6),
535--543.

\hypertarget{ref-brooks2008}{}
Brooks, R., \& Meltzoff, A. N. (2008). Infant gaze following and
pointing predict accelerated vocabulary growth through two years of age:
A longitudinal, growth curve modeling study. \emph{Journal of Child
Language}, \emph{35}(1), 207--220.

\hypertarget{ref-bruner1975}{}
Bruner, J. (1975). From communication to language: A psychological
perspective. \emph{Cognition}, \emph{3}(3), 255--287.

\hypertarget{ref-cao2017realtime}{}
Cao, Z., Simon, T., Wei, S.-E., \& Sheikh, Y. (2017). Realtime
multi-person 2D pose estimation using part affinity fields. In
\emph{CVPR}.

\hypertarget{ref-carpenter1998}{}
Carpenter, M., Nagell, K., \& Tomasello, M. (1998). Social cognition,
joint attention, and communicative competence from 9 to 15 months of
age. \emph{Monographs of the Society for Research in Child Development},
\emph{63}(4).

\hypertarget{ref-clerkin2017}{}
Clerkin, E. M., Hart, E., Rehg, J. M., Yu, C., \& Smith, L. B. (2017).
Real-world visual statistics and infants' first-learned object names.
\emph{Phil. Trans. R. Soc. B}, \emph{372}(1711), 20160055.

\hypertarget{ref-cummings1988}{}
Cummings, M., Van Hof-Van Duin, J., Mayer, D., Hansen, R., \& Fulton, A.
(1988). Visual fields of young children. \emph{Behavioural and Brain
Research}, \emph{29}(1), 7--16.

\hypertarget{ref-farroni2002eye}{}
Farroni, T., Csibra, G., Simion, F., \& Johnson, M. H. (2002). Eye
contact detection in humans from birth. \emph{Proceedings of the
National Academy of Sciences}, \emph{99}(14), 9602--9605.

\hypertarget{ref-fausey2016}{}
Fausey, C. M., Jayaraman, S., \& Smith, L. B. (2016). From faces to
hands: Changing visual input in the first two years. \emph{Cognition},
\emph{152}, 101--107.

\hypertarget{ref-franchak2017see}{}
Franchak, J. M., Kretch, K. S., \& Adolph, K. E. (2017). See and be
seen: Infant--caregiver social looking during locomotor free play.
\emph{Developmental Science}.

\hypertarget{ref-franchak2011}{}
Franchak, J. M., Kretch, K. S., Soska, K. C., \& Adolph, K. E. (2011).
Head-mounted eye tracking: A new method to describe infant looking.
\emph{Child Development}, \emph{82}(6), 1738--1750.

\hypertarget{ref-frank2013}{}
Frank, M. C., Simmons, K., Yurovsky, D., \& Pusiol, G. (2013).
Developmental and postural changes in children's visual access to faces.
In \emph{Proceedings of the 35th annual meeting of the cognitive science
society} (pp. 454--459).

\hypertarget{ref-gogate2000study}{}
Gogate, L. J., Bahrick, L. E., \& Watson, J. D. (2000). A study of
multimodal motherese: The role of temporal synchrony between verbal
labels and gestures. \emph{Child Development}, \emph{71}(4), 878--894.

\hypertarget{ref-gogate2006attention}{}
Gogate, L. J., Bolzani, L. H., \& Betancourt, E. A. (2006). Attention to
maternal multimodal naming by 6-to 8-month-old infants and learning of
word--object relations. \emph{Infancy}, \emph{9}(3), 259--288.

\hypertarget{ref-gredeback2010development}{}
Gredeback, G., Fikke, L., \& Melinder, A. (2010). The development of
joint visual attention: A longitudinal study of gaze following during
interactions with mothers and strangers. \emph{Developmental Science},
\emph{13}(6), 839--848.

\hypertarget{ref-iverson2010}{}
Iverson, J. M. (2010). Developing language in a developing body: The
relationship between motor development and language development.
\emph{Journal of Child Language}, \emph{37}(2), 229--261.

\hypertarget{ref-jayaraman2017faces}{}
Jayaraman, S., Fausey, C. M., \& Smith, L. B. (2017). Why are faces
denser in the visual experiences of younger than older infants?
\emph{Developmental Psychology}, \emph{53}(1), 38.

\hypertarget{ref-karasik2014}{}
Karasik, L. B., Tamis-LeMonda, C. S., \& Adolph, K. E. (2014). Crawling
and walking infants elicit different verbal responses from mothers.
\emph{Developmental Science}, \emph{17}(3), 388--395.

\hypertarget{ref-kietzmann2018deep}{}
Kietzmann, T. C., McClure, P., \& Kriegeskorte, N. (2018). Deep neural
networks in computational neuroscience. \emph{BioRxiv}, 133504.

\hypertarget{ref-kretch2014}{}
Kretch, K. S., Franchak, J. M., \& Adolph, K. E. (2014). Crawling and
walking infants see the world differently. \emph{Child Development},
\emph{85}(4), 1503--1518.

\hypertarget{ref-lampl1993evidence}{}
Lampl, M. (1993). Evidence of saltatory growth in infancy.
\emph{American Journal of Human Biology}, \emph{5}(6), 641--652.

\hypertarget{ref-libertus2016sit}{}
Libertus, K., \& Violi, D. A. (2016). Sit to talk: Relation between
motor skills and language development in infancy. \emph{Frontiers in
Psychology}, \emph{7}, 475.

\hypertarget{ref-mayer1988}{}
Mayer, D., Fulton, A., \& Cummings, M. (1988). Visual fields of infants
assessed with a new perimetric technique. \emph{Investigative
Ophthalmology \& Visual Science}, \emph{29}(3), 452--459.

\hypertarget{ref-moore2019point}{}
Moore, C., Dailey, S., Garrison, H., Amatuni, A., \& Bergelson, E.
(2019). Point, walk, talk: Links between three early milestones, from
observation and parental report. \emph{Developmental Psychology}.

\hypertarget{ref-nakagawa2013general}{}
Nakagawa, S., \& Schielzeth, H. (2013). A general and simple method for
obtaining r2 from generalized linear mixed-effects models. \emph{Methods
in Ecology and Evolution}, \emph{4}(2), 133--142.

\hypertarget{ref-patrick2012developmental}{}
Patrick, S. K., Noah, J. A., \& Yang, J. F. (2012). Developmental
constraints of quadrupedal coordination across crawling styles in human
infants. \emph{Journal of Neurophysiology}, \emph{107}(11), 3050--3061.

\hypertarget{ref-peterson2018evaluating}{}
Peterson, J. C., Abbott, J. T., \& Griffiths, T. L. (2018). Evaluating
(and improving) the correspondence between deep neural networks and
human representations. \emph{Cognitive Science}, \emph{42}(8),
2648--2669.

\hypertarget{ref-sanchez2018postural}{}
Sanchez, A., Long, B., Kraus, A. M., \& Frank, M. C. (2018). Postural
developments modulate children's visual access to social information.

\hypertarget{ref-scheirer2014perceptual}{}
Scheirer, W. J., Anthony, S. E., Nakayama, K., \& Cox, D. D. (2014).
Perceptual annotation: Measuring human vision to improve computer
vision. \emph{IEEE Transactions on Pattern Analysis and Machine
Intelligence}, \emph{36}(8), 1679--1686.

\hypertarget{ref-simon2015databrary}{}
Simon, D. A., Gordon, A. S., Steiger, L., \& Gilmore, R. O. (2015).
Databrary: Enabling sharing and reuse of research video. In
\emph{Proceedings of the 15th acm/ieee-cs joint conference on digital
libraries} (pp. 279--280).

\hypertarget{ref-simon2017hand}{}
Simon, T., Joo, H., Matthews, I., \& Sheikh, Y. (2017). Hand keypoint
detection in single images using multiview bootstrapping. In
\emph{CVPR}.

\hypertarget{ref-simonyan2014very}{}
Simonyan, K., \& Zisserman, A. (2014). Very deep convolutional networks
for large-scale image recognition. \emph{ArXiv Preprint
ArXiv:1409.1556}.

\hypertarget{ref-soska2010systems}{}
Soska, K. C., Adolph, K. E., \& Johnson, S. P. (2010). Systems in
development: Motor skill acquisition facilitates three-dimensional
object completion. \emph{Developmental Psychology}, \emph{46}(1), 129.

\hypertarget{ref-vanrullen2017perception}{}
VanRullen, R. (2017). Perception science in the age of deep neural
networks. \emph{Frontiers in Psychology}, \emph{8}, 142.

\hypertarget{ref-viola2004robust}{}
Viola, P., \& Jones, M. J. (2004). Robust real-time face detection.
\emph{International Journal of Computer Vision}, \emph{57}(2), 137--154.

\hypertarget{ref-walle2016infant}{}
Walle, E. A. (2016). Infant social development across the transition
from crawling to walking. \emph{Frontiers in Psychology}, \emph{7}, 960.

\hypertarget{ref-walle2014}{}
Walle, E. A., \& Campos, J. J. (2014). Infant language development is
related to the acquisition of walking. \emph{Developmental Psychology},
\emph{50}(2), 336.

\hypertarget{ref-wei2016cpm}{}
Wei, S.-E., Ramakrishna, V., Kanade, T., \& Sheikh, Y. (2016).
Convolutional pose machines. In \emph{CVPR}.

\hypertarget{ref-yoshida2008}{}
Yoshida, H., \& Smith, L. (2008). What's in view for toddlers? Using a
head camera to study visual experience. \emph{Infancy}, \emph{13},
229--248.

\hypertarget{ref-yu2013joint}{}
Yu, C., \& Smith, L. B. (2013). Joint attention without gaze following:
Human infants and their parents coordinate visual attention to objects
through eye-hand coordination. \emph{PloS One}, \emph{8}(11).

\hypertarget{ref-yu2017hand}{}
Yu, C., \& Smith, L. B. (2017). Hand--eye coordination predicts joint
attention. \emph{Child Development}, \emph{88}(6), 2060--2078.

\hypertarget{ref-yurovsky2018communicative}{}
Yurovsky, D. (2018). A communicative approach to early word learning.
\emph{New Ideas in Psychology}, \emph{50}, 73--79.

\hypertarget{ref-yurovsky2017beyond}{}
Yurovsky, D., \& Frank, M. C. (2017). Beyond naïve cue combination:
Salience and social cues in early word learning. \emph{Developmental
Science}, \emph{20}(2), e12349.

\hypertarget{ref-zhang2016}{}
Zhang, K., Zhang, Z., Li, Z., \& Qiao, Y. (2016). Joint face detection
and alignment using multitask cascaded convolutional networks.
\emph{IEEE Signal Processing Letters}, \emph{23}(10), 1499--1503.

\hypertarget{ref-zhou2017places}{}
Zhou, B., Lapedriza, A., Khosla, A., Oliva, A., \& Torralba, A. (2017).
Places: A 10 million image database for scene recognition. \emph{IEEE
Transactions on Pattern Analysis and Machine Intelligence},
\emph{40}(6), 1452--1464.

\endgroup

\end{document}