### 2020-08-11 analysis/0_sanity_checks/check_posture_frames.Rmd
port analysis scripts to analysis folder; rename

---
title: "check_posture_times"
author: "Bria Long"
date: "2/15/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Step 0: Libraries and helper functions
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls()) # remove all objects from current workspace
library(dplyr)
library(readr)
library(magrittr)
library(tidyr)
library(assertthat)
source("helper.R") # Contains critical helper functions for merging annotations
```

## Step 3. Merge demographics and manual annotations
```{r}
# load demographics and pose / orientation / timing
demo.data <- read_csv("../data/demographics/demographics.csv") %>% 
  select(-ra, -assist, -len) ## -dot -dib fields did not exist, deleted bll
```

```{r}
######## ADD POSTURE CODING #########
add.posture <- function(x) {
  fname <- paste("../data/posture/",
                 x$subid[1],
                 ".csv",
                 sep="")

  if (file.exists(fname)) {
    print(fname)
    postures <- read_csv(fname)
    postures <- postures[order(postures$start),] ## add line to sort by time in case weird coding order
    postures <- regularize.postures(subset(postures,code=="posture"))
    x$posture <- factor(NA,levels=levels(postures$posture))
    x$orientation <- factor(NA,levels=levels(postures$orientation))

    # for each row of p, populate posture and orientation to x
    for (i in 1:nrow(postures)) {
      range <- x$time > postures$start[i] & x$time <= postures$end[i]
      x$posture[range] <- postures$posture[i]
      x$orientation[range] <- postures$orientation[i]
    }
  } else {
    print(paste0("**** ", x$subid[1], " is missing posture ****"))
    x$posture <- NA
    x$orientation <- NA
  }

  return(x)
}
```