### 2020-08-11 analysis/0_sanity_checks/length_sanity_check.Rmd

---
title: "xsface - file length sanity checking"
author: "Mike Frank, continued by Bria Long & Alessandro Sanchez"
date: "July 11, 2016"
output:
  html_document:
  toc: true
---

The goal of this document is to document processing steps for the xsface data and ensure that three different datasets match up:

1. source files for movies
2. demographic sheet (containing session lengths)
3. detector outputs (MTCNN/ OpenPose)
4. timing outputs from ffprobe

```{r}
rm(list=ls())
library(readr)
library(magrittr)
library(dplyr)
library(lubridate)
library(stringr)

```

read in lengths on movies, demographics, detections, and frame time outputs

```{r}
demo <- read_csv("../data/demographics/demographics.csv")
movies <- read_csv("../data/video_stats/new_movie_lengths.csv") 
MTCNN <- read_csv("../data/final_detector_output/mtcnn3.csv")
openPose <- read_csv("../data/final_detector_output/openpose_results_truncated_2.csv")
frameTimes <- read_csv("../data/video_stats/frametimechecks_showframes.csv")
```

Clean up measurements.

```{r}
demo %<>% 
  rename(recorded_len = len_seconds)

## movie lengths
movies$frameCount=movies$fps*movies$length
movies %<>% 
  rename(last_frame_movie = frameCount, 
         length_movie = length)

# standardize
MTCNN$frame=as.numeric(MTCNN$frame)
MTCNN$subid=MTCNN$video
openPose$frame=as.numeric(openPose$frame)
openPose$subid=openPose$name

## get out max frame
MTCNN.lens <- MTCNN %>%
  group_by(subid) %>%
  summarise(frames_dets_mtcnn = max(frame))

openPose.lens <- openPose %>%
  group_by(subid) %>%
  summarise(frames_dets_op = max(frame))

## get frames from frame time output (ffprobe)
frameTimes <- frameTimes %>%
  mutate(last_frame_times = last.frame) %>%
  mutate(subid = str_replace_all(subid,"-","_")) %>%
  mutate(subid = str_sub(as.character(subid),start=1,end=7))

```

Merge.

```{r}
all_data <- left_join(demo, movies) %>%
  left_join(MTCNN.lens) %>%
  left_join(openPose.lens) %>%
  left_join(frameTimes) %>%
  select(subid, recorded_len, length_movie, last_frame_movie, last_frame_times, frames_dets_mtcnn, frames_dets_op) %>%
  mutate(det_frame_err = abs(last_frame_movie - frames_dets_mtcnn) > 100,
         demo_time_err = abs(recorded_len - length_movie) > 1,
         frame_time_diff = last_frame_times - frames_dets_mtcnn) %>%
  data.frame

sum(all_data$frame_time_diff)
```

Print.

```{r}
knitr::kable(all_data, digits = 2)
```

=====

<!-- XS_1229 has been trimmed (headcam came off) and therefore there are different movie lenghts. -->

<!-- Old notes pre-cleaning of dataset (irrelevant) -->

<!-- + 0825 - unknown, ~500 frames off (15s), spot checks seem to match decently well.  -->
<!-- + 1205, 1211, and 1225 were post-trimed, hypothesis is that the _QT files are wrappers to source files that trim length to appropriate?  -->
<!-- + 1628 and 1639 both off by a lot of frames, unknown. -->

<!-- to check in the videos -->
<!-- + 1229 is reported to be much longer in the demographics file, perhaps this is post-trimming?  -->
<!-- + 1612 is reported to be shorter than it is, check this.  -->
<!-- + 1622 is reported to be longer than it is, check this.  -->