### 2020-08-11 analysis/4_extra_plots/extra_plotting_code.Rmd

---
title: "extra_plotting_code"
author: "Bria Long"
date: "3/18/2020"
output: html_document
---

```{r}
# ggplot(posture, aes(y = posture, x = prop.time, fill=posture, col=posture)) + 
#   ggridges::geom_density_ridges(alpha=.7) +
#   theme_few() +
#   facet_grid(~age.grp) 
# coord_flip()

# p1=ggplot(posture, aes(x = factor(age.grp), y = prop.time, col = posture)) + 
#   geom_violin() + 
#   ylab("Proportion Time") + 
#   xlab("") + 
#   theme_few() +
#   scale_x_discrete( labels=c("8 m.", "12 m.", "16 m."))
# 
# p2=ggplot(orientation, aes(x = factor(age.grp), y = prop.time, col = orientation)) + 
#   geom_boxplot() + 
#   ylab("Proportion Time") + 
#   xlab("") + 
#   theme_few() +
#   scale_x_discrete( labels=c("8 m.", "12 m.", "16 m.")) +
#   theme(axis.ticks = element_blank())
# 
# ggarrange(p1,p2)
```

```{r detByPosOrientOld, fig.env="figure*", fig.pos = "h", fig.align = "center", fig.width=6, fig.height=3, fig.cap = "Proportion of face / wrist detections by children's age, posture (top panel), and orientation (bottom panel)."}
# ages_inc_subs <- inc_subs %>%
#   select(subid, age.at.test, age.grp)
# 
# detectionsByOrientation <- dlongPlot %>%
#   filter(!is.na(orientation)) %>%
#   group_by(age.grp, orientation, subid, detectorType) %>%
#   summarise(prop_detected = mean(detection), num_detections = length(detection), detection_size = mean(w*h, na.rm=TRUE)) %>%
#   left_join(ages_inc_subs) %>%
#   mutate(age.at.test = as.numeric(age.at.test))
# 
# ggplot(detectionsByOrientation, aes(x = age.at.test, y = detection_size, col = detectorType, size=num_detections)) + 
#   geom_point(alpha=.5) + 
#   labs(y = "Prop. detected", x = "Age (months)") +
#   facet_grid(~detectorType) +
#   theme_few() +
#   theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
#   # scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6)) +
#   facet_grid(~orientation) +
#   geom_smooth(span=10, alpha=.2, aes(weight = num_detections), show.legend = FALSE)

```

```{r}
detectionsByPosture <- dlongPlot %>%
  filter(!is.na(posture)) %>%
  group_by(age.grp, posture, subid, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  left_join(ages_inc_subs) %>%
  mutate(age.at.test = as.numeric(age.at.test))

ggplot(detectionsByPosture, aes(x = age.at.test, y = prop_detected, col = detectorType, size=num_detections)) +
  geom_point(alpha=.5) +
  labs(y = "Prop. detected", x = "Age (months)") +
  facet_grid(~detectorType) +
  theme_few() +
  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
  scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6)) +
  facet_grid(~posture) +
  geom_smooth(span=10, alpha=.2, show.legend = FALSE)

```

```{r eval=FALSE}
p1=ggplot(detectionsByPosture, aes(x = factor(age.grp), y = face, col = posture)) + 
  geom_boxplot(outlier.size = 0.5) + 
  labs(y = "Prop. detected", x = element_blank()) +
  facet_grid(~detectorType) +
  theme_few() +
  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm")) +
  scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6))

detectionsByOrientation <- dlongPlot %>%
  filter(!is.na(orientation)) %>%
  group_by(age.grp, orientation, subid, detectorType) %>%
  summarise(detections = mean(detection)) %>%
  group_by(age.grp)

p2=ggplot(detectionsByOrientation, aes(x = factor(age.grp), y = detections, col = orientation)) + 
  geom_boxplot(outlier.size = 0.5) + 
  labs(y = "Prop. detected", x = element_blank()) +
  facet_grid(~detectorType) +
  theme_few() +
  theme(legend.position="right", legend.text=element_text(size = 8), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"))  +
  scale_y_continuous(breaks=seq(0,.60,.20), limits=c(0, .6))

ggarrange(p1, p2)

```

```{r}
# naming_rate_posorient <- aroundNamings_Faces %>%
#   group_by(subid, age.grp, orientation, posture) %>%
#   summarize(count_namings= length(naming.instance)) %>%
#   left_join(posture_by_orientation) %>%
#   filter(!is.na(posture)) %>%
#   filter(!is.na(orientation)) %>%
#   mutate(naming_rate = count_namings / time) %>%
#   left_join(ages_inc_subs) %>%
#   ungroup() %>%
#   mutate(subid = fct_reorder(subid, age.at.test))

# ggplot(naming_rate_posorient, aes(x = subid, y = naming_rate)) + 
#   geom_col(aes(fill = pos_orient), position=position_stack()) +
#   ylab("Objects named  / second ") + 
#   xlab("Individual subjects (ordered by age)") +
#   theme_few() +
#   theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
#   scale_fill_manual(name = "Posture & CG Orientation", values=c("#8878cc",  "#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc",  "#e67300","#ffa64d", "#ffe6cc","#d1b673")) 
# 
# ggplot(naming_rate_posorient, aes(x = age.grp, y = naming_rate, col=posture,size=prop.time)) + 
#   geom_boxplot(alpha=.4) +
#   ylab("Objects named  / second ") + 
#   # xlab("Age (months)") +
#   theme_few() +
#   theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
#   facet_grid(~age.grp)
#   scale_color_manual(name = "Posture & CG Orientation", values=c("#8878cc",  "#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc",  "#e67300","#ffa64d", "#ffe6cc","#d1b673"))

```

```{r}
namings_prop <- aroundNamings_Faces %>%
  group_by(subid, age.grp, orientation, posture) %>%
  summarize(count_namings=length(naming.instance)) %>%
  group_by(subid) %>%
  mutate(total_namings=sum(count_namings)) %>%
  group_by(subid, age.grp, orientation, posture) %>%
  mutate(prop_namings = count_namings / total_namings) %>%
  # filter(!is.na(posture)) %>%
  # filter(!is.na(orientation)) %>%
  left_join(ages_inc_subs) %>%
  ungroup() %>%
  mutate(subid = fct_reorder(subid, age.at.test)) %>%
  mutate(pos_orient = as.factor(paste0(posture,'_',orientation)))

ggplot(namings_prop, aes(x = subid, y = prop_namings)) + 
  geom_col(aes(fill = pos_orient), position=position_stack()) +
  ylab("Proportion naming events") + 
  xlab("Individual subjects (ordered by age)") +
  theme_few() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  scale_fill_manual(name = "Posture & CG Orientation", values=c("#8878cc","#ebebeb","#ebebeb", "#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc", "#98a69f", "#e67300","#ffa64d", "#ffe6cc","#d1b673")) 
```

```{r}
## does naming rate differ by posture?

## number of namimgs in each posture/orientation combo by sub
posture_orientation_naming_count <- aroundNamings_Faces %>%
  group_by(subid, posture, orientation) %>%
  summarize(num_namings = length(naming.instance))

## time spent in each posture/orient combo for each sub
posture_orientation_time <- d %>%
  group_by(subid, posture, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE))

naming_rate_by_posture <- posture_orientation_naming_count %>%
  left_join(posture_orientation_time) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  mutate(naming_rate = num_namings / time) %>%
  group_by(posture, orientation) %>%
  multi_boot_standard(col = 'naming_rate')

## nope, not really
ggplot(naming_rate_by_posture, aes(x=posture, y=mean, col=orientation)) +
  geom_pointrange(aes(ymin=ci_lower, ymax=ci_upper), position=position_dodge(width=1)) +
  theme_few()
```

```{r}
countNamings_Raw <- aroundNamings_Faces %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE), num_namings=length(detections)) %>%
  left_join(ages_inc_subs)

countNamings_2 <- countNamings_Raw %>%
  mutate(pos_orient = as.factor(paste0('Infant ',posture, ', CG ', orientation))) %>%
  left_join(ages_inc_subs) %>%
  ungroup() %>%
  mutate(subid = fct_reorder(subid, age.at.test))

ggplot(countNamings_2, aes(x = subid, y = num_namings)) + 
  geom_col(aes(fill = pos_orient), position=position_stack()) +
  ylab("Number of naming events") + 
  xlab("Individual subjects (ordered by age)") +
  theme_few() +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank()) +
  scale_fill_manual(name = "Posture & CG Orientation", values=c("#8878cc", "#ebebeb", "#ebebeb", "#006080", "#0099cc", "#99e6ff","#87b5c4", "#339966","#66cc99", "#b3e6cc",  "#e67300","#ffa64d", "#ffe6cc","#d1b673"))

## more naming events wtih age

```