### 2020-08-12 analysis/4_extra_plots/SI_extraplots.Rmd
port old plotting code to backup

---
title: "SI_ExtraPlots"
author: "Bria Long"
date: "4/8/2020"
output: pdf_document
---

```{r, libraries}
library(knitr)
knitr::opts_chunk$set(fig.width=6, fig.height=5, fig.crop = FALSE, 
                      fig.path='figs/', echo=FALSE, warning = FALSE, 
                      cache=TRUE, message = FALSE, sanitize = TRUE)
library(readr)
library(dplyr)
library(tidyr)
library(xtable)
library(lubridate)
library(langcog)
library(ggplot2)
library(ggthemes)
library(lme4)
library(gridExtra)
library(ggridges)
library(assertthat)
library(tidyverse)
library(viridis)
library(papaja)
```

```{r, message=FALSE, warning=FALSE, echo=FALSE, include=FALSE}
### READ IN DETECTIONS
d <- read_csv("../data/consolidated_data/consolidated_data_4detectors_march2020.csv") %>%
  mutate(posture = factor(posture), orientation = factor(orientation))

## make long form version for plotting both face detections on same graph
dlong <- d %>%
  mutate(id = paste(subid,frame)) %>%
  gather(key = detectorType, value = detection, faceMT, faceOP, faceVJ, wristOP) %>%
  mutate(detectorType=as.factor(detectorType))

# for better plotting names
levels(dlong$detectorType) 
# [1] "faceMT"  "faceOP"  "faceVJ"  "wristOP"
levels(dlong$detectorType) <- c("MTCNN-Faces", "OpenPose-Faces","ViolaJones-Faces","OpenPose-Wrists")

```

```{r}
source("helper.R")
## summarize detections around naming events
aroundNamings_Wrists <- d %>%
  group_by(subid) %>%
  mutate(detections = wristOP) %>%
  do(summarize.naming(.))

aroundNamings_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceOP) %>%
  do(summarize.naming(.)) 
```

```{r eval=FALSE,include=FALSE}

### SUMMARIES FOR SI PLOTS
### Faces/Wrists by naming by posture/orientation

## summarize faces/wrists detected by orientation/posture/naming events
aroundNamings_Faces_PosOrient_Group <- aroundNamings_Faces %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(propFacesDetected = mean(detections, na.rm=TRUE)) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  group_by(orientation, posture, familiarity, age.grp) %>%
  multi_boot_standard(col = "propFacesDetected")

aroundNamings_Wrists_PosOrient_Group <- aroundNamings_Wrists %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(propWristsDetected = mean(detections, na.rm=TRUE)) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  group_by(orientation, posture, familiarity, age.grp) %>%
  multi_boot_standard(col = "propWristsDetected")

## for plotting
aroundNamings_Faces_PosOrient_Raw <- aroundNamings_Faces %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE), num_namings= length(detections)) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  left_join(ages_inc_subs)

aroundNamings_Wrists_PosOrient_Raw <- aroundNamings_Wrists %>%
  group_by(subid, age.grp, orientation, posture, familiarity) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE), num_namings=length(detections)) %>%
  filter(!is.na(orientation)) %>%
  filter(!is.na(posture)) %>%
  left_join(ages_inc_subs)

```

```{r eval=FALSE,include=FALSE}

### SI PLOT 1
### Faces by naming by posture/orientation

# plots means by familairity and posture
ggplot(aroundNamings_Faces_PosOrient_Group, aes(x=age.grp, y=mean, color=familiarity)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper, color=familiarity), position = position_dodge(width=3), alpha=.9) +
    # plot indiv data points
  geom_point(data=aroundNamings_Faces_PosOrient_Raw, aes(x=age.grp, y=prop_detected, size=num_namings, color=familiarity), position = position_dodge(width=3), alpha=.2) +
  # # plot baseline
    geom_pointrange(data = (detectionsByPostureAndOrientByGroup %>% filter(detectorType == 'OpenPose-Faces')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=0), alpha=.9) +
  theme_few() +
  ylim(0,.6) +
  facet_grid(orientation ~ posture) +
  labs(y = "Prop. face detections around naming events", x = "Age (months)") +
  theme_few(base_size=16) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")
```
```{r eval=FALSE,include=FALSE}

### SI PLOT 2
### Wrists by naming by posture/orientation

# plots means by familairity and posture
ggplot(aroundNamings_Wrists_PosOrient_Group, aes(x=age.grp, y=mean, color=familiarity)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper, color=familiarity), position = position_dodge(width=3), alpha=.9) +
    # plot indiv data points
  geom_point(data=aroundNamings_Wrists_PosOrient_Raw, aes(x=age.grp, y=prop_detected, size=num_namings, color=familiarity), position = position_dodge(width=3), alpha=.2) +
  # # plot baseline
    geom_pointrange(data = (detectionsByPostureAndOrientByGroup %>% filter(detectorType == 'OpenPose-Wrists')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=0), alpha=.9) +
  theme_few() +
  ylim(0,.6) +
  facet_grid(orientation ~ posture) +
  labs(y = "Prop. wrist detections around naming events", x = "Age (months)") +
  theme_few(base_size=16) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")
```

```{r eval=FALSE}
### Parallel plot to detections -- not included
# better labels 
orientation_labels <- c(
  behind ='Caregiver behind',
  close = 'Caregiver close',
  far = 'Caregiver far')

posture_labels = c(
  prone = 'Infant prone',
  sit = 'Infant sitting',
  stand ='Infant standing'
)
posture_by_orientation_subset <- d %>%
  filter(!is.na(posture)) %>%
  filter(!is.na(orientation)) %>%
  group_by(subid, posture, orientation) %>%
  summarise(time = sum(dt, na.rm=TRUE)) %>%
  group_by(subid) %>%
  mutate(prop.time = time/sum(time)) %>%
  left_join(ages_inc_subs) %>%
  mutate(pos_orient = paste0('Infant ',posture, ', CG ', orientation)) # for colors

ggplot(posture_by_orientation_subset, aes(x = age.at.test, y = prop.time, col=pos_orient, size=time)) +
  geom_point(alpha=.9) +
  ylab("Proportion Time") +
  xlab("Age (months)") +
  theme_few(base_size=18) +
  ylim(0,1) +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position='none') +
  scale_color_manual(name = "Posture & CG Orientation", values=c("#006080", "#0099cc", "#99e6ff","#339966","#66cc99", "#b3e6cc",  "#e67300","#ffa64d", "#ffe6cc")) +
  facet_grid(orientation ~ posture, labeller = labeller(posture = posture_labels, orientation= orientation_labels)) +
  geom_smooth(span=10, alpha=.2)

```

```{r afterNamingsExplore, eval=FALSE, include=FALSE, echo=FALSE}
afterNamings_Faces <- d %>%
  group_by(subid) %>%
  mutate(detections = faceOP) %>%
  do(summarize.naming((.),window = c(0,4)))

detectionsByGroup <- dlongPlot %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection), num_detections = length(detection)) %>%
  group_by(age.grp, detectorType) %>%
  multi_boot_standard(col = 'prop_detected')

afterNamings_Detections_BySub_Faces <- afterNamings_Faces %>%
  group_by(subid, familiarity, age.grp) %>%
  summarize(prop_detected = mean(detections, na.rm=TRUE))

afterNamings_Detections_Grouped_Faces <- afterNamings_Detections_BySub_Faces %>%
  group_by(familiarity, age.grp) %>%
  multi_boot_standard(col = 'prop_detected')

faces_afternamings_plot <- ggplot(afterNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
  # plot baseline
  geom_pointrange(data = (detectionsByGroup %>% filter(detectorType == 'OpenPose-Faces')), aes(ymin = ci_lower, ymax = ci_upper, color='baseline'), position = position_dodge(width=3), alpha=.9) +
  # plot averages by familiarity 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
  # plot familairity detections for each sub
  geom_point(data=afterNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
  theme_few() +
  labs(y = "Prop. face detections", x = "Age (months)") +
  theme_few(base_size=plot_size) +
  ylim(0,.6) +
  theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
  scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
  scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) +
  scale_size_continuous(name="# of naming events")

```

```{r}
### get individual points
baseline_diffs_faces <- dlong %>%
  group_by(age.grp, subid, detectorType) %>%
  summarise(prop_detected = mean(detection)) %>%
  mutate(familiarity = 'baseline') %>%
  full_join(aroundNamings_Detections_BySub_Faces) %>%
  spread(familiarity,  prop_detected) %>%
  summarize(fam_vs_baseline = familiar - baseline, novel_vs_baseline = novel - baseline)

baseline_by_group <-  baseline_diffs_faces %>%
  group_by(age.grp) %>%
  filter(!is.na(novel_vs_baseline)) %>%
  multi_boot_standard(col='novel_vs_baseline')

ggplot(baseline_by_group, aes(x=age.grp, y=mean)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=3), alpha=.9) +
  # plot familairity detections for each sub
  geom_point(data=afterNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=3), alpha=.2) +
  theme_few() +
```

### Analyze detections around naming events (without posture/orientation)

```{r include=FALSE}
# difference plots can be confusing, ommitting
# faces_namings_diffs_indiv <- aroundNamings_Detections_BySub_Fam_Faces_All %>%
#   spread(familiarity, prop_detected) %>%
#   mutate(fam_dif = familiar - baseline, nov_diff = novel - baseline) 
#
# ggplot(faces_namings_diffs_indiv, aes(x=age.grp, y=nov_diff, group=age.grp)) +
#   geom_boxplot() +
#   theme_few()

## explored looking at namings while acconting for posture and orientation
# aroundNamings_Detections_BySub_Posture_Faces <- aroundNamings_Faces %>%
#   group_by(subid, familiarity, posture, orientation, age.grp) %>%
#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
#   mutate(detectorType = 'OpenPose-Wrists')
# 
# baseline_diffs_by_posture <- dlong %>%
#   filter(detectorType == 'OpenPose-Wrists') %>%
#   group_by(age.grp, subid, posture, orientation,detectorType) %>%
#   summarise(prop_detected = mean(detection)) %>%
#   mutate(familiarity = 'baseline') %>%
#   full_join(aroundNamings_Detections_BySub_Posture_Faces)
# 
# naming = lmer(data = aroundNamings_Detections_BySub_Posture_Faces, prop_detected ~ familiarity + posture * orientation + (1|subid))
# summary(naming)
```

```{r setUpNamingsPlot}

## first get baseline
# baselineFaces <- dlongPlot %>%
#   filter(detectorType == 'OpenPose-Faces') %>%
#   group_by(age.grp, subid, detectorType) %>%
#   summarise(prop_detected = mean(detection)) %>%
#   mutate(familiarity='baseline')
# 
# baselineWrists <- dlongPlot %>%
#   filter(detectorType == 'OpenPose-Wrists') %>%
#   group_by(age.grp, subid, detectorType) %>%
#   summarise(prop_detected = mean(detection)) %>%
#   mutate(familiarity='baseline')
# 
# 
# ### then by faces
# aroundNamings_Detections_BySub_Faces <- aroundNamings_Faces %>%
#   group_by(subid, familiarity, age.grp) %>%
#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
#   mutate(detectorType = 'OpenPose-Faces') %>%
#   full_join(baselineFaces)
# 
# aroundNamings_Detections_Grouped_Faces <- aroundNamings_Detections_BySub_Faces %>%
#   group_by(familiarity, age.grp) %>%
#   multi_boot_standard(col = 'prop_detected')  
# 
# ### now by wrists 
# aroundNamings_Detections_BySub_Wrists <- aroundNamings_Wrists %>%
#   group_by(subid, familiarity, age.grp) %>%
#   summarize(prop_detected = mean(detections, na.rm=TRUE)) %>%
#   mutate(detectorType = 'OpenPose-Wrists') %>%
#   full_join(baselineWrists)
# 
# aroundNamings_Detections_Grouped_Wrists <- aroundNamings_Detections_BySub_Wrists %>%
#   group_by(familiarity, age.grp) %>%
#   multi_boot_standard(col = 'prop_detected')

```

```{r detByNamingPlotPanels}
## Plot faces/wrists around namings by age

# faces_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Faces, aes(x=age.grp, y=mean, color=familiarity)) +
#   # plot averages by familiarity 
#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=2), alpha=.9) +
#   # plot familairity detections for each sub
#   geom_point(data=aroundNamings_Detections_BySub_Faces, aes(x=age.grp, y=prop_detected), position = position_dodge(width=2), alpha=.2) +
#   theme_few() +
#   labs(y = "Prop. face detections", x = "Age (months)") +
#   theme_few(base_size=plot_size) +
#   ylim(0,.4) +
#   theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
#   scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
#   scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18))

## 
# wrists_aroundnamings_plot <- ggplot(aroundNamings_Detections_Grouped_Wrists, aes(x=age.grp, y=mean, color=familiarity)) +
#   # plot averages by familiarity 
#   geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), position = position_dodge(width=2), alpha=.9) +
#   # plot familairity detections for each sub
#   geom_point(data=aroundNamings_Detections_BySub_Wrists, aes(x=age.grp, y=prop_detected), position = position_dodge(width=2), alpha=.2) +
#   theme_few() +
#   labs(y = "Prop. wrist detections", x = "Age (months)") +
#   theme_few(base_size=plot_size) +
#   ylim(0,.4) +
#   theme(legend.position="bottom", legend.text=element_text(size = 12), legend.margin=margin(t=0, r=0, b=0, l=0, unit="cm"), legend.title=element_text(size = 14)) +
#   scale_colour_viridis(name = 'Object familiarity', discrete=TRUE, option='A', begin=.2, end=.8) +
#   scale_x_continuous(breaks=seq(8, 16, 4), limits=c(7, 18)) 
```